---
title: "NEDLAME Alert Pilot: TX (Trim) vs Control"
editor: source
format:
 html:
   embed-resources: true
   toc: true
   toc-location: left
   code-fold: true
execute:
    echo: false
    message: false
    warning: false
params:
  milk_window_pre: 14
  milk_window_post: 30
  milk_baseline_days: 7
  q4_lookback_days: 21
  history_after_date: "2026-06-01"
  recent_trim_suppression_days: 28
  post_trim_exclusion_days: 7
  analysis_end_date: "2026-08-27"
---

```{r}
#| label: setup

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  arrow,
  lubridate,
  gt,
  gtsummary,
  survival,
  survminer,
  ggsurvfit,
  scales,
  viridis
)

source(file.path('functions/fxn_read_milk_folder.R'))
source(file.path('functions/fxn_read_nedap_attentions.R'))

# GC's lame functions (fxn_collapse_lesions, fxn_trim_vars, fxn_dz_status) -
# fetched live from GitHub, same mechanism report_explore_lame_new.qmd uses.
# Two local overrides are sourced AFTER fxn_load_os_fxns() specifically so
# they win over the GitHub versions in the global environment:
#  - functions/fxn_code_lesions.R: the GitHub version's `trimonly` logic
#    only checks the *protocols* field for "NONE" - this farm's
#    routine/no-lesion trims are coded "NONE" in the *remark* field
#    instead, which the GitHub version doesn't check, so it silently
#    misclassifies every routine trim as having a lesion.
#  - functions/fxn_collapse_lesions.R: the GitHub version reassigns its
#    own `data` variable to an already-one-row-per-date subset before
#    computing the max() union across that date's lesion columns, so
#    when a cow has multiple event rows on the same date coding
#    DIFFERENT lesions, only whichever row survives the earlier
#    reduction keeps its lesion flags - the rest are silently dropped
#    (confirmed against a real cow with 4 same-day LAME rows: only 1 of
#    her 3 distinct lesion types survived under the GitHub version).
source(file.path('functions/fxn_load_os_fxns.R'))
fxn_load_os_fxns()
source(file.path('functions/fxn_code_lesions.R'))
source(file.path('functions/fxn_collapse_lesions.R'))

events_all <- read_parquet('data/intermediate_files/events_all_columns.parquet')
events_formatted <- read_parquet('data/intermediate_files/events_formatted.parquet')
animals <- read_parquet('data/intermediate_files/animals.parquet')
animal_lactations <- read_parquet('data/intermediate_files/animal_lactations.parquet')

date_max_pull <- max(events_all$date_event, na.rm = TRUE)
farm_name <- unique(events_all$location_event)

# NEDLAME go-live date, per the DairyComp setup: enrollment protocols were
# only turned on starting ~June 1 2026, so no NEDLAME alert can exist
# before this regardless of how "missed" a case looks.
nedlame_start_date <- min(events_all$date_event[events_all$Event == "NEDLAME"], na.rm = TRUE)
```

```{r}
#| label: cohort

# One row per cow (id_animal) that ever got a NEDLAME alert - grouped by
# id_animal alone since a cow only has one "first" NEDLAME ever. Treatment
# group is read off the MNFRS value on the FIRST NEDLAME event for that cow
# - MNFRS is stable across a cow's NEDLAME rows even though it drifts over
# her broader event history (0 before enrollment, 1/2 once enrolled).
# Hard analysis cut-off. The raw Nedap sensor export ends on
# params$analysis_end_date, so alerts and lesion cases after it cannot be
# checked against what the camera actually saw. Rather than carry a partly
# checkable tail, entry into the analysis stops there for BOTH alerts (here)
# and lesion cases (Q4). Follow-up events are deliberately NOT capped - a
# cow who entered on/before the cut-off keeps every trim, cull and milk
# record recorded after it, which is what time-to-event needs.
analysis_end <- ymd(params$analysis_end_date)

first_nedlame_raw <- events_all |>
  filter(Event == "NEDLAME", date_event <= analysis_end) |>
  group_by(id_animal) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  transmute(
    id_animal, id,
    first_nedlame_date = date_event,
    alert_type = Remark,
    tx_group = case_when(
      MNFRS == "1" ~ "TX (Trim)",
      MNFRS == "2" ~ "Control",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(tx_group))

# lact_number is deliberately NOT taken from the raw NEDLAME event's own
# field - right at a freshening transition (dry -> fresh, same day or one
# day apart), DC's raw event can carry a stale lact_number from the
# lactation that just ended rather than the one that just started (found
# in 2 of 718 cows once the dataset grew past ~700 alerts: their alert
# date fell after date_next_fresh/on date_fresh of the FOLLOWING
# lactation, yet the raw event said the OLD lact_number). Resolving to
# whichever lactation was actually active as of the alert date - the same
# nearest-preceding-fresh-date pattern already used for the milk join -
# fixes this without needing to trust the raw field.
first_nedlame <- first_nedlame_raw |>
  inner_join(
    animal_lactations |>
      filter(!is.na(date_fresh)) |>
      select(id_animal, lact_number, date_fresh, date_dry, date_next_fresh),
    by = "id_animal",
    relationship = "many-to-many"
  ) |>
  filter(date_fresh <= first_nedlame_date) |>
  group_by(id_animal) |>
  slice_max(date_fresh, n = 1, with_ties = FALSE) |>
  ungroup()

# Flag cows whose FIRST NEDLAME fired 1-`r params$post_trim_exclusion_days`
# days AFTER a recorded trim, same cow-lactation - per the farm's
# confirmed data-load schedule (NEDLAME's daily batch loads at ~5am and
# stamps each alert with the day it was LOADED, not necessarily the day
# the underlying sensor attention happened - an attention any time after
# ~5am on day D gets logged with date D+1), an alert dated on or before
# the trim date is a genuine independent catch, but one dated 1+ days
# AFTER the trim is very likely the camera picking up that cow's own
# post-trim gait change, not a new problem. It got through DC's own
# FTDAT suppression gate only because the trim record itself is commonly
# entered a day or two late (backdated to the actual trim date) - it
# simply wasn't in the system yet when NEDLAME's batch ran. See the new
# "Excluded" section below for the full writeup and examples.
#
# "Trim" here means LAME, FOOTRIM, *or* TRIM - same definition already
# established for Q4's FTDAT suppression gate (Round 9): a LAME visit
# IS the trim whenever the trimmer finds and codes a lesion at that
# same encounter, it just doesn't also get a separate FOOTRIM code that
# day. An earlier version of this exact chunk only checked
# FOOTRIM/TRIM, which missed exactly this case - cow 7207's LAME (White
# Line) visit on 06/11 was her actual trim, and NEDLAME fired the very
# next day (confirmed directly against her real cowcard).
trims_all <- events_all |>
  filter(Event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, trim_date = date_event) |>
  distinct()

post_trim_check <- first_nedlame |>
  select(id_animal, lact_number, first_nedlame_date) |>
  inner_join(trims_all, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(trim_date < first_nedlame_date, trim_date >= first_nedlame_date - params$post_trim_exclusion_days) |>
  group_by(id_animal, lact_number) |>
  summarize(days_since_nearest_trim = min(as.numeric(first_nedlame_date - trim_date)), .groups = "drop")

first_nedlame <- first_nedlame |>
  left_join(post_trim_check, by = c("id_animal", "lact_number")) |>
  mutate(flagged_post_trim = !is.na(days_since_nearest_trim))

# Attach lactation-stage context (date_fresh/date_dry) so every downstream
# join can be scoped to id_animal + lact_number, never crossing into a
# different lactation for the same cow.
all_nedlame_cows <- first_nedlame |>
  left_join(animals |> select(id_animal, date_left), by = "id_animal") |>
  mutate(
    dim_at_nedlame = as.numeric(first_nedlame_date - date_fresh),
    date_lact_end = pmin(
      coalesce(date_dry, date_next_fresh, date_max_pull),
      date_max_pull,
      na.rm = TRUE
    ),
    # Observation actually ends when the cow leaves the herd, if that comes
    # first: a culled/dead cow can't have a subsequent trim, so counting her
    # as still at risk inflates the denominator and understates the event
    # rate. date_lact_end alone ignored this. pmax() floors it at her alert
    # date so a data error can't produce negative follow-up.
    date_obs_end = pmax(
      pmin(date_lact_end, coalesce(date_left, date_max_pull), na.rm = TRUE),
      first_nedlame_date
    )
  )

# `cohort` (used throughout Q1-Q3, Q5, Q6) excludes the flagged cows;
# `cohort_excluded_post_trim` keeps them, visibly, for the "Excluded"
# section - nothing is silently dropped. Q4 is deliberately NOT scoped
# to `cohort` (it pulls NEDLAME/lesion history from the whole herd
# directly), so this exclusion does not affect Q4's miss-rate numbers.
cohort <- all_nedlame_cows |> filter(!flagged_post_trim)
cohort_excluded_post_trim <- all_nedlame_cows |> filter(flagged_post_trim)

n_tx <- sum(cohort$tx_group == "TX (Trim)")
n_control <- sum(cohort$tx_group == "Control")
n_excluded_post_trim <- nrow(cohort_excluded_post_trim)
```

```{r}
#| label: lesion-history-dataset

# Same lesion-coding / chronicity pipeline as report_explore_lame_new.qmd,
# just kept at the lactation grain (lact_number carried through) so the
# "history at enrollment" lookup below never crosses lactations.
all_lesions <- c("dd", "footrot", 'wld', "sole_ulcer", "injury",
                  "cork", "other", "hemorrhage",
                  "sole_fracture", "toe_ulcer", "thin",
                  "inf", "noninf", "toe", "axial", "lesion")
diseases <- c("inf", "noninf", "lesion")

lame_data <- events_formatted |>
  select(id_animal, lact_number, date_event, event,
         remark_letters1, protocols_letters1, location_event, protocols) |>
  filter(event %in% c("LAME", "TRIM", "FOOTRIM")) |>
  fxn_code_lesions(protocol_var = protocols_letters1, remark_var = remark_letters1) |>
  filter(!is.na(trimonly)) |>
  mutate(
    noninf = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 1,
      .default = noninf),
    other = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 0,
      .default = other)
  ) |>
  fxn_collapse_lesions(lesions = all_lesions) |>
  fxn_trim_vars(location_event, id_animal, date_var = date_event, trimonly_var = trimonly) |>
  fxn_dz_status(disease_cols = diseases, event_filter = trimonly == 0)

# Each NEDLAME cow's lameness-history label as of (at/before) her first
# NEDLAME date, using her FULL lifetime record (all lactations - lame_data
# goes back to 2021), not scoped to her current lactation. Unlike every
# other join in this report, this one is intentionally NOT lactation-scoped
# per explicit user direction: chronicity is a lifetime concept, and the
# ~5 years of history available should be used rather than artificially
# resetting it at each freshening.
#
# IMPORTANT: candidate history rows are filtered to "at/before enrollment"
# BEFORE joining to cohort, not after (same reasoning as fxn_time_to_next()
# above) - filtering after a left_join would silently DROP a cow from
# `cohort` entirely instead of correctly labeling her. An earlier version
# of this exact chunk had that bug and was quietly dropping 49 of 434 cows
# (~11%) from the ENTIRE cohort - not just from Q5, but from every section,
# since `cohort` is the shared base table.
history_lookup <- lame_data |>
  select(id_animal, date_event, status_lesion) |>
  filter(!is.na(status_lesion))

valid_history <- history_lookup |>
  inner_join(
    cohort |> select(id_animal, lact_number, first_nedlame_date),
    by = "id_animal"
  ) |>
  filter(date_event <= first_nedlame_date) |>
  select(id_animal, lact_number, date_event, status_lesion)

# A cow with no LESION ever diagnosed before her NEDLAME alert has no
# status_lesion to inherit - fxn_dz_status() only assigns a status to
# trimonly == 0 rows (an actual lesion found), so a cow's routine
# trim-only visits (trimonly == 1) never produce a status_lesion, no
# matter how many she's had. This is NOT the same as "no LAME/TRIM/
# FOOTRIM event at all" - checked directly, most of these cows (206 of
# 353 in the current cohort) DO have prior trim/LAME events, just never
# one where a lesion was found; only a small remainder (~21 of 353)
# truly have zero foot-care events before their alert. Rather than
# invent a 4th bucket for either case, she's folded into "New" - the
# same label her eventual first-ever diagnosis would get, and the
# closest existing category to "no established chronic pattern."
cohort <- cohort |>
  left_join(valid_history, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  group_by(id_animal, lact_number) |>
  slice_max(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
  ungroup() |>
  mutate(history_group = if_else(is.na(status_lesion), "New", status_lesion)) |>
  select(-date_event, -status_lesion)

history_levels <- c("New", "Repeat", "Chronic")
cohort <- cohort |>
  mutate(history_group = factor(history_group, levels = history_levels))
```

```{r}
#| label: milk-data

# data/milk/ is not wired into any existing pipeline script. Raw milk rows
# only carry a raw `id` (not the composite id_animal), so they're resolved
# to a specific animal-lactation by nearest-preceding date_fresh (see
# fxn_read_milk_folder.R for why a naive date_dry window isn't safe here).
milk_long <- fxn_read_milk_folder()
milk_resolved <- fxn_resolve_milk_to_lactation(milk_long, animal_lactations)
```

```{r}
#| label: nedap-attentions-data

# Raw Nedap sensor export - what the SENSOR actually flagged, independent
# of whether that flag ever reached DairyComp as a NEDLAME event. Used in
# Q4 to split "NEDAP missed it" cases into true sensor misses vs pipeline
# losses (a real cowcard/Pulse cross-check first surfaced this pattern -
# see the plan doc's Round 8/10). This is a one-time export, not a live
# feed, so its coverage window is whatever was current when it was
# pulled - captured here explicitly rather than assumed, since it can
# lag behind the DairyComp/milk data read in above.
attentions_raw <- fxn_read_nedap_attentions()
attentions_resolved <- fxn_resolve_nedap_to_lactation(attentions_raw, animal_lactations)

nedap_coverage_start <- min(attentions_resolved$attention_date, na.rm = TRUE)
nedap_coverage_end <- max(attentions_resolved$attention_date, na.rm = TRUE)
```

```{r}
#| label: platform-enrollment-data

# Nedap runs two separate platform instances at this farm - "Activity"
# (used by farm staff daily) and "SmartSight" (mobility monitoring,
# generates NEDLAME) - kept separate at the farm's request to avoid
# confusion when checking the Nedap platform. A cow absent from
# SmartSight's animal list was never being monitored for lameness at
# all, regardless of what the raw attentions export's date coverage
# shows - this check is animal-level enrollment, not date-bound, so it
# can resolve cases the date-bound attentions export above can't (see
# the Platform Enrollment Check subsection in Q4).
cows_activity <- read_csv("data/cows_activity.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

cows_smartsight <- read_csv("data/cows_smartsight.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

# Animal numbers get reused and reassigned. DairyComp records an ID change as
# an XID event whose Remark holds the OLD number, and 150 of those old numbers
# are still present in this extract as animals in their own right. A cow
# renumbered after she was enrolled therefore looks ABSENT from SmartSight
# under her current number while being present under her old one - confirmed
# on cow 23143, renumbered from 14347 on 2026-08-06 and enrolled all along as
# 14347. Checking the current number alone would call her unmonitored.
id_aliases <- events_all |>
  filter(Event == "XID") |>
  transmute(id = as.character(id), alias = trimws(as.character(Remark))) |>
  filter(!is.na(alias), grepl("^[0-9]+$", alias)) |>
  distinct()

smartsight_ids <- unique(cows_smartsight$id)
# a cow counts as enrolled if her current OR any prior number is on the list
enrolled_ids <- union(
  smartsight_ids,
  id_aliases$id[id_aliases$alias %in% smartsight_ids]
)
n_rescued_by_alias <- length(setdiff(
  id_aliases$id[id_aliases$alias %in% smartsight_ids], smartsight_ids))
```

# Overview

To determine if using the camera benefited cows we set up a study that trimmed 1/2 the cows that the camera flagged and left the other group to be trimmed if your farm staff found them or if they showed up on the trim list.

- Nedap flags cows in 3 ways and the setup in DairyComp section describes the flags and how this was handled in DairyComp

To compare the groups we looked at the following outcomes after being flagged:

- Milk Production

- Culling

- Cows with lesions not in the treatment groups that the camera missed

## Setup in DairyComp

For this study the follow things were created in DairyComp to handle the file data from NEDAP daily.

- A command to clear all filein items before new data was brought in as a workaround to deal with historical flag issues

- New NEDAP lame event with 3 protocols to flag cows and criteria to create them

  - Low (score \<30 today, not trimmed or an alarm in last 90 days not a cow with history of upper leg lameness)

    - ENTER EC=70\\P215 FOR LFTYP=30102 LMFLG=1 LMSV=1-30 DSNLM=90-1 FTDAT\<-90 UPLAM\<\>1

  - Low Declining (less than 30 score, declining score and and no alarm in last 7 days and not trimmed in last 28 days)

    - ENTER EC=70\\P214 FOR LFTYP=30101 LMFLG=1 DSNLM=7-1 FTDAT\<-28

  - Strong Decline (any score and a strong decline in scores with no alarm in last 7 days, not trimmed in last 28 days) second OR statement allows cows with a low alert already to get another alert

    - ENTER EC=70\\P213 FOR LFTYP=30100 LMFLG=1 (DSNLM=7-1 FTDAT\<-28)(FTDAT\<-28 %73.NDLAME.101\>LOW)

- Cows were added to trim list with a command that uses an existing item (MNFRS left over from Luciano's work) that randomizes cows by lactation group.

  - Enter FTDAT=1.1.1 for MNFRS=1 %.70NEDLAME.101=TODAY

Above commands were all run daily around 5 am after NEDAP file-in occurred.

## Results

- `r nrow(first_nedlame)` cows received a `NEDLAME` alert between `r format(min(first_nedlame$first_nedlame_date))` and `r format(max(first_nedlame$first_nedlame_date))`

- `r n_excluded_post_trim` were excluded as alerts occurred after a recent hoof trimming.

  - These were still coded in DairyComp due to the delay in importing hoof trimming data a few days after the actual hoof trimming date.

This results in

- `r n_tx` cows enrolled in the **TX (Trim)** arm (`MNFRS = 1`, automatically trimmed after alert)

- `r n_control` in the **Control** arm (`MNFRS = 2`, staff decide whether/when to act or trimmed because of regular trim list),

# Description of Cows

The tables below describe the cows enrolled by treatment group

## By Alert Type

::: panel-tabset
```{r}
#| label: overview-table

cohort |>
  count(tx_group, alert_type) |>
  pivot_wider(names_from = alert_type, values_from = n, values_fill = 0) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Most cows were flagged as Low Declines") |>
  tab_stubhead(label = "Treatment Group")
```

## By Lactation Number

```{r}
#| label: lact-group-table

cohort |>
  mutate(lact_group_display = case_when(
    lact_number == 1 ~ "1",
    lact_number == 2 ~ "2",
    lact_number >= 3 ~ "3+",
    TRUE ~ NA_character_
  )) |>
  count(tx_group, lact_group_display) |>
  pivot_wider(names_from = lact_group_display, values_from = n, values_fill = 0) |>
  select(tx_group, `1`, `2`, `3+`) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Most cows enrolled were Lact 3+",
    ) |>
  tab_stubhead(label = "Treatment Group") |>
  cols_label(`3+` = "3+")
```

## By Lactation Stage (DIM)

```{r}
#| label: dim-balance-table

cohort |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_dim = round(mean(dim_at_nedlame, na.rm = TRUE), 0),
    median_dim = round(median(dim_at_nedlame, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Similar DIM between groups",
    ) |>
  cols_label(n = "N Cows", mean_dim = "Mean DIM", median_dim = "Median DIM")
```

```{r}
#| label: q1-dim-check
#| fig-height: 4

cohort |>
  ggplot(aes(x = tx_group, y = dim_at_nedlame, fill = tx_group)) +
  geom_violin(alpha = 0.5, color = NA, trim = FALSE) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.4, size = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "DIM at 1st NEDLAME Alert")
```
:::

# Trimming

## Time to next trimming after alert

To determine the accuracy of the alert set up for each cow, we looked for the timing of the first trimming event after her first NEDLAME alert.

The figures below show the time to the next trimming event split by treatment group and alerts types.

- It is clear that most of the TRIM treatment cows got trimmed within 7-10 days and about 20% of the cows in the Control group get trimmed withing that same time period.
- Over time the control group reaches about 75% of cows trimmed.

```{r}
#| label: q2-build

next_events <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event, event) |>
  distinct()

lesion_lookup <- lame_data |> select(id_animal, lact_number, date_event, event, lesion)

# IMPORTANT: candidate events are filtered to the valid post-NEDLAME window
# BEFORE joining to cohort, not after. Filtering after a left_join would
# silently DROP a cow entirely (rather than correctly censoring her)
# whenever she has trim/LAME events somewhere in her history but none
# inside the valid window - an earlier version of this code had exactly
# that bug and was quietly excluding ~45% of cows from the "any event"
# curve (and a different ~25% from the "with lesion" curve), which is
# also why the two curves looked identical - they weren't being computed
# over the same denominator.
fxn_time_to_next <- function(candidate_events) {
  valid_candidates <- candidate_events |>
    inner_join(
      cohort |> select(id_animal, lact_number, first_nedlame_date, date_obs_end),
      by = c("id_animal", "lact_number")
    ) |>
    filter(date_event > first_nedlame_date, date_event <= date_obs_end) |>
    select(id_animal, lact_number, date_event)

  cohort |>
    left_join(valid_candidates, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
    group_by(id_animal, lact_number) |>
    slice_min(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
    ungroup() |>
    mutate(
      event_occurred = as.integer(!is.na(date_event)),
      days_to_event = if_else(
        event_occurred == 1,
        as.numeric(date_event - first_nedlame_date),
        as.numeric(date_obs_end - first_nedlame_date)
      )
    )
}

q2_any <- fxn_time_to_next(next_events)

next_events_lesion <- next_events |>
  left_join(lesion_lookup, by = c("id_animal", "lact_number", "date_event", "event")) |>
  filter(!is.na(lesion) & lesion == 1)

q2_lesion <- fxn_time_to_next(next_events_lesion)
```

::: panel-tabset
## Time to Any Trim/LAME Event

```{r}
#| label: q2-any-plot
#| fig-height: 5

km_any <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_any)

km_any |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-any-table

km_any |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Trim/LAME Event")
```

## By Alert Type

Decline cows mostly all get trimmed

```{r}
#| label: q2-alert-type-plot
#| fig-height: 6

alert_levels <- c("LOW", "LOW DECLINE", "STRONG DECLINE")

# ggsurvfit only exposes a single combined `strata` column, so a KM plot
# faceted by alert_type is built manually: fit one survfit2() per
# alert level (skipping any level without both treatment arms
# represented) and stack the tidy() estimates before plotting.
fittable_alert_levels <- alert_levels |>
  keep(\(a) {
    d <- q2_any |> filter(alert_type == a)
    n_distinct(d$tx_group) == 2 && nrow(d) >= 10
  })

q2_alert_type_km_data <- fittable_alert_levels |>
  map(\(a) {
    d <- q2_any |> filter(alert_type == a)
    fit <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = d)
    broom::tidy(fit) |>
      mutate(
        tx_group = str_remove(strata, "^tx_group="),
        alert_type = a,
        risk = 1 - estimate,
        # survival's CI bounds flip when converted to risk (1 - survival):
        # the upper bound on survival becomes the LOWER bound on risk.
        risk_low = 1 - conf.high,
        risk_high = 1 - conf.low
      )
  }) |>
  list_rbind() |>
  mutate(alert_type = factor(alert_type, levels = alert_levels))

skipped_alert_levels <- setdiff(alert_levels, fittable_alert_levels)

ggplot(q2_alert_type_km_data, aes(x = time, y = risk, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = risk_low, ymax = risk_high), alpha = 0.15, color = NA) +
  geom_step(linewidth = 1) +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event", color = "Treatment Group", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```

`r if (length(skipped_alert_levels) > 0) paste0("*Alert type(s) omitted for insufficient data in one or both arms: ", paste(skipped_alert_levels, collapse = ", "), ".*")`
:::

## How many trims with no lesions?

The table below shows the number of cows who had a first NEDAP alert and where trimmed **but** did not have a lesions.

**This table's population is only the number of cows in the `r nrow(cohort)`-cow NEDLAME cohort that were trimmed.**

Overall staff identified cows have fewer cows with no lesion but both have a high % of cows without a lesion and the camera is not that much worse

```{r}
#| label: q2-trimonly-summary

q2_first_event_detail <- q2_any |>
  filter(event_occurred == 1) |>
  left_join(lame_data |> select(id_animal, lact_number, date_event, event, trimonly),
            by = c("id_animal", "lact_number", "date_event"))

q2_first_event_detail |>
  group_by(tx_group) |>
  summarize(n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100, .groups = "drop") |>
  bind_rows(
    q2_first_event_detail |> summarize(tx_group = "Overall", n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100)
  ) |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_trimonly, decimals = 1) |>
  cols_label(n = "N Cows with Trim", n_trimonly = "No Lesion", pct_trimonly = "% Trim-Only") |>
  tab_header(title = "43-55% of Trims do not have a lesion")
```

# How does the Farm staff do?

The `CHKLAME` event is used to flag cows in the need of trimming by farm staff.

The table below compares what % of cows with a NEDAP alert were flagged by farms staff \<=7 days of the NEDAP alert.

Overall farm staff flags about 20% of cows overall with more cows with low scores or strong declines being flagged.

```{r}
#| label: q3-build

chklame <- events_formatted |>
  filter(event == "CHKLAME") |>
  select(id_animal, lact_number, date_event) |>
  distinct()

# "Trim" means LAME + FOOTRIM + TRIM, the house rule throughout this report:
# when the trimmer finds a lesion at the visit, that visit is coded LAME and
# gets no separate FOOTRIM code, so a FOOTRIM/TRIM-only filter silently misses
# exactly the trims most likely to matter. (This chunk used FOOTRIM/TRIM alone
# until an audit caught it - the third time that same mistake appeared in this
# report. Impact here was one cow; the fix is for consistency, not magnitude.)
trims <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event) |>
  distinct()

# A CHKLAME on the SAME day as the NEDLAME alert counts as staff-identified
# too (folded into the "pre-week" bucket via <= rather than <), rather than
# falling through the gap between "before" (strict <) and "after" (strict >).
pre_week <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event >= first_nedlame_date - 7, date_event <= first_nedlame_date) |>
  distinct(id_animal, lact_number)

first_post_chklame <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  rename(chklame_date = date_event)

first_post_trim <- cohort |>
  inner_join(trims, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(id_animal, lact_number, trim_date = date_event)

post_flag <- first_post_chklame |>
  left_join(first_post_trim, by = c("id_animal", "lact_number")) |>
  filter(is.na(trim_date) | chklame_date < trim_date) |>
  distinct(id_animal, lact_number)

staff_id <- cohort |>
  mutate(
    staff_identified = as.integer(
      paste(id_animal, lact_number) %in% paste(pre_week$id_animal, pre_week$lact_number) |
      paste(id_animal, lact_number) %in% paste(post_flag$id_animal, post_flag$lact_number)
    )
  )
```

::: panel-tabset
## By Treatment Group

```{r}
#| label: q3-treatment-group-plot
#| fig-height: 4

staff_id |>
  group_by(tx_group) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = tx_group, y = pct_staff_id, fill = tx_group)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "% Staff-Identified (CHKLAME)")
```

## By Alert Type

Staff catches a bigger % of Low and Strong Decline alerts

```{r}
#| label: q3-alert-type-by-group-plot
#| fig-height: 4

staff_id |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = alert_type, y = pct_staff_id, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")),
            position = position_dodge(width = 0.7), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  labs(x = "Alert Type", y = "% Staff-Identified (CHKLAME)", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Milk Production After 1st NEDLAME Alert

```{r}
#| label: q1-build

milk_cohort <- cohort |>
  inner_join(
    milk_resolved |> select(id_animal, lact_number, milk_date, dmlk1),
    by = c("id_animal", "lact_number"),
    relationship = "many-to-many"
  ) |>
  mutate(day_rel = as.numeric(milk_date - first_nedlame_date)) |>
  filter(day_rel >= -params$milk_window_pre, day_rel <= params$milk_window_post)

baseline <- milk_cohort |>
  filter(day_rel >= -params$milk_baseline_days, day_rel <= -1) |>
  group_by(id_animal, lact_number) |>
  summarize(baseline_dmlk1 = mean(dmlk1, na.rm = TRUE), .groups = "drop")

milk_norm <- milk_cohort |>
  inner_join(baseline, by = c("id_animal", "lact_number")) |>
  filter(baseline_dmlk1 >= 5) |> # drop the rare near-zero baseline (unreliable % denominator)
  mutate(pct_of_baseline = dmlk1 / baseline_dmlk1 * 100)

n_milk_cows <- n_distinct(milk_norm$id_animal)

# The milk extract is pulled separately from the DairyComp events and lags
# behind them, so the number of cows contributing FALLS as day_rel grows: a
# cow alerted less than 30 days before the milk data ends simply has no day-30
# record. n_milk_cows counts anyone with a usable baseline and at least one
# post-alert record, which is a much larger group than the one behind the
# right-hand end of the curve - quote both or the reader assumes the whole
# cohort is behind the day-30 figure. The n >= 5 filter on the plots below
# shortens the line rather than making it noisy, so the drop-off is invisible
# unless it is stated.
milk_coverage_end <- max(milk_norm$milk_date, na.rm = TRUE)
n_milk_day30 <- n_distinct(milk_norm$id_animal[milk_norm$day_rel == params$milk_window_post])
n_no_milk_at_all <- sum(cohort$first_nedlame_date > milk_coverage_end)

trend <- milk_norm |>
  group_by(tx_group, day_rel) |>
  summarize(
    mean_pct = mean(pct_of_baseline, na.rm = TRUE),
    se_pct = sd(pct_of_baseline, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )
```

The figures below show the impact of the treatment group on milk production.

- `r n_milk_cows` of the `r nrow(cohort)` cohort cows had at least one milk record post trimming and a usable pre-alert baseline. Not all cows have made it to the 30 day mark so there is more variability at those time points.

Overall numerically there is an advantage to the trimmed group in milk production after an alert but there is a lot of variation (the shaded part aroudn the solid line) so this not something to put a lot of trust in.

**NOTE:** *Due to the fact that cows were at very different lactation stages when flagged milk production is not directly comparable across cows. Each cow's production is normalized to **% of her own pre-alert baseline** (mean production over the `r params$milk_baseline_days` days before the alert) before averaging by day and treatment group. This allow us to isolates the shape of the trend/recovery from each cow's absolute production level.*

::: panel-tabset
## By Treatment Group

```{r}
#| label: q1-plot
#| fig-height: 5

ggplot(trend, aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    x = "Days Relative to First NEDLAME Alert",
    y = "Milk Production (% of Own Pre-Alert Baseline)",
    color = "Treatment Group", fill = "Treatment Group"
  )
```

## By Alert Type

Does the milk trend differ by the type of the Nedap alert?

```{r}
#| label: q1-alert-type-plot
#| fig-height: 5

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Time to Culling After NEDAP Enrollment

To evaluate the impact of treatment group on culling a time to culling graph is show below.

No real impact is seen.

```{r}
#| label: q6-build

# `date_left` is already attached to `cohort` upstream (it's needed there to
# censor Q2 at culling), so this chunk must NOT re-join it - doing so yields
# date_left.x / date_left.y and the bare name disappears. date_sold/date_died
# were only ever selected here, never used.
q6 <- cohort |>
  mutate(
    culled_ever = as.integer(!is.na(date_left)),
    days_to_censor_cull = if_else(
      culled_ever == 1,
      as.numeric(date_left - first_nedlame_date),
      as.numeric(date_max_pull - first_nedlame_date)
    )
  ) |>
  filter(days_to_censor_cull >= 0)

km_cull <- survfit2(Surv(days_to_censor_cull, culled_ever) ~ tx_group, data = q6)
```

```{r}
#| label: q6-plot
#| fig-height: 5

km_cull |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% Culled") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q6-table

km_cull |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% Culled After First NEDLAME Alert")
```

# What cows were missed by the system?

This section looks at cows with lesion regardless of treatment group.

Among cows with a `LAME`-diagnosed lesion on or after `r params$history_after_date`, this checks whether that cow had a `NEDLAME` alert in the preceding lookback window (same lactation).

A "miss" means staff/routine trimming found a lesion that NEDAP hadn't already flagged.

**Note** *Lame events that were in heifers, dried off cows or cows trimmed within 28 days of another trim were excluded as the system does not monitor those cows or they were automatically excluded*.

```{r}
#| label: q4-build

# Confirmed directly with the farm: the DairyComp FTDAT gate tracks days
# since the most recent LAME diagnosis OR trim - not trims alone. A cow
# with frequent LAME diagnoses but no formal trim in between is just as
# ineligible for a new alert as a recently-trimmed one.
trims_for_suppression <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, trim_date = date_event) |>
  distinct()

lame_lesion_recent_all <- lame_data |>
  filter(event == "LAME", lesion == 1,
         date_event >= ymd(params$history_after_date),
         date_event <= analysis_end) |>
  left_join(animal_lactations |> select(id_animal, lact_number, date_fresh, date_dry), by = c("id_animal", "lact_number"))

# Days since this cow's most recent PRIOR trim, for each lesion case - computed
# so that a case with no prior trim at all still gets exactly one row (Inf),
# rather than a naive filter-then-min silently dropping it (see lessons-learned.md
# in mcp_server/knowledge/ for why that pattern is dangerous).
recent_trim_check <- lame_lesion_recent_all |>
  select(id_animal, lact_number, date_event) |>
  left_join(trims_for_suppression, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(days_since_trim = if_else(!is.na(trim_date) & trim_date < date_event, as.numeric(date_event - trim_date), Inf)) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(days_since_last_trim = min(days_since_trim), .groups = "drop")

lame_lesion_recent_all <- lame_lesion_recent_all |>
  left_join(recent_trim_check, by = c("id_animal", "lact_number", "date_event"))

# exclude lesions diagnosed on a heifer who never freshened (lact_number ==
# 0, no date_fresh on record - confirmed this is exactly how this herd's
# data represents "hasn't calved yet" - a real cowcard cross-check (cow
# 31377) caught this: she's lact_number 0 and was being counted as a
# "miss" despite never having been eligible for NEDAP monitoring at all),
# after the cow was already dried off, or diagnosed while NEDAP would
# have been structurally suppressed by a too-recent trim
n_excluded_prefresh <- sum(is.na(lame_lesion_recent_all$date_fresh))
n_excluded_dry <- sum(!is.na(lame_lesion_recent_all$date_fresh) & !is.na(lame_lesion_recent_all$date_dry) & lame_lesion_recent_all$date_event > lame_lesion_recent_all$date_dry)
n_excluded_recent_trim <- sum(
  !is.na(lame_lesion_recent_all$date_fresh) &
    lame_lesion_recent_all$days_since_last_trim <= params$recent_trim_suppression_days &
    (is.na(lame_lesion_recent_all$date_dry) | lame_lesion_recent_all$date_event <= lame_lesion_recent_all$date_dry)
)

# A cow who was never enrolled in SmartSight was never monitored for
# mobility at all, so a lesion found on her cannot be a camera miss - the
# same reasoning that already excludes dry cows and unfreshened heifers.
# Enrollment is judged renumber-aware via enrolled_ids (see the setup chunk).
lame_lesion_recent_all <- lame_lesion_recent_all |>
  left_join(events_all |> select(id_animal, id) |> distinct(), by = "id_animal") |>
  mutate(enrolled_smartsight = as.character(id) %in% enrolled_ids)

# Counted at the SAME grain as q4 (one row per cow-lactation, after the
# first-qualifying-case slice), so that nrow(q4) + this reconciles against the
# pre-exclusion total. Counting raw lesion rows instead double-counts a cow
# with two qualifying events.
excluded_not_enrolled_cases <- lame_lesion_recent_all |>
  filter(!is.na(date_fresh),
         is.na(date_dry) | date_event <= date_dry,
         days_since_last_trim > params$recent_trim_suppression_days,
         !enrolled_smartsight) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup()
n_excluded_not_enrolled <- nrow(excluded_not_enrolled_cases)

lame_lesion_recent <- lame_lesion_recent_all |>
  filter(!is.na(date_fresh)) |>
  filter(is.na(date_dry) | date_event <= date_dry) |>
  filter(days_since_last_trim > params$recent_trim_suppression_days) |>
  filter(enrolled_smartsight) |>
  select(-enrolled_smartsight, -id) |>
  select(-date_fresh, -date_dry, -days_since_last_trim) |>
  # first qualifying lesion case per cow-lactation only
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup()

nedlame_all <- events_all |>
  filter(Event == "NEDLAME") |>
  select(id_animal, lact_number, ned_date = date_event) |>
  distinct()

q4 <- lame_lesion_recent |>
  mutate(
    lookback_days_available = pmax(as.numeric(date_event - nedlame_start_date), 0),
    lookback_days_used = pmin(params$q4_lookback_days, lookback_days_available)
  ) |>
  left_join(nedlame_all, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(lookback_hit = !is.na(ned_date) & ned_date <= date_event & ned_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event, lookback_days_used) |>
  summarize(caught_by_nedap = any(lookback_hit), .groups = "drop") |>
  left_join(cohort |> select(id_animal, lact_number, tx_group), by = c("id_animal", "lact_number")) |>
  mutate(tx_group = if_else(is.na(tx_group), "Never Alerted (Not in Pilot Cohort)", tx_group))

n_truncated_lookback <- sum(q4$lookback_days_used < params$q4_lookback_days)

# The "Never Alerted" label is assigned by absence from `cohort`, and
# `cohort` now excludes the post-trim-artifact cows - so that bucket is NOT
# purely "never got an alert." Split it explicitly rather than describing it
# loosely: an earlier draft called the non-pure share "a small number" when
# it is actually ~18% of the bucket.
# Must match the figure's "Never Alerted" band exactly, which means carrying
# the !caught_by_nedap test too - a handful of cows are outside the cohort yet
# were still flagged in time, and they belong in "Caught in Time", not here.
# (detection_group itself is built later, so the rule is repeated rather than
# referenced; if one changes, change both.)
never_alerted_rows <- q4 |>
  filter(!caught_by_nedap, tx_group == "Never Alerted (Not in Pilot Cohort)")
n_never_alerted_bucket <- nrow(never_alerted_rows)
n_never_alerted_true <- never_alerted_rows |>
  anti_join(all_nedlame_cows |> select(id_animal, lact_number), by = c("id_animal", "lact_number")) |>
  nrow()
n_never_alerted_excluded <- n_never_alerted_bucket - n_never_alerted_true
```

- A total of `r nrow(q4)` cows had a lesion diagnosed after the project start date (`r params$history_after_date`).

  - Of those `r nrow(q4)` cows, `r sum(!q4$caught_by_nedap)` (`r round(100 * mean(!q4$caught_by_nedap), 1)`%) had no NEDAP alert entered in DairyComp in the `r params$q4_lookback_days` days before their lesion was diagnosed.

## What Lesions are being missed?

The figure below shows the lesion types found at these first-diagnosis cases, broken out by NEDAP detection status:

- **Never Alerted** — no usable NEDLAME alert. Most were never alerted at all (`r n_never_alerted_true` of `r n_never_alerted_bucket`); the other `r n_never_alerted_excluded` had an alert that was excluded because it came after trimming.
- **Alerted, Missed** — she was alerted at some point, but not in the lookback window before *this* lesion. Some of these came after her first lesion.
- **Alerted, Caught in Time** — an alert landed in the window before the lesion. NEDAP flagged her first.

From the figure below it is clear that DD and other are the main categories being missed. Other is typically Corkscrews. The missing of DD is not suprising as the camera was trained on locomotion scoring that also misses cows with DD.

```{r}
#| label: q4-lesion-types

lesion_type_cols <- c("dd", "footrot", "wld", "sole_ulcer", "injury", "cork",
                       "hemorrhage", "sole_fracture", "toe_ulcer", "thin", "other")

q4_group <- q4 |>
  # Order matters: test caught_by_nedap FIRST, then cohort membership.
  #
  # caught_by_nedap looks at ANY NEDLAME event, while cohort membership also
  # requires a valid MNFRS and survival of the post-trim exclusion. So a cow
  # can be flagged in time yet sit outside the cohort. Testing cohort
  # membership first labelled those cows "Never Alerted" even though an alert
  # did reach DairyComp inside the window - 4 cases, and it stopped the two
  # missed categories summing to the miss total (55 + 288 = 343, not 339).
  #
  # NB when editing: compare tx_group against the value it actually holds (set
  # in q4-build), not the short display label - the short label matches nothing
  # and silently dumps every never-alerted cow into "Alerted, Missed".
  mutate(detection_group = case_when(
    caught_by_nedap ~ "Alerted, Caught in Time",
    tx_group == "Never Alerted (Not in Pilot Cohort)" ~ "Never Alerted",
    TRUE ~ "Alerted, Missed"
  ))

# The two missed categories must account for every miss. If this ever fails,
# the labels have drifted away from caught_by_nedap again.
stopifnot(
  sum(q4_group$detection_group %in% c("Never Alerted", "Alerted, Missed")) ==
    sum(!q4_group$caught_by_nedap)
)

# and the prose counters above must describe the same band as the figure
stopifnot(sum(q4_group$detection_group == "Never Alerted") == n_never_alerted_bucket)

q4_lesion_types <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(q4_group |> select(id_animal, lact_number, date_event, detection_group),
             by = c("id_animal", "lact_number", "date_event")) |>
  pivot_longer(cols = all_of(lesion_type_cols), names_to = "lesion_type", values_to = "has_lesion") |>
  filter(has_lesion == 1) |>
  count(detection_group, lesion_type) |>
  # Shown as a share within each lesion type, not as counts. On a count axis a
  # common lesion looks like a detection problem simply because it is common -
  # the same base-rate trap the DD-by-foot table further down was built to
  # avoid. The case count is printed beside each bar so a lesion with only a
  # handful of cases can be discounted rather than read as a real rate.
  group_by(lesion_type) |>
  mutate(total_cases = sum(n), pct = 100 * n / total_cases) |>
  ungroup() |>
  mutate(lesion_type = fct_reorder(lesion_type, total_cases))

lesion_totals <- q4_lesion_types |> distinct(lesion_type, total_cases)

ggplot(q4_lesion_types, aes(x = lesion_type, y = pct, fill = detection_group)) +
  geom_col(width = 0.7) +
  geom_text(data = lesion_totals, inherit.aes = FALSE,
            aes(x = lesion_type, y = 101, label = paste0("n=", total_cases)),
            hjust = 0, size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 100, 25), labels = function(x) paste0(x, "%"),
                     limits = c(0, 118), expand = expansion(mult = c(0, 0))) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases with that lesion", fill = "NEDAP Detection Status") +
  theme(legend.position = "bottom")
```

## Of the Misses, Did Farm Staff Catch Them Anyway?

```{r}
#| label: q4-staff-catch-build

missed_cases <- q4 |> filter(!caught_by_nedap)

staff_catch <- missed_cases |>
  left_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(staff_flagged = !is.na(date_event.y) & date_event.y <= date_event.x & date_event.y >= date_event.x - 7) |>
  group_by(id_animal, lact_number, date_event = date_event.x, tx_group) |>
  summarize(staff_flagged = any(staff_flagged), .groups = "drop")
```

Of the `r nrow(staff_catch)` cows that were missed by the camera but had a lesion (trimmed due to routine trim or farm staff reasons),

- `r sum(staff_catch$staff_flagged)` (`r round(100 * mean(staff_catch$staff_flagged), 1)`%) were flagged by the farm staff in the 7 days before diagnosis.

# **The bottom line**

- NEDAP camera preforms better than CattleEye

- Cows with DD are being missed by the camera

- NEDAP camera catches more cows than just farm staff/routine trimming

- Catching these extra cows does not impact short term milk production and culling in the short term

- The study was not long enough to determine if more cows being caught by camera do better long term as in less repeat lesions.