---
title: "NEDLAME Alert Pilot: TX (Trim) vs Control"
editor: source
format:
 html:
   embed-resources: true
   toc: true
   toc-location: left
   code-fold: true
execute:
    echo: false
    message: false
    warning: false
params:
  milk_window_pre: 14
  milk_window_post: 30
  milk_baseline_days: 7
  q4_lookback_days: 21
  history_after_date: "2026-06-01"
  recent_trim_suppression_days: 28
  post_trim_exclusion_days: 7
  analysis_end_date: "2026-08-27"
---

```{r}
#| label: setup

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  arrow,
  lubridate,
  gt,
  gtsummary,
  survival,
  survminer,
  ggsurvfit,
  scales,
  viridis
)

source(file.path('functions/fxn_read_milk_folder.R'))
source(file.path('functions/fxn_read_nedap_attentions.R'))

# GC's lame functions (fxn_collapse_lesions, fxn_trim_vars, fxn_dz_status) -
# fetched live from GitHub, same mechanism report_explore_lame_new.qmd uses.
# Two local overrides are sourced AFTER fxn_load_os_fxns() specifically so
# they win over the GitHub versions in the global environment:
#  - functions/fxn_code_lesions.R: the GitHub version's `trimonly` logic
#    only checks the *protocols* field for "NONE" - this farm's
#    routine/no-lesion trims are coded "NONE" in the *remark* field
#    instead, which the GitHub version doesn't check, so it silently
#    misclassifies every routine trim as having a lesion.
#  - functions/fxn_collapse_lesions.R: the GitHub version reassigns its
#    own `data` variable to an already-one-row-per-date subset before
#    computing the max() union across that date's lesion columns, so
#    when a cow has multiple event rows on the same date coding
#    DIFFERENT lesions, only whichever row survives the earlier
#    reduction keeps its lesion flags - the rest are silently dropped
#    (confirmed against a real cow with 4 same-day LAME rows: only 1 of
#    her 3 distinct lesion types survived under the GitHub version).
source(file.path('functions/fxn_load_os_fxns.R'))
fxn_load_os_fxns()
source(file.path('functions/fxn_code_lesions.R'))
source(file.path('functions/fxn_collapse_lesions.R'))

events_all <- read_parquet('data/intermediate_files/events_all_columns.parquet')
events_formatted <- read_parquet('data/intermediate_files/events_formatted.parquet')
animals <- read_parquet('data/intermediate_files/animals.parquet')
animal_lactations <- read_parquet('data/intermediate_files/animal_lactations.parquet')

date_max_pull <- max(events_all$date_event, na.rm = TRUE)
farm_name <- unique(events_all$location_event)

# NEDLAME go-live date, per the DairyComp setup: enrollment protocols were
# only turned on starting ~June 1 2026, so no NEDLAME alert can exist
# before this regardless of how "missed" a case looks.
nedlame_start_date <- min(events_all$date_event[events_all$Event == "NEDLAME"], na.rm = TRUE)
```

```{r}
#| label: cohort

# One row per cow (id_animal) that ever got a NEDLAME alert - grouped by
# id_animal alone since a cow only has one "first" NEDLAME ever. Treatment
# group is read off the MNFRS value on the FIRST NEDLAME event for that cow
# - MNFRS is stable across a cow's NEDLAME rows even though it drifts over
# her broader event history (0 before enrollment, 1/2 once enrolled).
# Hard analysis cut-off. The raw Nedap sensor export ends on
# params$analysis_end_date, so alerts and lesion cases after it cannot be
# checked against what the camera actually saw. Rather than carry a partly
# checkable tail, entry into the analysis stops there for BOTH alerts (here)
# and lesion cases (Q4). Follow-up events are deliberately NOT capped - a
# cow who entered on/before the cut-off keeps every trim, cull and milk
# record recorded after it, which is what time-to-event needs.
analysis_end <- ymd(params$analysis_end_date)

first_nedlame_raw <- events_all |>
  filter(Event == "NEDLAME", date_event <= analysis_end) |>
  group_by(id_animal) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  transmute(
    id_animal, id,
    first_nedlame_date = date_event,
    alert_type = Remark,
    tx_group = case_when(
      MNFRS == "1" ~ "TX (Trim)",
      MNFRS == "2" ~ "Control",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(tx_group))

# lact_number is deliberately NOT taken from the raw NEDLAME event's own
# field - right at a freshening transition (dry -> fresh, same day or one
# day apart), DC's raw event can carry a stale lact_number from the
# lactation that just ended rather than the one that just started (found
# in 2 of 718 cows once the dataset grew past ~700 alerts: their alert
# date fell after date_next_fresh/on date_fresh of the FOLLOWING
# lactation, yet the raw event said the OLD lact_number). Resolving to
# whichever lactation was actually active as of the alert date - the same
# nearest-preceding-fresh-date pattern already used for the milk join -
# fixes this without needing to trust the raw field.
first_nedlame <- first_nedlame_raw |>
  inner_join(
    animal_lactations |>
      filter(!is.na(date_fresh)) |>
      select(id_animal, lact_number, date_fresh, date_dry, date_next_fresh),
    by = "id_animal",
    relationship = "many-to-many"
  ) |>
  filter(date_fresh <= first_nedlame_date) |>
  group_by(id_animal) |>
  slice_max(date_fresh, n = 1, with_ties = FALSE) |>
  ungroup()

# Flag cows whose FIRST NEDLAME fired 1-`r params$post_trim_exclusion_days`
# days AFTER a recorded trim, same cow-lactation - per the farm's
# confirmed data-load schedule (NEDLAME's daily batch loads at ~5am and
# stamps each alert with the day it was LOADED, not necessarily the day
# the underlying sensor attention happened - an attention any time after
# ~5am on day D gets logged with date D+1), an alert dated on or before
# the trim date is a genuine independent catch, but one dated 1+ days
# AFTER the trim is very likely the camera picking up that cow's own
# post-trim gait change, not a new problem. It got through DC's own
# FTDAT suppression gate only because the trim record itself is commonly
# entered a day or two late (backdated to the actual trim date) - it
# simply wasn't in the system yet when NEDLAME's batch ran. See the new
# "Excluded" section below for the full writeup and examples.
#
# "Trim" here means LAME, FOOTRIM, *or* TRIM - same definition already
# established for Q4's FTDAT suppression gate (Round 9): a LAME visit
# IS the trim whenever the trimmer finds and codes a lesion at that
# same encounter, it just doesn't also get a separate FOOTRIM code that
# day. An earlier version of this exact chunk only checked
# FOOTRIM/TRIM, which missed exactly this case - cow 7207's LAME (White
# Line) visit on 06/11 was her actual trim, and NEDLAME fired the very
# next day (confirmed directly against her real cowcard).
trims_all <- events_all |>
  filter(Event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, trim_date = date_event) |>
  distinct()

post_trim_check <- first_nedlame |>
  select(id_animal, lact_number, first_nedlame_date) |>
  inner_join(trims_all, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(trim_date < first_nedlame_date, trim_date >= first_nedlame_date - params$post_trim_exclusion_days) |>
  group_by(id_animal, lact_number) |>
  summarize(days_since_nearest_trim = min(as.numeric(first_nedlame_date - trim_date)), .groups = "drop")

first_nedlame <- first_nedlame |>
  left_join(post_trim_check, by = c("id_animal", "lact_number")) |>
  mutate(flagged_post_trim = !is.na(days_since_nearest_trim))

# Attach lactation-stage context (date_fresh/date_dry) so every downstream
# join can be scoped to id_animal + lact_number, never crossing into a
# different lactation for the same cow.
all_nedlame_cows <- first_nedlame |>
  left_join(animals |> select(id_animal, date_left), by = "id_animal") |>
  mutate(
    dim_at_nedlame = as.numeric(first_nedlame_date - date_fresh),
    date_lact_end = pmin(
      coalesce(date_dry, date_next_fresh, date_max_pull),
      date_max_pull,
      na.rm = TRUE
    ),
    # Observation actually ends when the cow leaves the herd, if that comes
    # first: a culled/dead cow can't have a subsequent trim, so counting her
    # as still at risk inflates the denominator and understates the event
    # rate. date_lact_end alone ignored this. pmax() floors it at her alert
    # date so a data error can't produce negative follow-up.
    date_obs_end = pmax(
      pmin(date_lact_end, coalesce(date_left, date_max_pull), na.rm = TRUE),
      first_nedlame_date
    )
  )

# `cohort` (used throughout Q1-Q3, Q5, Q6) excludes the flagged cows;
# `cohort_excluded_post_trim` keeps them, visibly, for the "Excluded"
# section - nothing is silently dropped. Q4 is deliberately NOT scoped
# to `cohort` (it pulls NEDLAME/lesion history from the whole herd
# directly), so this exclusion does not affect Q4's miss-rate numbers.
cohort <- all_nedlame_cows |> filter(!flagged_post_trim)
cohort_excluded_post_trim <- all_nedlame_cows |> filter(flagged_post_trim)

n_tx <- sum(cohort$tx_group == "TX (Trim)")
n_control <- sum(cohort$tx_group == "Control")
n_excluded_post_trim <- nrow(cohort_excluded_post_trim)
```

```{r}
#| label: lesion-history-dataset

# Same lesion-coding / chronicity pipeline as report_explore_lame_new.qmd,
# just kept at the lactation grain (lact_number carried through) so the
# "history at enrollment" lookup below never crosses lactations.
all_lesions <- c("dd", "footrot", 'wld', "sole_ulcer", "injury",
                  "cork", "other", "hemorrhage",
                  "sole_fracture", "toe_ulcer", "thin",
                  "inf", "noninf", "toe", "axial", "lesion")
diseases <- c("inf", "noninf", "lesion")

lame_data <- events_formatted |>
  select(id_animal, lact_number, date_event, event,
         remark_letters1, protocols_letters1, location_event, protocols) |>
  filter(event %in% c("LAME", "TRIM", "FOOTRIM")) |>
  fxn_code_lesions(protocol_var = protocols_letters1, remark_var = remark_letters1) |>
  filter(!is.na(trimonly)) |>
  mutate(
    noninf = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 1,
      .default = noninf),
    other = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 0,
      .default = other)
  ) |>
  fxn_collapse_lesions(lesions = all_lesions) |>
  fxn_trim_vars(location_event, id_animal, date_var = date_event, trimonly_var = trimonly) |>
  fxn_dz_status(disease_cols = diseases, event_filter = trimonly == 0)

# Each NEDLAME cow's lameness-history label as of (at/before) her first
# NEDLAME date, using her FULL lifetime record (all lactations - lame_data
# goes back to 2021), not scoped to her current lactation. Unlike every
# other join in this report, this one is intentionally NOT lactation-scoped
# per explicit user direction: chronicity is a lifetime concept, and the
# ~5 years of history available should be used rather than artificially
# resetting it at each freshening.
#
# IMPORTANT: candidate history rows are filtered to "at/before enrollment"
# BEFORE joining to cohort, not after (same reasoning as fxn_time_to_next()
# above) - filtering after a left_join would silently DROP a cow from
# `cohort` entirely instead of correctly labeling her. An earlier version
# of this exact chunk had that bug and was quietly dropping 49 of 434 cows
# (~11%) from the ENTIRE cohort - not just from Q5, but from every section,
# since `cohort` is the shared base table.
history_lookup <- lame_data |>
  select(id_animal, date_event, status_lesion) |>
  filter(!is.na(status_lesion))

valid_history <- history_lookup |>
  inner_join(
    cohort |> select(id_animal, lact_number, first_nedlame_date),
    by = "id_animal"
  ) |>
  filter(date_event <= first_nedlame_date) |>
  select(id_animal, lact_number, date_event, status_lesion)

# A cow with no LESION ever diagnosed before her NEDLAME alert has no
# status_lesion to inherit - fxn_dz_status() only assigns a status to
# trimonly == 0 rows (an actual lesion found), so a cow's routine
# trim-only visits (trimonly == 1) never produce a status_lesion, no
# matter how many she's had. This is NOT the same as "no LAME/TRIM/
# FOOTRIM event at all" - checked directly, most of these cows (206 of
# 353 in the current cohort) DO have prior trim/LAME events, just never
# one where a lesion was found; only a small remainder (~21 of 353)
# truly have zero foot-care events before their alert. Rather than
# invent a 4th bucket for either case, she's folded into "New" - the
# same label her eventual first-ever diagnosis would get, and the
# closest existing category to "no established chronic pattern."
cohort <- cohort |>
  left_join(valid_history, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  group_by(id_animal, lact_number) |>
  slice_max(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
  ungroup() |>
  mutate(history_group = if_else(is.na(status_lesion), "New", status_lesion)) |>
  select(-date_event, -status_lesion)

history_levels <- c("New", "Repeat", "Chronic")
cohort <- cohort |>
  mutate(history_group = factor(history_group, levels = history_levels))
```

```{r}
#| label: milk-data

# data/milk/ is not wired into any existing pipeline script. Raw milk rows
# only carry a raw `id` (not the composite id_animal), so they're resolved
# to a specific animal-lactation by nearest-preceding date_fresh (see
# fxn_read_milk_folder.R for why a naive date_dry window isn't safe here).
milk_long <- fxn_read_milk_folder()
milk_resolved <- fxn_resolve_milk_to_lactation(milk_long, animal_lactations)
```

```{r}
#| label: nedap-attentions-data

# Raw Nedap sensor export - what the SENSOR actually flagged, independent
# of whether that flag ever reached DairyComp as a NEDLAME event. Used in
# Q4 to split "NEDAP missed it" cases into true sensor misses vs pipeline
# losses (a real cowcard/Pulse cross-check first surfaced this pattern -
# see the plan doc's Round 8/10). This is a one-time export, not a live
# feed, so its coverage window is whatever was current when it was
# pulled - captured here explicitly rather than assumed, since it can
# lag behind the DairyComp/milk data read in above.
attentions_raw <- fxn_read_nedap_attentions()
attentions_resolved <- fxn_resolve_nedap_to_lactation(attentions_raw, animal_lactations)

nedap_coverage_start <- min(attentions_resolved$attention_date, na.rm = TRUE)
nedap_coverage_end <- max(attentions_resolved$attention_date, na.rm = TRUE)
```

```{r}
#| label: platform-enrollment-data

# Nedap runs two separate platform instances at this farm - "Activity"
# (used by farm staff daily) and "SmartSight" (mobility monitoring,
# generates NEDLAME) - kept separate at the farm's request to avoid
# confusion when checking the Nedap platform. A cow absent from
# SmartSight's animal list was never being monitored for lameness at
# all, regardless of what the raw attentions export's date coverage
# shows - this check is animal-level enrollment, not date-bound, so it
# can resolve cases the date-bound attentions export above can't (see
# the Platform Enrollment Check subsection in Q4).
cows_activity <- read_csv("data/cows_activity.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

cows_smartsight <- read_csv("data/cows_smartsight.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

# Animal numbers get reused and reassigned. DairyComp records an ID change as
# an XID event whose Remark holds the OLD number, and 150 of those old numbers
# are still present in this extract as animals in their own right. A cow
# renumbered after she was enrolled therefore looks ABSENT from SmartSight
# under her current number while being present under her old one - confirmed
# on cow 23143, renumbered from 14347 on 2026-08-06 and enrolled all along as
# 14347. Checking the current number alone would call her unmonitored.
id_aliases <- events_all |>
  filter(Event == "XID") |>
  transmute(id = as.character(id), alias = trimws(as.character(Remark))) |>
  filter(!is.na(alias), grepl("^[0-9]+$", alias)) |>
  distinct()

smartsight_ids <- unique(cows_smartsight$id)
# a cow counts as enrolled if her current OR any prior number is on the list
enrolled_ids <- union(
  smartsight_ids,
  id_aliases$id[id_aliases$alias %in% smartsight_ids]
)
n_rescued_by_alias <- length(setdiff(
  id_aliases$id[id_aliases$alias %in% smartsight_ids], smartsight_ids))
```

# Overview

## Setup in DairyComp

For this study the follow things were created in DairyComp to handle the file data from NEDAP daily.

- A command to clear all filein items before new data was brought in as a workaround to deal with historical flag issues

- New NEDAP lame event with 3 protocols to flag cows and criteria to create them

  - Low (score \<30 today, no alarm in the last 90 days, not trimmed in last 90 days, not a cow with history of upper leg lameness)

    - ENTER EC=70\\P215 FOR LFTYP=30102 LMFLG=1 LMSV=1-30 DSNLM=90-1 FTDAT\<-90 UPLAM\<\>1

  - Low Declining (no alarm in the last 7 days, so a cow already alerted is not re-flagged, and not trimmed in last 28 days)

    - ENTER EC=70\\P214 FOR LFTYP=30101 LMFLG=1 DSNLM=7-1 FTDAT\<-28

  - Strong Decline (no alarm in the last 7 days, not trimmed in last 28 days); the second OR statement lets a cow who already has a Low alert still raise a Strong Decline

    - ENTER EC=70\\P213 FOR LFTYP=30100 LMFLG=1 (DSNLM=7-1 FTDAT\<-28)(FTDAT\<-28 %73.NDLAME.101\>LOW)

- Cows were added to trim list with a command that uses an existing item (MNFRS) that randomizes cows by lactation group.

  - Enter FTDAT=1.1.1 for MNFRS=1 %.70NEDLAME.101=TODAY

Above commands were all run daily around 5 am after NEDAP file in occurred.

## Results

- `r nrow(first_nedlame)` cows received a `NEDLAME` alert between `r format(min(first_nedlame$first_nedlame_date))` and `r format(max(first_nedlame$first_nedlame_date))`

- The analysis stops at `r format(analysis_end)`, the last day the raw Nedap sensor export covers. DairyComp holds alerts past that date, but they cannot be checked against what the camera actually saw, so they are left out of every section rather than being partly checkable. Follow-up is not capped: a cow who entered on or before the cut-off keeps her later trims, culls and milk records.

- `r n_excluded_post_trim` were excluded as alerts occurred after a recent hoof trimming. These were still coded in DairyComp as the farm imports hoof trimming data a few days after the actual hoof trimming date. See @tbl-post-trim-exclusion-table.

This results in

- `r n_tx` cows enrolled in the **TX (Trim)** arm (`MNFRS = 1`, automatically trimmed after alert)

- `r n_control` in the **Control** arm (`MNFRS = 2`, staff decide whether/when to act),

**NOTE**: no cow is counted more than once in any table below. Where this report looks for "the" trim, LAME diagnosis, staff check, or NEDLAME alert for a cow, it either uses her **first** qualifying occurrence in the relevant window or asks only whether *at least one* occurred. Therefore a cow flagged or diagnosed repeatedly isn't counted more than once. When a single trimming date has more than one distinct lesion type coded (e.g. white line disease *and* digital dermatitis found at the same trimming), each distinct type is kept; the same lesion type coded more than once on one day only counts once.

### Description of Cows

The tables below describe the cows enrolled by treatment group

## By Alert Type

::: panel-tabset
```{r}
#| label: overview-table

cohort |>
  count(tx_group, alert_type) |>
  pivot_wider(names_from = alert_type, values_from = n, values_fill = 0) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Most cows were flagged as Low Declines") |>
  tab_stubhead(label = "Treatment Group")
```

## By Lactation Number

```{r}
#| label: lact-group-table

cohort |>
  mutate(lact_group_display = case_when(
    lact_number == 1 ~ "1",
    lact_number == 2 ~ "2",
    lact_number >= 3 ~ "3+",
    TRUE ~ NA_character_
  )) |>
  count(tx_group, lact_group_display) |>
  pivot_wider(names_from = lact_group_display, values_from = n, values_fill = 0) |>
  select(tx_group, `1`, `2`, `3+`) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Most cows enrolled were Lact 3+",
    ) |>
  tab_stubhead(label = "Treatment Group") |>
  cols_label(`3+` = "3+")
```

## By Lactation Stage (DIM)

```{r}
#| label: dim-balance-table

cohort |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_dim = round(mean(dim_at_nedlame, na.rm = TRUE), 0),
    median_dim = round(median(dim_at_nedlame, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Some skewness in DIM",
    ) |>
  cols_label(n = "N Cows", mean_dim = "Mean DIM", median_dim = "Median DIM")
```

```{r}
#| label: q1-dim-check
#| fig-height: 4

cohort |>
  ggplot(aes(x = tx_group, y = dim_at_nedlame, fill = tx_group)) +
  geom_violin(alpha = 0.5, color = NA, trim = FALSE) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.4, size = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "DIM at 1st NEDLAME Alert")
```
:::

### Details about excluded cows with alerts right 1-7 days after trimming

```{r}
#| label: tbl-post-trim-exclusion-table

cohort_excluded_post_trim |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_days_after_trim = round(mean(days_since_nearest_trim), 1),
    .groups = "drop"
  ) |>
  bind_rows(
    cohort_excluded_post_trim |> summarize(tx_group = "Overall", n = n(), mean_days_after_trim = round(mean(days_since_nearest_trim), 1))
  ) |>
  gt(rowname_col = "tx_group") |>
  cols_label(n = "N Excluded", mean_days_after_trim = "Mean Days After Trim") |>
  tab_header(title = "Systems flags cows recenlty trimmed quickly")
```

```{r}
#| label: post-trim-exclusion-plot
#| fig-height: 4

cohort_excluded_post_trim |>
  count(alert_type, days_since_nearest_trim) |>
  ggplot(aes(x = factor(days_since_nearest_trim), y = n, fill = alert_type)) +
  geom_col() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = "Days Between Trim and NEDLAME Alert", y = "N Cows Excluded", fill = "Alert Type") +
  theme(legend.position = "bottom")
```

```{r}
#| label: post-trim-followup-check

# The real evidence for the artifact reading: did anything ever come of
# these alerts? Computed rather than asserted - an earlier draft described
# this as "a separate check (not shown here)" and instead leaned on the
# share of decline-type alerts, which turns out to be just the cohort's
# base rate (see prose below) and therefore argues nothing either way.
excluded_followup <- cohort_excluded_post_trim |>
  select(id_animal, lact_number, first_nedlame_date, date_lact_end) |>
  left_join(
    events_formatted |>
      filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
      select(id_animal, lact_number, followup_date = date_event) |>
      distinct(),
    by = c("id_animal", "lact_number"), relationship = "many-to-many"
  ) |>
  mutate(is_followup = !is.na(followup_date) &
           followup_date > first_nedlame_date &
           followup_date <= date_lact_end) |>
  group_by(id_animal, lact_number) |>
  summarize(had_followup = any(is_followup), .groups = "drop")

n_excluded_no_followup <- sum(!excluded_followup$had_followup)
pct_excluded_no_followup <- round(100 * n_excluded_no_followup / nrow(excluded_followup), 0)

# Base-rate comparison, so the decline-share figure can be reported honestly.
pct_decline_excluded <- round(100 * mean(cohort_excluded_post_trim$alert_type %in% c("LOW DECLINE", "STRONG DECLINE")), 1)
pct_decline_cohort <- round(100 * mean(cohort$alert_type %in% c("LOW DECLINE", "STRONG DECLINE")), 1)
```

# Milk Production After 1st NEDLAME Alert

```{r}
#| label: q1-build

milk_cohort <- cohort |>
  inner_join(
    milk_resolved |> select(id_animal, lact_number, milk_date, dmlk1),
    by = c("id_animal", "lact_number"),
    relationship = "many-to-many"
  ) |>
  mutate(day_rel = as.numeric(milk_date - first_nedlame_date)) |>
  filter(day_rel >= -params$milk_window_pre, day_rel <= params$milk_window_post)

baseline <- milk_cohort |>
  filter(day_rel >= -params$milk_baseline_days, day_rel <= -1) |>
  group_by(id_animal, lact_number) |>
  summarize(baseline_dmlk1 = mean(dmlk1, na.rm = TRUE), .groups = "drop")

milk_norm <- milk_cohort |>
  inner_join(baseline, by = c("id_animal", "lact_number")) |>
  filter(baseline_dmlk1 >= 5) |> # drop the rare near-zero baseline (unreliable % denominator)
  mutate(pct_of_baseline = dmlk1 / baseline_dmlk1 * 100)

n_milk_cows <- n_distinct(milk_norm$id_animal)

# The milk extract is pulled separately from the DairyComp events and lags
# behind them, so the number of cows contributing FALLS as day_rel grows: a
# cow alerted less than 30 days before the milk data ends simply has no day-30
# record. n_milk_cows counts anyone with a usable baseline and at least one
# post-alert record, which is a much larger group than the one behind the
# right-hand end of the curve - quote both or the reader assumes the whole
# cohort is behind the day-30 figure. The n >= 5 filter on the plots below
# shortens the line rather than making it noisy, so the drop-off is invisible
# unless it is stated.
milk_coverage_end <- max(milk_norm$milk_date, na.rm = TRUE)
n_milk_day30 <- n_distinct(milk_norm$id_animal[milk_norm$day_rel == params$milk_window_post])
n_no_milk_at_all <- sum(cohort$first_nedlame_date > milk_coverage_end)

trend <- milk_norm |>
  group_by(tx_group, day_rel) |>
  summarize(
    mean_pct = mean(pct_of_baseline, na.rm = TRUE),
    se_pct = sd(pct_of_baseline, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )
```

The figures below show the impact of the treatment group on milk production.

- `r n_milk_cows` of the `r nrow(cohort)` cohort cows had at least one milk record post trimming and a usable pre-alert baseline.

- The milk extract runs to `r format(milk_coverage_end)`, behind the event data. A cow needs `r params$milk_window_post` days of records after her alert to appear at the right-hand end of the curves, so the number of cows contributing falls across the x-axis: **`r n_milk_day30` cows sit behind the day-`r params$milk_window_post` point, not `r n_milk_cows`.** `r n_no_milk_at_all` cohort cows were alerted after the milk data ends and contribute nothing. The day-`r params$milk_window_post` figures therefore describe the earlier-alerting cows, and will change as milk data catches up.

Due to the fact that cows were at very different lactation stages when flagged milk production is not directly comparable across cows. Each cow's production is normalized to **% of her own pre-alert baseline** (mean production over the `r params$milk_baseline_days` days before the alert) before averaging by day and treatment group. This allow us to isolates the *shape* of the trend/recovery from each cow's absolute production level.

::: panel-tabset
## By Treatment Group

```{r}
#| label: q1-plot
#| fig-height: 5

ggplot(trend, aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    x = "Days Relative to First NEDLAME Alert",
    y = "Milk Production (% of Own Pre-Alert Baseline)",
    color = "Treatment Group", fill = "Treatment Group"
  )
```

## By Alert Type

Does the milk trend differ by the type of the Nedap alert?

```{r}
#| label: q1-alert-type-plot
#| fig-height: 5

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Trimming

## Time to next trimming after alert

To determine the accuracy of the alert for each cow, we looked for the first trimming event after her first NEDLAME alert. Cows were censored at the end of their lactation or culling.

The figures below show the time to the next trimming event split by treatment group and for various lesion or alerts types.

- It is clear that most of the trim cows got trimmed within 7-10 days.

```{r}
#| label: q2-build

next_events <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event, event) |>
  distinct()

lesion_lookup <- lame_data |> select(id_animal, lact_number, date_event, event, lesion)

# IMPORTANT: candidate events are filtered to the valid post-NEDLAME window
# BEFORE joining to cohort, not after. Filtering after a left_join would
# silently DROP a cow entirely (rather than correctly censoring her)
# whenever she has trim/LAME events somewhere in her history but none
# inside the valid window - an earlier version of this code had exactly
# that bug and was quietly excluding ~45% of cows from the "any event"
# curve (and a different ~25% from the "with lesion" curve), which is
# also why the two curves looked identical - they weren't being computed
# over the same denominator.
fxn_time_to_next <- function(candidate_events) {
  valid_candidates <- candidate_events |>
    inner_join(
      cohort |> select(id_animal, lact_number, first_nedlame_date, date_obs_end),
      by = c("id_animal", "lact_number")
    ) |>
    filter(date_event > first_nedlame_date, date_event <= date_obs_end) |>
    select(id_animal, lact_number, date_event)

  cohort |>
    left_join(valid_candidates, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
    group_by(id_animal, lact_number) |>
    slice_min(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
    ungroup() |>
    mutate(
      event_occurred = as.integer(!is.na(date_event)),
      days_to_event = if_else(
        event_occurred == 1,
        as.numeric(date_event - first_nedlame_date),
        as.numeric(date_obs_end - first_nedlame_date)
      )
    )
}

q2_any <- fxn_time_to_next(next_events)

next_events_lesion <- next_events |>
  left_join(lesion_lookup, by = c("id_animal", "lact_number", "date_event", "event")) |>
  filter(!is.na(lesion) & lesion == 1)

q2_lesion <- fxn_time_to_next(next_events_lesion)
```

::: panel-tabset
## Time to Any Trim/LAME Event

```{r}
#| label: q2-any-plot
#| fig-height: 5

km_any <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_any)

km_any |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-any-table

km_any |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Trim/LAME Event")
```

## Time to a Trim/LAME Event *With a Lesion*

```{r}
#| label: q2-lesion-plot
#| fig-height: 5

km_lesion <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_lesion)

km_lesion |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Lesion Found") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-lesion-table

km_lesion |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Lesion Found")
```

## By Alert Type

Decline cows mostly all get trimmed

```{r}
#| label: q2-alert-type-plot
#| fig-height: 6

alert_levels <- c("LOW", "LOW DECLINE", "STRONG DECLINE")

# ggsurvfit only exposes a single combined `strata` column, so a KM plot
# faceted by alert_type is built manually: fit one survfit2() per
# alert level (skipping any level without both treatment arms
# represented) and stack the tidy() estimates before plotting.
fittable_alert_levels <- alert_levels |>
  keep(\(a) {
    d <- q2_any |> filter(alert_type == a)
    n_distinct(d$tx_group) == 2 && nrow(d) >= 10
  })

q2_alert_type_km_data <- fittable_alert_levels |>
  map(\(a) {
    d <- q2_any |> filter(alert_type == a)
    fit <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = d)
    broom::tidy(fit) |>
      mutate(
        tx_group = str_remove(strata, "^tx_group="),
        alert_type = a,
        risk = 1 - estimate,
        # survival's CI bounds flip when converted to risk (1 - survival):
        # the upper bound on survival becomes the LOWER bound on risk.
        risk_low = 1 - conf.high,
        risk_high = 1 - conf.low
      )
  }) |>
  list_rbind() |>
  mutate(alert_type = factor(alert_type, levels = alert_levels))

skipped_alert_levels <- setdiff(alert_levels, fittable_alert_levels)

ggplot(q2_alert_type_km_data, aes(x = time, y = risk, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = risk_low, ymax = risk_high), alpha = 0.15, color = NA) +
  geom_step(linewidth = 1) +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event", color = "Treatment Group", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```

`r if (length(skipped_alert_levels) > 0) paste0("*Alert type(s) omitted for insufficient data in one or both arms: ", paste(skipped_alert_levels, collapse = ", "), ".*")`
:::

## What actually brought the Control cows in?

In the Control arm a cow is only trimmed if someone decides to act, so the trims above are a mixture of two different things: staff spotting her and putting her on the list, or her simply coming up on the routine schedule. A `CHKLAME` recorded between her alert and her trim marks the first.

```{r}
#| label: q2-chklame-build

# Same pre-filter-before-join pattern used throughout: restrict the CHKLAME
# candidates to the window FIRST, then join, so a cow with no qualifying
# CHKLAME is correctly labelled "routine" rather than dropped from the table.
chklame_events <- events_formatted |>
  filter(event == "CHKLAME") |>
  select(id_animal, lact_number, chk_date = date_event) |>
  distinct()

first_post_alert_trim <- next_events |>
  inner_join(cohort |> select(id_animal, lact_number, first_nedlame_date, date_obs_end, tx_group),
             by = c("id_animal", "lact_number")) |>
  filter(date_event > first_nedlame_date, date_event <= date_obs_end) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  rename(trim_date = date_event)

chk_before_trim <- first_post_alert_trim |>
  select(id_animal, lact_number, first_nedlame_date, trim_date) |>
  inner_join(chklame_events, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(chk_date >= first_nedlame_date, chk_date <= trim_date) |>
  distinct(id_animal, lact_number) |>
  mutate(staff_flagged = TRUE)

trim_driver <- first_post_alert_trim |>
  left_join(chk_before_trim, by = c("id_animal", "lact_number")) |>
  left_join(lame_data |> select(id_animal, lact_number, date_event, trimonly) |> distinct(),
            by = c("id_animal", "lact_number", "trim_date" = "date_event")) |>
  mutate(staff_flagged = coalesce(staff_flagged, FALSE),
         Driver = if_else(staff_flagged, "Staff flagged her (CHKLAME)", "Routine trim, no CHKLAME"),
         days_to_trim = as.numeric(trim_date - first_nedlame_date),
         found_lesion = coalesce(trimonly, 1) == 0)

control_trims <- trim_driver |> filter(tx_group == "Control")
n_ctl_trimmed <- nrow(control_trims)
n_ctl_staff   <- sum(control_trims$staff_flagged)
n_ctl_routine <- n_ctl_trimmed - n_ctl_staff
pct_ctl_staff <- round(100 * n_ctl_staff / n_ctl_trimmed, 1)
pct_lesion_staff   <- round(100 * mean(control_trims$found_lesion[control_trims$staff_flagged]), 0)
pct_lesion_routine <- round(100 * mean(control_trims$found_lesion[!control_trims$staff_flagged]), 0)
med_days_staff   <- median(control_trims$days_to_trim[control_trims$staff_flagged])
med_days_routine <- median(control_trims$days_to_trim[!control_trims$staff_flagged])
```

```{r}
#| label: tbl-trim-driver

control_trims |>
  group_by(Driver) |>
  summarize(`Cows` = n(),
            `% of trims` = round(100 * n() / n_ctl_trimmed, 1),
            `% finding a lesion` = round(100 * mean(found_lesion), 0),
            `Median days alert to trim` = median(days_to_trim),
            .groups = "drop") |>
  gt() |>
  tab_header(title = "Most Control trims were routine, and most of those found nothing") |>
  tab_source_note("Control cows only, counting each cow once at her first trim after the alert.")
```

- Of the `r n_ctl_trimmed` Control cows trimmed after their alert, **`r n_ctl_staff` (`r pct_ctl_staff`%) had a `CHKLAME` first**; the other `r n_ctl_routine` were trimmed with no staff check recorded.

- **The two groups are not doing the same job.** Staff-flagged trims found a lesion `r pct_lesion_staff`% of the time against `r pct_lesion_routine`% for routine ones, and happened at a median `r med_days_staff` days after the alert rather than `r med_days_routine`.

- So the Control curve above is mostly the routine trim schedule catching up with these cows, not staff responding to them. That is worth holding onto when comparing the arms: TX is being compared against a mixture, and only about a third of it is an active decision.

*This is a description of the trims that happened, not a third survival curve. Whether a cow gets a `CHKLAME` is decided after her alert, during follow-up, so splitting the earlier curves on it would compare groups defined by something that happens after the clock starts and would overstate the difference.*

## How many trims with no lesions?

The table below shows the number of cows who had a first NEDAP alert and where trimmed **but** did not have a lesions.

**This table's population is only the number of cows in the `r nrow(cohort)`-cow NEDLAME cohort that were trimmed.**

```{r}
#| label: q2-trimonly-summary

q2_first_event_detail <- q2_any |>
  filter(event_occurred == 1) |>
  left_join(lame_data |> select(id_animal, lact_number, date_event, event, trimonly),
            by = c("id_animal", "lact_number", "date_event"))

q2_first_event_detail |>
  group_by(tx_group) |>
  summarize(n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100, .groups = "drop") |>
  bind_rows(
    q2_first_event_detail |> summarize(tx_group = "Overall", n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100)
  ) |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_trimonly, decimals = 1) |>
  cols_label(n = "N Cows with Trim", n_trimonly = "No Lesion", pct_trimonly = "% Trim-Only") |>
  tab_header(title = "43-55% of Trims do not have a lesion")
```

# How does the Farm staff do?

The dairy uses a `CHKLAME` event to flag cows in the need of trimming. The tables below compares what % of cows with a NEDAP alert were flagged by farms staff \<=7 days of the NEDAP alert.

```{r}
#| label: q3-build

chklame <- events_formatted |>
  filter(event == "CHKLAME") |>
  select(id_animal, lact_number, date_event) |>
  distinct()

# "Trim" means LAME + FOOTRIM + TRIM, the house rule throughout this report:
# when the trimmer finds a lesion at the visit, that visit is coded LAME and
# gets no separate FOOTRIM code, so a FOOTRIM/TRIM-only filter silently misses
# exactly the trims most likely to matter. (This chunk used FOOTRIM/TRIM alone
# until an audit caught it - the third time that same mistake appeared in this
# report. Impact here was one cow; the fix is for consistency, not magnitude.)
trims <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event) |>
  distinct()

# A CHKLAME on the SAME day as the NEDLAME alert counts as staff-identified
# too (folded into the "pre-week" bucket via <= rather than <), rather than
# falling through the gap between "before" (strict <) and "after" (strict >).
pre_week <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event >= first_nedlame_date - 7, date_event <= first_nedlame_date) |>
  distinct(id_animal, lact_number)

first_post_chklame <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  rename(chklame_date = date_event)

first_post_trim <- cohort |>
  inner_join(trims, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(id_animal, lact_number, trim_date = date_event)

post_flag <- first_post_chklame |>
  left_join(first_post_trim, by = c("id_animal", "lact_number")) |>
  filter(is.na(trim_date) | chklame_date < trim_date) |>
  distinct(id_animal, lact_number)

staff_id <- cohort |>
  mutate(
    staff_identified = as.integer(
      paste(id_animal, lact_number) %in% paste(pre_week$id_animal, pre_week$lact_number) |
      paste(id_animal, lact_number) %in% paste(post_flag$id_animal, post_flag$lact_number)
    )
  )
```

::: panel-tabset
## By Treatment Group

```{r}
#| label: q3-treatment-group-plot
#| fig-height: 4

staff_id |>
  group_by(tx_group) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = tx_group, y = pct_staff_id, fill = tx_group)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "% Staff-Identified (CHKLAME)")
```

## By Alert Type

Staff catches a bigger % of Low and Strong Decline alerts

```{r}
#| label: q3-alert-type-by-group-plot
#| fig-height: 4

staff_id |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = alert_type, y = pct_staff_id, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")),
            position = position_dodge(width = 0.7), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  labs(x = "Alert Type", y = "% Staff-Identified (CHKLAME)", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Influence of Lameness History

Lameness history can play an important role in camera performance. @fig-chronic shows the breakdown of new vs chronic cows that were enrolled. @fig-chronic-milk shows the impact on milk production between the lesion history types.

Each cow's lameness-history label (`New` / `Chronic`) reflects her **lifetime record** as of her first NEDLAME alert. The `lame_data` goes back to 2021, and this deliberately isn't restricted to her current lactation, since chronicity for lameness should be considered a lifetime event.

```{r}
#| label: fig-chronic
#| fig-height: 4

cohort |>
  count(tx_group, history_group) |>
  group_by(tx_group) |>
  mutate(pct = n / sum(n) * 100) |>
  ungroup() |>
  mutate(history_group = fct_relabel(history_group, \(x) str_wrap(x, width = 14))) |>
  ggplot(aes(x = history_group, y = n, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(n, " (", round(pct, 0), "%)")),
            position = position_dodge(width = 0.7), vjust = -0.3, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal() +
  labs(x = NULL, y = "N Cows", fill = "Treatment Group",
       title = "Lameness-History Distribution at Enrollment") +
  theme(legend.position = "bottom",
        axis.text.x = element_text(lineheight = 0.9),
        plot.margin = margin(10, 15, 10, 15))
```

::: callout-note
`Repeat` is left out of the plot below but kept in the chart above. All `r sum(cohort$history_group == "Repeat")` `Repeat` cows got that label the same way: their most recent lesion before enrollment landed exactly one day after an earlier one. That is a timing artifact, not a clinical category.
:::

## Milk Trend by History Group

```{r}
#| label: fig-chronic-milk
#| fig-height: 6

milk_norm |>
  filter(history_group %in% c("New", "Chronic")) |>
  group_by(tx_group, history_group, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~history_group) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

# Time to Culling After NEDAP Enrollment

To evaluate the impact of treatment group on culling a time to culling graph is show below. No real impact is seen.

```{r}
#| label: q6-build

# `date_left` is already attached to `cohort` upstream (it's needed there to
# censor Q2 at culling), so this chunk must NOT re-join it - doing so yields
# date_left.x / date_left.y and the bare name disappears. date_sold/date_died
# were only ever selected here, never used.
q6 <- cohort |>
  mutate(
    culled_ever = as.integer(!is.na(date_left)),
    days_to_censor_cull = if_else(
      culled_ever == 1,
      as.numeric(date_left - first_nedlame_date),
      as.numeric(date_max_pull - first_nedlame_date)
    )
  ) |>
  filter(days_to_censor_cull >= 0)

km_cull <- survfit2(Surv(days_to_censor_cull, culled_ever) ~ tx_group, data = q6)
```

```{r}
#| label: q6-plot
#| fig-height: 5

km_cull |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% Culled") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q6-table

km_cull |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% Culled After First NEDLAME Alert")
```

# What cows were missed by the system?

This section looks at cows with lesion regardless of treatment group. Among cows with a `LAME`-diagnosed lesion on or after `r params$history_after_date`, this checks whether that cow had a `NEDLAME` alert in the preceding lookback window (same lactation). A "miss" means staff/routine trimming found a lesion that NEDAP hadn't already flagged. Each cow is counted once here, at her **first** qualifying lesion diagnosis on/after that date (consistent with using the first event throughout this report).

**Note** *Lame events that were in heifers, dried off cows or cows trimmed within 28 days of another trim were excluded as the system does not monitor those cows or they were automatically excluded*.

```{r}
#| label: q4-build

# Confirmed directly with the farm: the DairyComp FTDAT gate tracks days
# since the most recent LAME diagnosis OR trim - not trims alone. A cow
# with frequent LAME diagnoses but no formal trim in between is just as
# ineligible for a new alert as a recently-trimmed one.
trims_for_suppression <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, trim_date = date_event) |>
  distinct()

lame_lesion_recent_all <- lame_data |>
  filter(event == "LAME", lesion == 1,
         date_event >= ymd(params$history_after_date),
         date_event <= analysis_end) |>
  left_join(animal_lactations |> select(id_animal, lact_number, date_fresh, date_dry), by = c("id_animal", "lact_number"))

# Days since this cow's most recent PRIOR trim, for each lesion case - computed
# so that a case with no prior trim at all still gets exactly one row (Inf),
# rather than a naive filter-then-min silently dropping it (see lessons-learned.md
# in mcp_server/knowledge/ for why that pattern is dangerous).
recent_trim_check <- lame_lesion_recent_all |>
  select(id_animal, lact_number, date_event) |>
  left_join(trims_for_suppression, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(days_since_trim = if_else(!is.na(trim_date) & trim_date < date_event, as.numeric(date_event - trim_date), Inf)) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(days_since_last_trim = min(days_since_trim), .groups = "drop")

lame_lesion_recent_all <- lame_lesion_recent_all |>
  left_join(recent_trim_check, by = c("id_animal", "lact_number", "date_event"))

# exclude lesions diagnosed on a heifer who never freshened (lact_number ==
# 0, no date_fresh on record - confirmed this is exactly how this herd's
# data represents "hasn't calved yet" - a real cowcard cross-check (cow
# 31377) caught this: she's lact_number 0 and was being counted as a
# "miss" despite never having been eligible for NEDAP monitoring at all),
# after the cow was already dried off, or diagnosed while NEDAP would
# have been structurally suppressed by a too-recent trim
n_excluded_prefresh <- sum(is.na(lame_lesion_recent_all$date_fresh))
n_excluded_dry <- sum(!is.na(lame_lesion_recent_all$date_fresh) & !is.na(lame_lesion_recent_all$date_dry) & lame_lesion_recent_all$date_event > lame_lesion_recent_all$date_dry)
n_excluded_recent_trim <- sum(
  !is.na(lame_lesion_recent_all$date_fresh) &
    lame_lesion_recent_all$days_since_last_trim <= params$recent_trim_suppression_days &
    (is.na(lame_lesion_recent_all$date_dry) | lame_lesion_recent_all$date_event <= lame_lesion_recent_all$date_dry)
)

# A cow who was never enrolled in SmartSight was never monitored for
# mobility at all, so a lesion found on her cannot be a camera miss - the
# same reasoning that already excludes dry cows and unfreshened heifers.
# Enrollment is judged renumber-aware via enrolled_ids (see the setup chunk).
lame_lesion_recent_all <- lame_lesion_recent_all |>
  left_join(events_all |> select(id_animal, id) |> distinct(), by = "id_animal") |>
  mutate(enrolled_smartsight = as.character(id) %in% enrolled_ids)

# Counted at the SAME grain as q4 (one row per cow-lactation, after the
# first-qualifying-case slice), so that nrow(q4) + this reconciles against the
# pre-exclusion total. Counting raw lesion rows instead double-counts a cow
# with two qualifying events.
excluded_not_enrolled_cases <- lame_lesion_recent_all |>
  filter(!is.na(date_fresh),
         is.na(date_dry) | date_event <= date_dry,
         days_since_last_trim > params$recent_trim_suppression_days,
         !enrolled_smartsight) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup()
n_excluded_not_enrolled <- nrow(excluded_not_enrolled_cases)

lame_lesion_recent <- lame_lesion_recent_all |>
  filter(!is.na(date_fresh)) |>
  filter(is.na(date_dry) | date_event <= date_dry) |>
  filter(days_since_last_trim > params$recent_trim_suppression_days) |>
  filter(enrolled_smartsight) |>
  select(-enrolled_smartsight, -id) |>
  select(-date_fresh, -date_dry, -days_since_last_trim) |>
  # first qualifying lesion case per cow-lactation only
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup()

nedlame_all <- events_all |>
  filter(Event == "NEDLAME") |>
  select(id_animal, lact_number, ned_date = date_event) |>
  distinct()

q4 <- lame_lesion_recent |>
  mutate(
    lookback_days_available = pmax(as.numeric(date_event - nedlame_start_date), 0),
    lookback_days_used = pmin(params$q4_lookback_days, lookback_days_available)
  ) |>
  left_join(nedlame_all, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(lookback_hit = !is.na(ned_date) & ned_date <= date_event & ned_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event, lookback_days_used) |>
  summarize(caught_by_nedap = any(lookback_hit), .groups = "drop") |>
  left_join(cohort |> select(id_animal, lact_number, tx_group), by = c("id_animal", "lact_number")) |>
  mutate(tx_group = if_else(is.na(tx_group), "Never Alerted (Not in Pilot Cohort)", tx_group))

n_truncated_lookback <- sum(q4$lookback_days_used < params$q4_lookback_days)

# The "Never Alerted" label is assigned by absence from `cohort`, and
# `cohort` now excludes the post-trim-artifact cows - so that bucket is NOT
# purely "never got an alert." Split it explicitly rather than describing it
# loosely: an earlier draft called the non-pure share "a small number" when
# it is actually ~18% of the bucket.
# Must match the figure's "Never Alerted" band exactly, which means carrying
# the !caught_by_nedap test too - a handful of cows are outside the cohort yet
# were still flagged in time, and they belong in "Caught in Time", not here.
# (detection_group itself is built later, so the rule is repeated rather than
# referenced; if one changes, change both.)
never_alerted_rows <- q4 |>
  filter(!caught_by_nedap, tx_group == "Never Alerted (Not in Pilot Cohort)")
n_never_alerted_bucket <- nrow(never_alerted_rows)
n_never_alerted_true <- never_alerted_rows |>
  anti_join(all_nedlame_cows |> select(id_animal, lact_number), by = c("id_animal", "lact_number")) |>
  nrow()
n_never_alerted_excluded <- n_never_alerted_bucket - n_never_alerted_true
```

A total of `r nrow(q4)` cows had a lesion diagnosed after the project start date (`r params$history_after_date`). This excludes `r n_excluded_recent_trim` cows diagnosed within `r params$recent_trim_suppression_days` days of a prior trim, when NEDAP couldn't have alerted a cow due to the set up in DairyComp. Of the `r nrow(q4)` cows, `r n_truncated_lookback` had a lookback window shorter than the full `r params$q4_lookback_days` days (because they fell within `r params$q4_lookback_days` days of NEDLAME go-live).

Of those `r nrow(q4)` cows, `r sum(!q4$caught_by_nedap)` (`r round(100 * mean(!q4$caught_by_nedap), 1)`%) had no NEDAP alert entered in DairyComp in the `r params$q4_lookback_days` days before their lesion was diagnosed.

## What Lesions are being missed?

The figure below shows the lesion types found at these first-diagnosis cases, broken out by NEDAP detection status:

- **Never Alerted** — no usable NEDLAME alert. Most were never alerted at all (`r n_never_alerted_true` of `r n_never_alerted_bucket`); the other `r n_never_alerted_excluded` had an alert that was excluded because it came after trimming.
- **Alerted, Missed** — she was alerted at some point, but not in the lookback window before *this* lesion. Some of these came after her first lesion.
- **Alerted, Caught in Time** — an alert landed in the window before the lesion. NEDAP flagged her first.

Where a cow had two different lesions on the same day, each type is counted. The same type coded twice on one day counts once.

From the figure below it is clear that DD and other are the main categories being missed. Other on this farm is typically Corkscrews.

```{r}
#| label: q4-lesion-types

lesion_type_cols <- c("dd", "footrot", "wld", "sole_ulcer", "injury", "cork",
                       "hemorrhage", "sole_fracture", "toe_ulcer", "thin", "other")

q4_group <- q4 |>
  # Order matters: test caught_by_nedap FIRST, then cohort membership.
  #
  # caught_by_nedap looks at ANY NEDLAME event, while cohort membership also
  # requires a valid MNFRS and survival of the post-trim exclusion. So a cow
  # can be flagged in time yet sit outside the cohort. Testing cohort
  # membership first labelled those cows "Never Alerted" even though an alert
  # did reach DairyComp inside the window - 4 cases, and it stopped the two
  # missed categories summing to the miss total (55 + 288 = 343, not 339).
  #
  # NB when editing: compare tx_group against the value it actually holds (set
  # in q4-build), not the short display label - the short label matches nothing
  # and silently dumps every never-alerted cow into "Alerted, Missed".
  mutate(detection_group = case_when(
    caught_by_nedap ~ "Alerted, Caught in Time",
    tx_group == "Never Alerted (Not in Pilot Cohort)" ~ "Never Alerted",
    TRUE ~ "Alerted, Missed"
  ))

# The two missed categories must account for every miss. If this ever fails,
# the labels have drifted away from caught_by_nedap again.
stopifnot(
  sum(q4_group$detection_group %in% c("Never Alerted", "Alerted, Missed")) ==
    sum(!q4_group$caught_by_nedap)
)

# and the prose counters above must describe the same band as the figure
stopifnot(sum(q4_group$detection_group == "Never Alerted") == n_never_alerted_bucket)

q4_lesion_types <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(q4_group |> select(id_animal, lact_number, date_event, detection_group),
             by = c("id_animal", "lact_number", "date_event")) |>
  pivot_longer(cols = all_of(lesion_type_cols), names_to = "lesion_type", values_to = "has_lesion") |>
  filter(has_lesion == 1) |>
  count(detection_group, lesion_type) |>
  # Shown as a share within each lesion type, not as counts. On a count axis a
  # common lesion looks like a detection problem simply because it is common -
  # the same base-rate trap the DD-by-foot table further down was built to
  # avoid. The case count is printed beside each bar so a lesion with only a
  # handful of cases can be discounted rather than read as a real rate.
  group_by(lesion_type) |>
  mutate(total_cases = sum(n), pct = 100 * n / total_cases) |>
  ungroup() |>
  mutate(lesion_type = fct_reorder(lesion_type, total_cases))

lesion_totals <- q4_lesion_types |> distinct(lesion_type, total_cases)

ggplot(q4_lesion_types, aes(x = lesion_type, y = pct, fill = detection_group)) +
  geom_col(width = 0.7) +
  geom_text(data = lesion_totals, inherit.aes = FALSE,
            aes(x = lesion_type, y = 101, label = paste0("n=", total_cases)),
            hjust = 0, size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 100, 25), labels = function(x) paste0(x, "%"),
                     limits = c(0, 118), expand = expansion(mult = c(0, 0))) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases with that lesion", fill = "NEDAP Detection Status") +
  theme(legend.position = "bottom")
```

## Which leg is being missed?

The same cases split by which foot the lesion was found on. A cow can have lesions on more than one foot, so the bars add to more than `r nrow(q4_group)` cases.

```{r}
#| label: fig-missed-by-leg
#| fig-height: 4

leg_names <- c(LF = "Left Front", RF = "Right Front", LR = "Left Rear", RR = "Right Rear")

# locate_lesion packs one or more two-character foot codes into one string
# ("RRLR" = right rear + left rear, "RRLRRFLF" = all four). It must be parsed
# in 2-character chunks, NOT with substring matching: str_detect("LRRF", "RR")
# is TRUE, but "LRRF" is LR + RF, so substring matching would invent a
# right-rear lesion that was never recorded. Checked every value in the data -
# all are even length and every chunk is one of LF/RF/LR/RR.
leg_long <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM"), !is.na(locate_lesion), locate_lesion != "") |>
  select(id_animal, lact_number, date_event, locate_lesion) |>
  mutate(leg = str_extract_all(locate_lesion, ".{2}")) |>
  unnest(leg) |>
  distinct(id_animal, lact_number, date_event, leg) |>
  # Map to readable names HERE, once, so everything downstream sees
  # "Left Rear" rather than "LR". Leaving the raw codes in place and
  # renaming later silently broke a front/rear split that tested
  # str_detect(leg, "Rear") - it matched nothing and called every foot Front.
  mutate(leg = unname(leg_names[leg]))

q4_legs <- leg_long |>
  inner_join(q4_group |> select(id_animal, lact_number, date_event, detection_group),
             by = c("id_animal", "lact_number", "date_event")) |>
  count(detection_group, leg) |>
  # Same reasoning as the lesion-type figure above: a share within each foot,
  # so the four feet are comparable regardless of how many lesions each carries.
  group_by(leg) |>
  mutate(total_cases = sum(n), pct = 100 * n / total_cases) |>
  ungroup() |>
  mutate(leg = fct_reorder(leg, total_cases))

leg_totals <- q4_legs |> distinct(leg, total_cases)

# leg is a factor now (ordered for the plot), and str_detect() rejects factors
n_legs_rear <- sum(q4_legs$n[str_detect(as.character(q4_legs$leg), "Rear")])
n_legs_front <- sum(q4_legs$n[str_detect(as.character(q4_legs$leg), "Front")])

ggplot(q4_legs, aes(x = leg, y = pct, fill = detection_group)) +
  geom_col(width = 0.7) +
  geom_text(data = leg_totals, inherit.aes = FALSE,
            aes(x = leg, y = 101, label = paste0("n=", total_cases)),
            hjust = 0, size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 100, 25), labels = function(x) paste0(x, "%"),
                     limits = c(0, 118), expand = expansion(mult = c(0, 0))) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases on that foot", fill = "NEDAP Detection Status") +
  theme(legend.position = "bottom")
```

Rear feet account for `r n_legs_rear` of the `r n_legs_rear + n_legs_front` foot-level lesions, front feet `r n_legs_front`. The never-alerted share is the largest block on every foot, so the misses are not concentrated on one limb.

## Alert Check: True Misses vs. Pipeline Losses

The figure above can only answer "did a `NEDLAME` event land in DairyComp in time". However, it can't tell a true alert miss apart from a case where the cow was flagged but the flag never reached DairyComp. The impact of this was checked with raw flag data from Nedap for the following time period (`r format(nedap_coverage_start)` to `r format(nedap_coverage_end)`).

The figure below compares among lesions what the reasons for not being flagged were. Some pipeline issues exist (likely low alerts with a score \>30) but in the never alerted section the majority are ones that where never flagged.

```{r}
#| label: q4-sensor-build

# Coverage is capped at nedap_coverage_end - the raw export is a one-time
# pull, not a live feed. The lookback window only looks BACKWARD from
# date_event, so a case is fully checkable as long as date_event itself
# is on/before the coverage end date; cases after that are flagged as
# "Not Checked" rather than silently miscounted as true misses just
# because we don't have sensor data that far out yet.
q4_sensor <- q4_group |>
  mutate(in_coverage = date_event <= nedap_coverage_end)

# Same pre-filter-before-join pattern used throughout this report:
# restrict candidates to the checkable subset FIRST, then join - a
# left_join on the full q4_sensor table followed by a filter would risk
# silently dropping the "Not Checked" rows instead of correctly keeping
# and labeling them.
sensor_hits <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> select(id_animal, lact_number, attention_date),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(sensor_hit = !is.na(attention_date) &
           attention_date <= date_event &
           attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(sensor_fired_in_window = any(sensor_hit), .groups = "drop")

q4_sensor <- q4_sensor |>
  left_join(sensor_hits, by = c("id_animal", "lact_number", "date_event")) |>
  mutate(
    # The strings in case_when() and in levels= must match EXACTLY or the
    # non-matching ones become NA and vanish from every count and plot.
    detection_group_3way = factor(case_when(
      !in_coverage ~ "No Data (Outside Coverage)",
      caught_by_nedap ~ "Flagged, Caught in Time",
      sensor_fired_in_window ~ "Flagged, Lost in Pipeline",
      TRUE ~ "Never Flagged (True Miss)"
    ), levels = c(
      "Flagged, Caught in Time", "Flagged, Lost in Pipeline",
      "Never Flagged (True Miss)", "No Data (Outside Coverage)"
    ))
  )

misses_in_coverage <- q4_sensor |> filter(in_coverage, !caught_by_nedap)
n_pipeline_loss <- sum(misses_in_coverage$detection_group_3way == "Flagged, Lost in Pipeline")
n_true_miss <- sum(misses_in_coverage$detection_group_3way == "Never Flagged (True Miss)")
pct_pipeline_loss <- round(100 * n_pipeline_loss / nrow(misses_in_coverage), 1)
n_outside_coverage <- sum(!q4_sensor$in_coverage)
# Misses specifically that fall outside coverage - the figure that belongs
# next to nrow(misses_in_coverage). n_outside_coverage counts ALL lesion
# cases outside coverage (caught ones included), so quoting it alongside a
# miss-denominated number mixes two bases.
n_misses_outside_coverage <- sum(!q4_sensor$in_coverage & !q4_sensor$caught_by_nedap)

# Sensitivity on is_completed. The raw export marks some attention episodes
# as not completed, and the loader does not filter them out, so an incomplete
# episode currently counts as the camera having flagged her. Rather than make
# that choice silently, re-run the same window test using only completed
# episodes and report how many pipeline losses depend on the incomplete ones.
sensor_hits_completed <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> filter(is_completed) |> select(id_animal, lact_number, attention_date),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(sensor_hit = !is.na(attention_date) &
           attention_date <= date_event &
           attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(fired_completed = any(sensor_hit), .groups = "drop")

n_pipeline_loss_completed_only <- misses_in_coverage |>
  filter(detection_group_3way == "Flagged, Lost in Pipeline") |>
  left_join(sensor_hits_completed, by = c("id_animal", "lact_number", "date_event")) |>
  summarize(n = sum(coalesce(fired_completed, FALSE))) |>
  pull(n)
n_pipeline_loss_incomplete_only <- n_pipeline_loss - n_pipeline_loss_completed_only
n_attentions_incomplete <- sum(!attentions_resolved$is_completed, na.rm = TRUE)
```

```{r}
#| label: q4-sensor-plot
#| fig-height: 4

q4_sensor |>
  count(tx_group, detection_group_3way) |>
  ggplot(aes(x = tx_group, y = n, fill = detection_group_3way)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_viridis_d() +
  guides(fill = guide_legend(nrow = 2)) +
  theme_minimal() +
  labs(x = NULL, y = "% of Lesion Cases", fill = "Alert Status") +
  theme(legend.position = "bottom")
```

Further investigation of the missed cows showed:

- Of the `r nrow(misses_in_coverage)` DC-recorded misses that fall within the raw sensor's coverage window, `r n_pipeline_loss` (`r pct_pipeline_loss`%) actually had an alert within the lookback window.

  - The flag never got into DairyComp as a `NEDLAME` event (potential pipeline issue)

- The remaining `r n_true_miss` had no flag at all in that window, a true miss.

`r if (n_misses_outside_coverage > 0) paste0("- The other ", n_misses_outside_coverage, " misses were diagnosed after ", format(nedap_coverage_end), ", which the current export can't see. These are left unclassified rather than assumed.") else paste0("Every miss here is checkable against the raw sensor feed: the analysis stops at ", format(analysis_end), ", the last day the Nedap export covers.")`

One judgement call sits behind these numbers. `r n_attentions_incomplete` of the resolved flags are marked not completed in the raw export, and they are counted here as the camera having flagged the cow. `r n_pipeline_loss_incomplete_only` of the `r n_pipeline_loss` pipeline losses rest only on such a flag; counting completed flags alone would give `r n_pipeline_loss_completed_only` instead. The choice matters at the margin, so it is stated rather than assumed.

## What is the camera actually missing?

```{r}
#| label: true-miss-build

checkable <- q4_sensor |>
  filter(in_coverage) |>
  mutate(camera = if_else(detection_group_3way == "Never Flagged (True Miss)",
                          "Camera never flagged", "Camera flagged her"))

n_cam_miss <- sum(checkable$camera == "Camera never flagged")
n_cam_flag <- sum(checkable$camera == "Camera flagged her")

lesion_by_camera <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(checkable |> select(id_animal, lact_number, date_event, camera),
             by = c("id_animal", "lact_number", "date_event")) |>
  pivot_longer(all_of(lesion_type_cols), names_to = "lesion_type", values_to = "has_lesion") |>
  filter(has_lesion == 1) |>
  count(camera, lesion_type) |>
  mutate(pct = if_else(camera == "Camera never flagged", 100 * n / n_cam_miss, 100 * n / n_cam_flag))

pct_dd_miss  <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "dd"  & lesion_by_camera$camera == "Camera never flagged"], 0)
pct_dd_flag  <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "dd"  & lesion_by_camera$camera == "Camera flagged her"], 0)
pct_wld_miss <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "wld" & lesion_by_camera$camera == "Camera never flagged"], 0)
pct_wld_flag <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "wld" & lesion_by_camera$camera == "Camera flagged her"], 0)

# Front/rear share within each group, computed here rather than inline so the
# prose stays readable.
leg_end <- leg_long |>
  mutate(end = if_else(str_detect(leg, "Rear"), "Rear", "Front")) |>
  distinct(id_animal, lact_number, date_event, end)

leg_by_camera <- leg_end |>
  inner_join(checkable |> select(id_animal, lact_number, date_event, camera),
             by = c("id_animal", "lact_number", "date_event")) |>
  count(camera, end) |>
  group_by(camera) |>
  mutate(pct = 100 * n / sum(n)) |>
  ungroup()

pct_rear_miss <- round(leg_by_camera$pct[leg_by_camera$camera == "Camera never flagged" & leg_by_camera$end == "Rear"], 0)
pct_rear_flag <- round(leg_by_camera$pct[leg_by_camera$camera == "Camera flagged her"   & leg_by_camera$end == "Rear"], 0)

# The marginal splits above (DD vs not, rear vs front) each hint at the answer
# but neither isolates it, because DD is mostly a rear-foot disease to begin
# with - so "DD is over-represented" and "rear feet are over-represented"
# could be one fact reported twice. Crossing the two is what separates them:
# a case is scored on whether it had DD on a REAR foot specifically, and the
# same share is computed for the cases the camera did flag.
dd_cases_all <- lame_lesion_recent |> filter(dd == 1) |> distinct(id_animal, lact_number, date_event)
case_key <- function(d) paste(d$id_animal, d$lact_number, d$date_event)

dd_rear_keys  <- case_key(leg_end |> filter(end == "Rear")  |> semi_join(dd_cases_all, by = c("id_animal","lact_number","date_event")))
dd_front_keys <- case_key(leg_end |> filter(end == "Front") |> semi_join(dd_cases_all, by = c("id_animal","lact_number","date_event")))
nodd_rear_keys <- case_key(leg_end |> filter(end == "Rear") |> anti_join(dd_cases_all, by = c("id_animal","lact_number","date_event")))

conj_share <- function(keys) {
  k <- case_key(checkable)
  c(miss    = round(100 * mean(k[checkable$camera == "Camera never flagged"] %in% keys), 1),
    flagged = round(100 * mean(k[checkable$camera == "Camera flagged her"]   %in% keys), 1))
}
s_ddrear  <- conj_share(dd_rear_keys)
s_ddfront <- conj_share(dd_front_keys)
s_noddrear <- conj_share(nodd_rear_keys)

conj_tbl <- tibble::tribble(
  ~`Lesion and foot`,            ~`% of misses`,      ~`% of flagged`,
  "DD on a rear foot",           s_ddrear[["miss"]],   s_ddrear[["flagged"]],
  "DD on a front foot",          s_ddfront[["miss"]],  s_ddfront[["flagged"]],
  "Anything else on a rear foot", s_noddrear[["miss"]], s_noddrear[["flagged"]]
) |>
  mutate(Difference = round(`% of misses` - `% of flagged`, 1))
```

The `r n_cam_miss` cases the camera never flagged are compared below against the `r n_cam_flag` it did flag. Both sit inside the sensor coverage window, so both have raw flag data behind them. Percentages are within each group, and a cow can have more than one lesion type and more than one affected foot, so they add to more than 100%.

```{r}
#| label: fig-true-miss-lesion
#| fig-height: 4

lesion_by_camera |>
  ggplot(aes(x = fct_reorder(lesion_type, pct, .fun = max), y = pct, fill = camera)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases in that group", fill = NULL) +
  theme(legend.position = "bottom")
```

- **Digital dermatitis is over-represented**: `r pct_dd_miss`% of the cases the camera never flagged, against `r pct_dd_flag`% of the ones it did flag.
- **White line is the opposite**: `r pct_wld_miss`% of misses vs `r pct_wld_flag`% of flagged cases. The camera picks white line up well.


- Rear feet are `r pct_rear_miss`% of the missed cases against `r pct_rear_flag`% of the flagged ones. That gap is small, and on its own it does not say much.

Those two splits look like two findings but may be one, since DD is mostly a rear-foot disease anyway. Crossing them separates the two:

```{r}
#| label: tbl-dd-by-foot

conj_tbl |>
  gt() |>
  tab_header(title = "The gap is DD on a rear foot specifically") |>
  tab_source_note("A case counts in a row if it had that lesion on that foot; a cow with several lesions or feet can count in more than one row, so the rows do not sum to 100%.")
```

- **DD on a rear foot is `r s_ddrear[["miss"]]`% of misses against `r s_ddrear[["flagged"]]`% of flagged cases** — a `r round(s_ddrear[["miss"]] - s_ddrear[["flagged"]], 1)` point gap, and the whole effect.
- **DD on a front foot is `r s_ddfront[["miss"]]`% against `r s_ddfront[["flagged"]]`%** — no difference. So it is not DD as such.
- **Anything other than DD on a rear foot is `r s_noddrear[["miss"]]`% against `r s_noddrear[["flagged"]]`%**, meaning those are caught *better* than average. So it is not rear feet as such either.

The blind spot is the combination: **digital dermatitis on a hind foot.** Neither DD anywhere nor a hind foot with anything else shows the same gap.

A likely reason, though this data cannot prove it: DD is often bilateral, and a cow sore on both hind feet walks symmetrically. A camera scoring mobility off asymmetry would be expected to under-detect exactly that. Worth putting to Nedap.

Since we are running a separate SmartSight instance, I wanted to make sure all cows were visible on it. A cow absent from SmartSight is not being watched for lameness at all, so a lesion found on her cannot be a camera miss.

```{r}
#| label: platform-enrollment-build

q4_sensor <- q4_sensor |>
  left_join(events_all |> select(id_animal, id) |> distinct(), by = "id_animal") |>
  mutate(
    # alias-aware, matching the Q4 eligibility rule exactly - a cow enrolled
    # under a previous animal number is enrolled (see enrolled_ids in setup)
    not_in_smartsight = !(id %in% enrolled_ids),
    # Checked, not assumed: the point of this section is that a cow can be
    # current on the Activity platform (which staff use daily) while absent
    # from SmartSight. Earlier drafts asserted "even though she's on
    # Activity" in prose without computing it - now computed so the claim
    # can't drift away from the data.
    in_activity = id %in% cows_activity$id
  )

# Unenrolled cows are excluded upstream in Q4, so this should now be empty.
# Kept as an assertion rather than deleted: if it ever stops being zero, the
# exclusion has silently stopped working.
n_residual_not_in_smartsight <- sum(!q4_sensor$caught_by_nedap & q4_sensor$not_in_smartsight)
n_total_misses <- sum(!q4_sensor$caught_by_nedap)

# Herd-level enrollment check. The misses above are only the unenrolled cows
# who also happened to get a lesion, so they understate the gap. This asks the
# direct question instead: of every first-lactation cow that freshened in the
# study window, what share is on the SmartSight list at all? Compared against
# the Activity list, which is the control - if both dropped together it would
# point at the export; if only SmartSight drops it points at enrollment.
ss_ids  <- unique(cows_smartsight$id)
act_ids <- unique(cows_activity$id)

l1_fresh <- animal_lactations |>
  filter(lact_number == 1, !is.na(date_fresh), date_fresh >= ymd(params$history_after_date) - 150,
         date_fresh <= analysis_end) |>
  distinct(id_animal, id, date_fresh) |>
  mutate(id = as.character(id),
         on_smartsight = id %in% ss_ids,
         on_activity   = id %in% act_ids,
         fresh_month   = format(date_fresh, "%Y-%m"))

enroll_by_month <- l1_fresh |>
  group_by(`Freshened` = fresh_month) |>
  summarize(`Cows` = n(),
            `% on SmartSight` = round(100 * mean(on_smartsight), 1),
            `% on Activity`   = round(100 * mean(on_activity), 1),
            .groups = "drop")

# the hard break: the most recent freshening among enrolled first-lactation cows
last_enrolled_fresh <- max(l1_fresh$date_fresh[l1_fresh$on_smartsight], na.rm = TRUE)
l1_after  <- l1_fresh |> filter(date_fresh > last_enrolled_fresh)
l1_before <- l1_fresh |> filter(date_fresh <= last_enrolled_fresh)
n_l1_after_unenrolled <- sum(!l1_after$on_smartsight)
pct_l1_before_enrolled <- round(100 * mean(l1_before$on_smartsight), 1)
n_l1_after_on_activity <- sum(l1_after$on_activity)

```


`r n_excluded_not_enrolled` lesion cases were excluded on this basis, and they are no longer counted as misses anywhere above.

Those are only the unenrolled cows who also happened to get a lesion, so they understate the problem. Asking the question directly - of every first-lactation cow that freshened in the window, how many are on each platform:

```{r}
#| label: tbl-enrollment-by-month

enroll_by_month |>
  gt() |>
  tab_header(title = "SmartSight enrollment of first-lactation cows stops; Activity does not") |>
  tab_source_note("Activity is the control here. Both lists were pulled on the same day, so a drop in one and not the other cannot be an export problem.")
```

- The most recent freshening among first-lactation cows who *are* on SmartSight is **`r format(last_enrolled_fresh)`**. Before that date `r pct_l1_before_enrolled`% were enrolled; after it, **`r n_l1_after_unenrolled` first-lactation cows are absent**, and `r n_l1_after_on_activity` of them are on Activity.

- Both lists were pulled on the same day, well after the cut-off, so this is not a stale-export artifact. A drop in SmartSight with Activity unaffected points at enrollment having stopped, not at data transfer.

- **Animal numbers get reassigned, and that had to be handled before this check could be trusted.** DairyComp logs an ID change as an `XID` event holding the old number, and `r nrow(id_aliases)` such changes appear in this extract. A cow renumbered after she was enrolled looks absent under her current number while being present under her old one - cow 23143, renumbered from 14347, is enrolled all along as 14347. Enrollment is therefore checked against a cow's current *and* prior numbers; `r n_rescued_by_alias` animals are only found that way.

## Pipeline Losses, How Many Are By-Design Exclusions?

Two rules limit which cows can become a `Low` alert: a score under 31, and no history of upper-leg lameness (`Leg-AboveFoot` in the trim `protocols`).

Score-at-flag isn't available in this data, so only the injury/`Leg-AboveFoot` half of this can be checked. It is shown in the figure below.

```{r}
#| label: pipeline-loss-injury-build

# Checked lifetime, same reasoning as Q5's chronicity check - this is
# about the cow's history, not scoped to her current lactation.
leg_abovefoot_injury_history <- lame_data |>
  filter(protocols == "Leg-AboveFoot" | injury == 1) |>
  select(id_animal, hist_date = date_event) |>
  distinct()

pipeline_loss_cases <- q4_sensor |> filter(detection_group_3way == "Flagged, Lost in Pipeline")

# Which alert type(s) actually fired within the lookback window for each
# pipeline-loss case - needed to restrict the injury/Leg-AboveFoot check
# to cases where only a LOW attention fired, since that's the specific
# enrollment gate this exclusion applies to (Strong/Low Decline aren't
# gated the same way, per the farm).
pipeline_loss_fired_types <- pipeline_loss_cases |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> select(id_animal, lact_number, attention_date, alert_type),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(attention_date <= date_event, attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(low_only = all(alert_type == "LOW"), .groups = "drop")

# Same pre-filter-before-join pattern used throughout this report.
injury_history_check <- pipeline_loss_cases |>
  select(id_animal, lact_number, date_event) |>
  left_join(leg_abovefoot_injury_history, by = "id_animal", relationship = "many-to-many") |>
  mutate(has_prior_injury = !is.na(hist_date) & hist_date < date_event) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(has_injury_leg_history = any(has_prior_injury), .groups = "drop")

# SECOND by-design mechanism: the FTDAT gate, evaluated AT THE ATTENTION
# DATE. Q4 already applies an FTDAT-style test, but only at the *lesion*
# date - which never asks whether the cow was eligible at the moment the
# camera actually fired. A trim falling between the attention and the lesion
# leaves the lesion-date test clean while making the attention itself
# un-enrollable. Gate widths come straight from the farm's DairyComp
# commands: FTDAT<-90 for the Low route, FTDAT<-28 for both decline routes.
trims_gate <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, gate_trim_date = date_event) |>
  distinct()

attention_gate_check <- pipeline_loss_cases |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> select(id_animal, lact_number, attention_date, alert_type),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(attention_date <= date_event, attention_date >= date_event - lookback_days_used) |>
  left_join(trims_gate, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(days_since_trim = if_else(!is.na(gate_trim_date) & gate_trim_date < attention_date,
                                    as.numeric(attention_date - gate_trim_date), Inf)) |>
  group_by(id_animal, lact_number, date_event, attention_date, alert_type) |>
  summarize(days_since_trim = min(days_since_trim), .groups = "drop") |>
  mutate(gate_days = if_else(alert_type == "LOW", 90, 28),
         attention_blocked = days_since_trim <= gate_days) |>
  group_by(id_animal, lact_number, date_event) |>
  # only "by design" if EVERY attention in the window was gated out; if even
  # one was eligible, DairyComp should have enrolled her.
  summarize(all_attentions_blocked = all(attention_blocked), .groups = "drop")

pipeline_loss_cases <- pipeline_loss_cases |>
  left_join(pipeline_loss_fired_types, by = c("id_animal", "lact_number", "date_event")) |>
  left_join(injury_history_check, by = c("id_animal", "lact_number", "date_event")) |>
  left_join(attention_gate_check, by = c("id_animal", "lact_number", "date_event")) |>
  mutate(
    all_attentions_blocked = coalesce(all_attentions_blocked, FALSE),
    injury_explained = low_only & has_injury_leg_history,
    by_design = all_attentions_blocked | injury_explained
  )

n_pipeline_loss_low_only <- sum(pipeline_loss_cases$low_only, na.rm = TRUE)
n_pipeline_loss_injury_explained <- sum(pipeline_loss_cases$injury_explained, na.rm = TRUE)
n_pipeline_loss_ftdat_blocked <- sum(pipeline_loss_cases$all_attentions_blocked, na.rm = TRUE)
n_pipeline_loss_by_design <- sum(pipeline_loss_cases$by_design, na.rm = TRUE)
n_pipeline_loss_unexplained <- nrow(pipeline_loss_cases) - n_pipeline_loss_by_design
```

```{r}
#| label: pipeline-loss-injury-plot
#| fig-height: 4

pipeline_loss_cases |>
  mutate(category = case_when(
    all_attentions_blocked ~ "By design: every flag inside its FTDAT window",
    injury_explained       ~ "By design: injury/upper-leg history blocks Low",
    !low_only              ~ "Unexplained: escalated past LOW",
    TRUE                   ~ "Unexplained: LOW-only, no injury history"
  )) |>
  count(category) |>
  ggplot(aes(x = fct_reorder(category, n), y = n, fill = category)) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "N Pipeline-Loss Cases") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

Two of the enrollment rules can be checked. Of the `r nrow(pipeline_loss_cases)` pipeline-loss cases:

- `r n_pipeline_loss_ftdat_blocked` had every flag land inside its own `FTDAT` window (90 days for `Low`, 28 for the declines). DairyComp was right to skip them.

- `r n_pipeline_loss_injury_explained` were `LOW`-only flags on cows with upper-leg lameness history, which blocks the `Low` route.

- That is **`r n_pipeline_loss_by_design` of `r nrow(pipeline_loss_cases)` working as designed**, leaving **`r n_pipeline_loss_unexplained` as the real pipeline losses**.

- This is a floor. We can't check score at flag, and the trim lookup only covers the current lactation while `FTDAT` likely looks back further. Both push the by-design count down.

## Of the Misses, Did Farm Staff Catch Them Anyway?

```{r}
#| label: q4-staff-catch-build

missed_cases <- q4 |> filter(!caught_by_nedap)

staff_catch <- missed_cases |>
  left_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(staff_flagged = !is.na(date_event.y) & date_event.y <= date_event.x & date_event.y >= date_event.x - 7) |>
  group_by(id_animal, lact_number, date_event = date_event.x, tx_group) |>
  summarize(staff_flagged = any(staff_flagged), .groups = "drop")
```

Of the `r nrow(staff_catch)` cows that were missed by the camera but had a lesion (trimmed due to routine trim or farm staff reasons),

- `r sum(staff_catch$staff_flagged)` (`r round(100 * mean(staff_catch$staff_flagged), 1)`%) were flagged by the farm staff in the 7 days before diagnosis.

## Summarizing the Misses

Each check above narrows a different group of cows, so the percentages don't share a denominator. The table below ties them together. Indented rows are splits of the row above, and the "% of" column always refers to the base named beside it.

```{r}
#| label: q4-reconciliation

n_caught_total <- sum(q4$caught_by_nedap)
n_staff_caught <- sum(staff_catch$staff_flagged)

q4_recon <- tibble::tribble(
  ~step, ~n, ~base, ~base_n,
  "Qualifying lesion cases (after eligibility exclusions)", nrow(q4), "—", NA_integer_,
  " Caught by NEDAP in time", n_caught_total, "qualifying cases", nrow(q4),
  " Not caught — \"misses\"", n_total_misses, "qualifying cases", nrow(q4),
  "  Within raw sensor coverage (checkable)", nrow(misses_in_coverage), "misses", n_total_misses,
  "  Outside coverage (not checkable yet)", n_misses_outside_coverage, "misses", n_total_misses,
  "   Flag fired anyway — candidate pipeline loss", n_pipeline_loss, "checkable misses", nrow(misses_in_coverage),
  "   Never flagged — true miss", n_true_miss, "checkable misses", nrow(misses_in_coverage),
  "    Of candidates: blocked by its own FTDAT window", n_pipeline_loss_ftdat_blocked, "candidates", n_pipeline_loss,
  "    Of candidates: LOW-only + injury/upper-leg history", n_pipeline_loss_injury_explained, "candidates", n_pipeline_loss,
  "    Of candidates: by design overall (either rule)", n_pipeline_loss_by_design, "candidates", n_pipeline_loss,
  "    Of candidates: UNEXPLAINED — real pipeline losses", n_pipeline_loss_unexplained, "candidates", n_pipeline_loss
) |>
  mutate(pct = if_else(is.na(base_n), NA_real_, 100 * n / base_n))

q4_recon |>
  select(step, n, pct, base) |>
  gt() |>
  fmt_number(columns = pct, decimals = 1) |>
  sub_missing(columns = pct, missing_text = "") |>
  cols_label(step = "", n = "N", pct = "% of base", base = "…of what") |>
  cols_align(align = "left", columns = step) |>
  tab_header(
    title = "Q4 Reconciliation: How the Populations Nest",
    subtitle = "Indented rows split the row above them"
  ) |>
  tab_source_note(source_note = paste0(
    "Cows never enrolled in SmartSight are excluded before this chain starts (",
    n_excluded_not_enrolled, " lesion cases), on the same basis as dry cows and unfreshened heifers - ",
    "so they are not a row here and the residual inside the chain is ", n_residual_not_in_smartsight, ". ",
    "One check does cut ACROSS the chain rather than extending it: ", n_staff_caught, " of the ", n_total_misses,
    " misses were independently flagged by staff, which can apply to any row."))
```

**The bottom line of this:**

- of the `r n_total_misses` cases DairyComp records as missed,

- only `r nrow(misses_in_coverage)` can currently be checked against the raw sensor feed at all.

  Within that checkable slice,

- `r n_true_miss` are genuine missing alerts

- `r n_pipeline_loss` are cases where the camera did flag her

  - `r n_pipeline_loss_by_design` were correct by-design suppressions (`FTDAT` window or injury/upper-leg rule), not integration failures,

  - leaving `r n_pipeline_loss_unexplained` as genuine pipeline losses — `r round(100 * n_pipeline_loss_unexplained / nrow(misses_in_coverage), 1)`% of the checkable misses.
