---
title: "NEDLAME Alert Pilot: TX (Trim) vs Control"
editor: source
format:
 html:
   embed-resources: true
   toc: true
   toc-location: left
   code-fold: true
execute:
    echo: false
    message: false
    warning: false
params:
  milk_window_pre: 14
  milk_window_post: 30
  milk_baseline_days: 7
  q4_lookback_days: 21
  history_after_date: "2026-06-01"
  recent_trim_suppression_days: 28
  post_trim_exclusion_days: 7
---

```{r}
#| label: setup

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  arrow,
  lubridate,
  gt,
  gtsummary,
  survival,
  survminer,
  ggsurvfit,
  scales,
  viridis
)

source(file.path('functions/fxn_read_milk_folder.R'))
source(file.path('functions/fxn_read_nedap_attentions.R'))

# GC's lame functions (fxn_collapse_lesions, fxn_trim_vars, fxn_dz_status) -
# fetched live from GitHub, same mechanism report_explore_lame_new.qmd uses.
# Two local overrides are sourced AFTER fxn_load_os_fxns() specifically so
# they win over the GitHub versions in the global environment:
#  - functions/fxn_code_lesions.R: the GitHub version's `trimonly` logic
#    only checks the *protocols* field for "NONE" - this farm's
#    routine/no-lesion trims are coded "NONE" in the *remark* field
#    instead, which the GitHub version doesn't check, so it silently
#    misclassifies every routine trim as having a lesion.
#  - functions/fxn_collapse_lesions.R: the GitHub version reassigns its
#    own `data` variable to an already-one-row-per-date subset before
#    computing the max() union across that date's lesion columns, so
#    when a cow has multiple event rows on the same date coding
#    DIFFERENT lesions, only whichever row survives the earlier
#    reduction keeps its lesion flags - the rest are silently dropped
#    (confirmed against a real cow with 4 same-day LAME rows: only 1 of
#    her 3 distinct lesion types survived under the GitHub version).
source(file.path('functions/fxn_load_os_fxns.R'))
fxn_load_os_fxns()
source(file.path('functions/fxn_code_lesions.R'))
source(file.path('functions/fxn_collapse_lesions.R'))

events_all <- read_parquet('data/intermediate_files/events_all_columns.parquet')
events_formatted <- read_parquet('data/intermediate_files/events_formatted.parquet')
animals <- read_parquet('data/intermediate_files/animals.parquet')
animal_lactations <- read_parquet('data/intermediate_files/animal_lactations.parquet')

date_max_pull <- max(events_all$date_event, na.rm = TRUE)
farm_name <- unique(events_all$location_event)

# NEDLAME go-live date, per the DairyComp setup: enrollment protocols were
# only turned on starting ~June 1 2026, so no NEDLAME alert can exist
# before this regardless of how "missed" a case looks.
nedlame_start_date <- min(events_all$date_event[events_all$Event == "NEDLAME"], na.rm = TRUE)
```

```{r}
#| label: cohort

# One row per cow (id_animal) that ever got a NEDLAME alert - grouped by
# id_animal alone since a cow only has one "first" NEDLAME ever. Treatment
# group is read off the MNFRS value on the FIRST NEDLAME event for that cow
# - MNFRS is stable across a cow's NEDLAME rows even though it drifts over
# her broader event history (0 before enrollment, 1/2 once enrolled).
first_nedlame_raw <- events_all |>
  filter(Event == "NEDLAME") |>
  group_by(id_animal) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  transmute(
    id_animal, id,
    first_nedlame_date = date_event,
    alert_type = Remark,
    tx_group = case_when(
      MNFRS == "1" ~ "TX (Trim)",
      MNFRS == "2" ~ "Control",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(tx_group))

# lact_number is deliberately NOT taken from the raw NEDLAME event's own
# field - right at a freshening transition (dry -> fresh, same day or one
# day apart), DC's raw event can carry a stale lact_number from the
# lactation that just ended rather than the one that just started (found
# in 2 of 718 cows once the dataset grew past ~700 alerts: their alert
# date fell after date_next_fresh/on date_fresh of the FOLLOWING
# lactation, yet the raw event said the OLD lact_number). Resolving to
# whichever lactation was actually active as of the alert date - the same
# nearest-preceding-fresh-date pattern already used for the milk join -
# fixes this without needing to trust the raw field.
first_nedlame <- first_nedlame_raw |>
  inner_join(
    animal_lactations |>
      filter(!is.na(date_fresh)) |>
      select(id_animal, lact_number, date_fresh, date_dry, date_next_fresh),
    by = "id_animal",
    relationship = "many-to-many"
  ) |>
  filter(date_fresh <= first_nedlame_date) |>
  group_by(id_animal) |>
  slice_max(date_fresh, n = 1, with_ties = FALSE) |>
  ungroup()

# Flag cows whose FIRST NEDLAME fired 1-`r params$post_trim_exclusion_days`
# days AFTER a recorded trim, same cow-lactation - per the farm's
# confirmed data-load schedule (NEDLAME's daily batch loads at ~5am and
# stamps each alert with the day it was LOADED, not necessarily the day
# the underlying sensor attention happened - an attention any time after
# ~5am on day D gets logged with date D+1), an alert dated on or before
# the trim date is a genuine independent catch, but one dated 1+ days
# AFTER the trim is very likely the camera picking up that cow's own
# post-trim gait change, not a new problem. It got through DC's own
# FTDAT suppression gate only because the trim record itself is commonly
# entered a day or two late (backdated to the actual trim date) - it
# simply wasn't in the system yet when NEDLAME's batch ran. See the new
# "Excluded" section below for the full writeup and examples.
#
# "Trim" here means LAME, FOOTRIM, *or* TRIM - same definition already
# established for Q4's FTDAT suppression gate (Round 9): a LAME visit
# IS the trim whenever the trimmer finds and codes a lesion at that
# same encounter, it just doesn't also get a separate FOOTRIM code that
# day. An earlier version of this exact chunk only checked
# FOOTRIM/TRIM, which missed exactly this case - cow 7207's LAME (White
# Line) visit on 06/11 was her actual trim, and NEDLAME fired the very
# next day (confirmed directly against her real cowcard).
trims_all <- events_all |>
  filter(Event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, trim_date = date_event) |>
  distinct()

post_trim_check <- first_nedlame |>
  select(id_animal, lact_number, first_nedlame_date) |>
  inner_join(trims_all, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(trim_date < first_nedlame_date, trim_date >= first_nedlame_date - params$post_trim_exclusion_days) |>
  group_by(id_animal, lact_number) |>
  summarize(days_since_nearest_trim = min(as.numeric(first_nedlame_date - trim_date)), .groups = "drop")

first_nedlame <- first_nedlame |>
  left_join(post_trim_check, by = c("id_animal", "lact_number")) |>
  mutate(flagged_post_trim = !is.na(days_since_nearest_trim))

# Attach lactation-stage context (date_fresh/date_dry) so every downstream
# join can be scoped to id_animal + lact_number, never crossing into a
# different lactation for the same cow.
all_nedlame_cows <- first_nedlame |>
  mutate(
    dim_at_nedlame = as.numeric(first_nedlame_date - date_fresh),
    date_lact_end = pmin(
      coalesce(date_dry, date_next_fresh, date_max_pull),
      date_max_pull,
      na.rm = TRUE
    )
  )

# `cohort` (used throughout Q1-Q3, Q5, Q6) excludes the flagged cows;
# `cohort_excluded_post_trim` keeps them, visibly, for the "Excluded"
# section - nothing is silently dropped. Q4 is deliberately NOT scoped
# to `cohort` (it pulls NEDLAME/lesion history from the whole herd
# directly), so this exclusion does not affect Q4's miss-rate numbers.
cohort <- all_nedlame_cows |> filter(!flagged_post_trim)
cohort_excluded_post_trim <- all_nedlame_cows |> filter(flagged_post_trim)

n_tx <- sum(cohort$tx_group == "TX (Trim)")
n_control <- sum(cohort$tx_group == "Control")
n_excluded_post_trim <- nrow(cohort_excluded_post_trim)
```

```{r}
#| label: lesion-history-dataset

# Same lesion-coding / chronicity pipeline as report_explore_lame_new.qmd,
# just kept at the lactation grain (lact_number carried through) so the
# "history at enrollment" lookup below never crosses lactations.
all_lesions <- c("dd", "footrot", 'wld', "sole_ulcer", "injury",
                  "cork", "other", "hemorrhage",
                  "sole_fracture", "toe_ulcer", "thin",
                  "inf", "noninf", "toe", "axial", "lesion")
diseases <- c("inf", "noninf", "lesion")

lame_data <- events_formatted |>
  select(id_animal, lact_number, date_event, event,
         remark_letters1, protocols_letters1, location_event) |>
  filter(event %in% c("LAME", "TRIM", "FOOTRIM")) |>
  fxn_code_lesions(protocol_var = protocols_letters1, remark_var = remark_letters1) |>
  filter(!is.na(trimonly)) |>
  mutate(
    noninf = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 1,
      .default = noninf),
    other = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 0,
      .default = other)
  ) |>
  fxn_collapse_lesions(lesions = all_lesions) |>
  fxn_trim_vars(location_event, id_animal, date_var = date_event, trimonly_var = trimonly) |>
  fxn_dz_status(disease_cols = diseases, event_filter = trimonly == 0)

# Each NEDLAME cow's lameness-history label as of (at/before) her first
# NEDLAME date, using her FULL lifetime record (all lactations - lame_data
# goes back to 2021), not scoped to her current lactation. Unlike every
# other join in this report, this one is intentionally NOT lactation-scoped
# per explicit user direction: chronicity is a lifetime concept, and the
# ~5 years of history available should be used rather than artificially
# resetting it at each freshening.
#
# IMPORTANT: candidate history rows are filtered to "at/before enrollment"
# BEFORE joining to cohort, not after (same reasoning as fxn_time_to_next()
# above) - filtering after a left_join would silently DROP a cow from
# `cohort` entirely instead of correctly labeling her. An earlier version
# of this exact chunk had that bug and was quietly dropping 49 of 434 cows
# (~11%) from the ENTIRE cohort - not just from Q5, but from every section,
# since `cohort` is the shared base table.
history_lookup <- lame_data |>
  select(id_animal, date_event, status_lesion) |>
  filter(!is.na(status_lesion))

valid_history <- history_lookup |>
  inner_join(
    cohort |> select(id_animal, lact_number, first_nedlame_date),
    by = "id_animal"
  ) |>
  filter(date_event <= first_nedlame_date) |>
  select(id_animal, lact_number, date_event, status_lesion)

# A cow with literally no LAME/TRIM/FOOTRIM diagnosis anywhere in her life
# before her NEDLAME alert (still ~37% of the cohort even using the full
# 5-year history) has no status_lesion to inherit. Rather than invent a
# 4th bucket for this, she's folded into "New" - the same label her
# eventual first-ever diagnosis would get, and the closest existing
# category to "no established chronic pattern."
cohort <- cohort |>
  left_join(valid_history, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  group_by(id_animal, lact_number) |>
  slice_max(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
  ungroup() |>
  mutate(history_group = if_else(is.na(status_lesion), "New", status_lesion)) |>
  select(-date_event, -status_lesion)

history_levels <- c("New", "Repeat", "Chronic")
cohort <- cohort |>
  mutate(history_group = factor(history_group, levels = history_levels))
```

```{r}
#| label: milk-data

# data/milk/ is not wired into any existing pipeline script. Raw milk rows
# only carry a raw `id` (not the composite id_animal), so they're resolved
# to a specific animal-lactation by nearest-preceding date_fresh (see
# fxn_read_milk_folder.R for why a naive date_dry window isn't safe here).
milk_long <- fxn_read_milk_folder()
milk_resolved <- fxn_resolve_milk_to_lactation(milk_long, animal_lactations)
```

```{r}
#| label: nedap-attentions-data

# Raw Nedap sensor export - what the SENSOR actually flagged, independent
# of whether that flag ever reached DairyComp as a NEDLAME event. Used in
# Q4 to split "NEDAP missed it" cases into true sensor misses vs pipeline
# losses (a real cowcard/Pulse cross-check first surfaced this pattern -
# see the plan doc's Round 8/10). This is a one-time export, not a live
# feed, so its coverage window is whatever was current when it was
# pulled - captured here explicitly rather than assumed, since it can
# lag behind the DairyComp/milk data read in above.
attentions_raw <- fxn_read_nedap_attentions()
attentions_resolved <- fxn_resolve_nedap_to_lactation(attentions_raw, animal_lactations)

nedap_coverage_start <- min(attentions_resolved$attention_date, na.rm = TRUE)
nedap_coverage_end <- max(attentions_resolved$attention_date, na.rm = TRUE)
```

# Overview

`r n_tx` cows were enrolled in the **TX (Trim)** arm (`MNFRS = 1`, automatically trimmed after alert) and `r n_control` in the **Control** arm (`MNFRS = 2`, staff decide whether/when to act), based on `r nrow(first_nedlame)` cows that ever received a `NEDLAME` alert between `r format(min(first_nedlame$first_nedlame_date))` and `r format(max(first_nedlame$first_nedlame_date))` — of whom `r n_excluded_post_trim` were excluded as likely post-trim artifacts rather than independent catches (see "Excluded" just below). This is a short (~`r round(as.numeric(max(first_nedlame$first_nedlame_date) - min(first_nedlame$first_nedlame_date))/7, 0)`-week) history since NEDLAME is a newly-added event stream, which caps every time-window analysis below.

**A methodology note that applies throughout**: wherever this report looks for "the" trim, LAME diagnosis, staff check, or NEDLAME alert for a cow, it always uses her **first** qualifying occurrence in the relevant window, not every occurrence — a cow flagged or diagnosed more than once doesn't get counted multiple times. When a single visit/date has more than one distinct lesion type coded (e.g. white line disease *and* digital dermatitis found at the same exam), each distinct type is kept; the same lesion type coded more than once on one day only counts once (see the note in Q4 for the specific chart this matters most for).

```{r}
#| label: overview-table

cohort |>
  count(tx_group, alert_type) |>
  pivot_wider(names_from = alert_type, values_from = n, values_fill = 0) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Cows per Treatment Group by Alert Type") |>
  tab_stubhead(label = "Treatment Group")
```

```{r}
#| label: dim-balance-table

cohort |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_dim = round(mean(dim_at_nedlame, na.rm = TRUE), 0),
    median_dim = round(median(dim_at_nedlame, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Lactation Stage at Enrollment",
    subtitle = "Days in milk (DIM) when first NEDLAME alert fired"
  ) |>
  cols_label(n = "N Cows", mean_dim = "Mean DIM", median_dim = "Median DIM")
```

## Excluded: Alerts Following a Recent Trim

`r n_excluded_post_trim` cows are excluded from every TX vs Control comparison below (Q1-Q3, Q5, Q6) because their **first-ever `NEDLAME` alert fired `r params$post_trim_exclusion_days` days or less *after* a recorded trim/footrim**, same cow-lactation — the sensor most likely caught that cow's own post-trim gait change, not a new, independent problem.

**Why the cutoff is "after the trim date," not "within N days of it either direction"**: `NEDLAME`'s daily batch import runs at ~5am and stamps each alert with the day it was *loaded*, not necessarily the day the underlying sensor attention happened — an attention any time after ~5am on day D gets logged with date D+1. That means an alert dated **on or before** the trim date is a genuine, independent catch (the camera flagged her, and a trim followed — exactly how the system is supposed to work), but an alert dated **any number of days after** the trim is a different story: it's very likely the camera picking up soreness/altered gait from the trim that already happened, and it only got through DC's own recent-trim suppression gate because the trim record itself is commonly entered a day or two late (correctly backdated to the actual trim date, but not yet in the system when `NEDLAME`'s batch ran that morning).

```{r}
#| label: post-trim-exclusion-table

cohort_excluded_post_trim |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_days_after_trim = round(mean(days_since_nearest_trim), 1),
    .groups = "drop"
  ) |>
  bind_rows(
    cohort_excluded_post_trim |> summarize(tx_group = "Overall", n = n(), mean_days_after_trim = round(mean(days_since_nearest_trim), 1))
  ) |>
  gt(rowname_col = "tx_group") |>
  cols_label(n = "N Excluded", mean_days_after_trim = "Mean Days After Trim") |>
  tab_header(title = paste0("Cows Excluded: First NEDLAME 1-", params$post_trim_exclusion_days, " Days After a Trim"))
```

```{r}
#| label: post-trim-exclusion-plot
#| fig-height: 4

cohort_excluded_post_trim |>
  count(alert_type, days_since_nearest_trim) |>
  ggplot(aes(x = factor(days_since_nearest_trim), y = n, fill = alert_type)) +
  geom_col() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = "Days Between Trim and NEDLAME Alert", y = "N Cows Excluded", fill = "Alert Type") +
  theme(legend.position = "bottom")
```

Consistent with the post-trim-gait theory, `r sum(cohort_excluded_post_trim$alert_type %in% c("LOW DECLINE", "STRONG DECLINE"))` of these `r n_excluded_post_trim` cows show a real decline signal (`LOW DECLINE`/`STRONG DECLINE`, not the mild `LOW` type — see chart above), and a separate check (not shown here) found the large majority never had *any* further `LAME`/`FOOTRIM`/`TRIM` event afterward — nothing else ever came of the alert, consistent with a transient, self-resolving cause rather than a new lameness case.

**Scope of this exclusion**: it only affects `cohort` (Q1-Q3, Q5, Q6's TX-vs-Control comparisons). Q4's herd-wide miss-rate analysis is deliberately unaffected — Q4 pulls `NEDLAME` and lesion history directly from the full event log, not from `cohort`, and a separate check found none of Q4's "caught by NEDAP" credits were driven by this pattern anyway.

# Q1: Milk Production Trend After First NEDLAME Alert

Cows are at very different lactation stages when flagged, so raw milk pounds aren't directly comparable across cows. Each cow's series is normalized to **% of her own pre-alert baseline** (mean production over the `r params$milk_baseline_days` days before the alert) before averaging by day and treatment group — this isolates the *shape* of the trend/recovery from each cow's absolute production level.

```{r}
#| label: q1-build

milk_cohort <- cohort |>
  inner_join(
    milk_resolved |> select(id_animal, lact_number, milk_date, dmlk1),
    by = c("id_animal", "lact_number"),
    relationship = "many-to-many"
  ) |>
  mutate(day_rel = as.numeric(milk_date - first_nedlame_date)) |>
  filter(day_rel >= -params$milk_window_pre, day_rel <= params$milk_window_post)

baseline <- milk_cohort |>
  filter(day_rel >= -params$milk_baseline_days, day_rel <= -1) |>
  group_by(id_animal, lact_number) |>
  summarize(baseline_dmlk1 = mean(dmlk1, na.rm = TRUE), .groups = "drop")

milk_norm <- milk_cohort |>
  inner_join(baseline, by = c("id_animal", "lact_number")) |>
  filter(baseline_dmlk1 >= 5) |> # drop the rare near-zero baseline (unreliable % denominator)
  mutate(pct_of_baseline = dmlk1 / baseline_dmlk1 * 100)

n_milk_cows <- n_distinct(milk_norm$id_animal)

trend <- milk_norm |>
  group_by(tx_group, day_rel) |>
  summarize(
    mean_pct = mean(pct_of_baseline, na.rm = TRUE),
    se_pct = sd(pct_of_baseline, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )
```

`r n_milk_cows` of the `r nrow(cohort)` cohort cows had at least one resolvable milk record in this window and a usable pre-alert baseline.

```{r}
#| label: q1-plot
#| fig-height: 5

ggplot(trend, aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    x = "Days Relative to First NEDLAME Alert",
    y = "Milk Production (% of Own Pre-Alert Baseline)",
    color = "Treatment Group", fill = "Treatment Group"
  )
```

## Milk Trend by Alert Type

Does the milk trend differ by how severe the Nedap alert was (`Strong Decline` vs `Low Decline` vs `Low`)?

```{r}
#| label: q1-alert-type-plot
#| fig-height: 5

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

## Lactation-Stage Balance

```{r}
#| label: q1-dim-check

cohort |>
  filter(id_animal %in% milk_norm$id_animal) |>
  ggplot(aes(x = tx_group, y = dim_at_nedlame, fill = tx_group)) +
  geom_violin(alpha = 0.5, color = NA, trim = FALSE) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.4, size = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "Days in Milk at First NEDLAME Alert")
```

# Q2: Time to Next LAME/FOOTRIM/TRIM Event

For each cow, we look for the first `LAME`, `FOOTRIM`, or `TRIM` event strictly after her first NEDLAME alert, **within the same lactation** — censored at that lactation's dry-off date or the data pull date, whichever comes first. Because MNFRS=1 cows are automatically trimmed, "time to any trim" mechanically favors TX; a second version restricts to events where a lesion was actually identified, to separate "got trimmed on schedule" from "a problem was actually found."

```{r}
#| label: q2-build

next_events <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event, event) |>
  distinct()

lesion_lookup <- lame_data |> select(id_animal, lact_number, date_event, event, lesion)

# IMPORTANT: candidate events are filtered to the valid post-NEDLAME window
# BEFORE joining to cohort, not after. Filtering after a left_join would
# silently DROP a cow entirely (rather than correctly censoring her)
# whenever she has trim/LAME events somewhere in her history but none
# inside the valid window - an earlier version of this code had exactly
# that bug and was quietly excluding ~45% of cows from the "any event"
# curve (and a different ~25% from the "with lesion" curve), which is
# also why the two curves looked identical - they weren't being computed
# over the same denominator.
fxn_time_to_next <- function(candidate_events) {
  valid_candidates <- candidate_events |>
    inner_join(
      cohort |> select(id_animal, lact_number, first_nedlame_date, date_lact_end),
      by = c("id_animal", "lact_number")
    ) |>
    filter(date_event > first_nedlame_date, date_event <= date_lact_end) |>
    select(id_animal, lact_number, date_event)

  cohort |>
    left_join(valid_candidates, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
    group_by(id_animal, lact_number) |>
    slice_min(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
    ungroup() |>
    mutate(
      event_occurred = as.integer(!is.na(date_event)),
      days_to_event = if_else(
        event_occurred == 1,
        as.numeric(date_event - first_nedlame_date),
        as.numeric(date_lact_end - first_nedlame_date)
      )
    )
}

q2_any <- fxn_time_to_next(next_events)

next_events_lesion <- next_events |>
  left_join(lesion_lookup, by = c("id_animal", "lact_number", "date_event", "event")) |>
  filter(!is.na(lesion) & lesion == 1)

q2_lesion <- fxn_time_to_next(next_events_lesion)
```

::: panel-tabset
## Time to Any Trim/LAME Event

```{r}
#| label: q2-any-plot
#| fig-height: 5

km_any <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_any)

km_any |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-any-table

km_any |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Trim/LAME Event")
```

## Time to a Trim/LAME Event *With a Lesion*

```{r}
#| label: q2-lesion-plot
#| fig-height: 5

km_lesion <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_lesion)

km_lesion |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Lesion Found") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-lesion-table

km_lesion |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Lesion Found")
```
:::

## Trim Outcomes: How Many Found Nothing?

Of the cows who had a first post-NEDLAME LAME/FOOTRIM/TRIM event (the "any" curve above), what fraction of those visits were trim-only (`trimonly` - no lesion coded) vs. an actual lesion found - by treatment group. **This table's population is only the `r nrow(cohort)`-cow NEDLAME cohort, narrowed further to those with an event inside their short post-alert observation window** — it is not the same population as Q4 below, which pulls lesion cases from the whole herd (including cows never alerted at all) with no such window restriction. Any apparent overlap between a count here and a count in Q4 is coincidental, not the same set of cows.

```{r}
#| label: q2-trimonly-summary

q2_first_event_detail <- q2_any |>
  filter(event_occurred == 1) |>
  left_join(lame_data |> select(id_animal, lact_number, date_event, event, trimonly),
            by = c("id_animal", "lact_number", "date_event"))

q2_first_event_detail |>
  group_by(tx_group) |>
  summarize(n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100, .groups = "drop") |>
  bind_rows(
    q2_first_event_detail |> summarize(tx_group = "Overall", n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100)
  ) |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_trimonly, decimals = 1) |>
  cols_label(n = "N Cows With a First Event", n_trimonly = "Trim-Only (No Lesion)", pct_trimonly = "% Trim-Only") |>
  tab_header(title = "First Post-NEDLAME Visit Outcome")
```

# Q3: % of Cows Independently Identified by Farm Staff

`CHKLAME` is the event that represents farm staff independently flagging a lame cow (distinct from the automated `NEDLAME` alert). A cow counts as **staff-identified** here if she has a `CHKLAME` event, in the same lactation, either (a) in the 7 days before her first NEDLAME alert **or on the same day as it**, or (b) any time after it — excluding cows whose post-alert `CHKLAME` came at/after she'd already been trimmed, since a trim-triggered exam isn't independent detection.

```{r}
#| label: q3-build

chklame <- events_formatted |>
  filter(event == "CHKLAME") |>
  select(id_animal, lact_number, date_event) |>
  distinct()

trims <- events_formatted |>
  filter(event %in% c("FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event) |>
  distinct()

# A CHKLAME on the SAME day as the NEDLAME alert counts as staff-identified
# too (folded into the "pre-week" bucket via <= rather than <), rather than
# falling through the gap between "before" (strict <) and "after" (strict >).
pre_week <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event >= first_nedlame_date - 7, date_event <= first_nedlame_date) |>
  distinct(id_animal, lact_number)

first_post_chklame <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  rename(chklame_date = date_event)

first_post_trim <- cohort |>
  inner_join(trims, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(id_animal, lact_number, trim_date = date_event)

post_flag <- first_post_chklame |>
  left_join(first_post_trim, by = c("id_animal", "lact_number")) |>
  filter(is.na(trim_date) | chklame_date < trim_date) |>
  distinct(id_animal, lact_number)

staff_id <- cohort |>
  mutate(
    staff_identified = as.integer(
      paste(id_animal, lact_number) %in% paste(pre_week$id_animal, pre_week$lact_number) |
      paste(id_animal, lact_number) %in% paste(post_flag$id_animal, post_flag$lact_number)
    )
  )
```

::: panel-tabset
## By Treatment Group

```{r}
#| label: q3-treatment-group-plot
#| fig-height: 4

staff_id |>
  group_by(tx_group) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = tx_group, y = pct_staff_id, fill = tx_group)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(round(pct_staff_id, 1), "%\n(n=", n, ")")), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "% Staff-Identified (CHKLAME)")
```

## By Alert Type

Does staff catch a cow on their own more often when the Nedap alert itself was more severe (`Strong Decline`) vs milder (`Low Decline`/`Low`)?

```{r}
#| label: q3-alert-type-by-group-plot
#| fig-height: 4

staff_id |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = alert_type, y = pct_staff_id, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(round(pct_staff_id, 1), "%\n(n=", n, ")")),
            position = position_dodge(width = 0.7), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  labs(x = "Alert Type", y = "% Staff-Identified (CHKLAME)", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Q4: % of Post-June-1 Lesion Cases NEDAP Didn't Catch First

Among cows with a `LAME`-diagnosed lesion on or after `r params$history_after_date`, this checks whether that cow had a `NEDLAME` alert in the preceding lookback window (same lactation). A "miss" means staff/routine trimming found a lesion that NEDAP hadn't already flagged. Each cow is counted once here, at her **first** qualifying lesion diagnosis on/after that date (consistent with using the first event throughout this report) — a cow diagnosed lame twice after June 1 doesn't get double-counted.

::: {.callout-warning}
**The table below can only see DairyComp's side of the story, and a real-world spot-check found that isn't the whole story.** Manually checking cow IDs against the Nedap/Pulse platform directly (not just what landed in DairyComp) found cases where the sensor *did* flag a cow but the flag never reached DairyComp as a `NEDLAME` event — a pipeline/integration gap, not a detection failure. (An earlier version of this caveat also flagged a possible eligibility-modeling gap, based on a cow whose recent-trim exclusion didn't look right - that's now been confirmed and fixed: the exclusion below correctly checks days since her last `LAME`/`FOOTRIM`/`TRIM` record of any kind, not just formal trims.) **The headline miss rate below is still best read as an upper bound on the true technology-miss rate, not a precise figure** — but the "Sensor-Level Check" subsection further down now uses Nedap's raw flag history to separate true sensor misses from pipeline losses directly, for the window that raw data covers. See the plan doc's "Round 8"/"Round 10" for the specific examples that motivated this.
:::

**This section's population is the whole herd**, not just the `r nrow(cohort)`-cow NEDLAME cohort from Q1-Q3 above — a cow who never received a NEDLAME alert at all can still show up here if she had a qualifying lesion diagnosis. This is deliberate: the question is how many real lameness cases NEDAP is catching across the herd, which necessarily includes cows it never flagged in the first place, not only the cows already in the pilot. Don't compare counts here directly against Q2/Q3's cohort-scoped counts.

Lesions diagnosed **on a heifer who hasn't freshened yet** (`lact_number == 0`, no `date_fresh` on record) are excluded entirely, for the same reason as the dry-off exclusion just below: NEDLAME's mobility/activity monitoring is production-linked, so a cow who was never enrolled in milking in the first place was never eligible to be caught either way.

Lesions diagnosed **after a cow was already dried off** are excluded entirely: NEDLAME's mobility/activity monitoring is production-linked, so a dry-cow lameness case isn't a fair "miss" to hold against it either way.

Lesions diagnosed **within `r params$recent_trim_suppression_days` days of the cow's previous `LAME`, `FOOTRIM`, or `TRIM` record** are also excluded entirely: per the DairyComp `NEDLAME` setup, a cow can't be enrolled for a new alert unless it's been more than `r params$recent_trim_suppression_days` days since her last lameness-related record of any of those types — not just a formal trim (this is by design — it stops a cow from being re-flagged over and over right after she's already been identified and is being addressed). Since NEDAP was structurally unable to fire an alert during that window regardless of how lame she actually was, a lesion found in that window isn't a fair "miss" either — same reasoning as the dry-off exclusion above.

NEDLAME alerting only went live on `r format(nedlame_start_date)` (per the DairyComp setup), so a fixed `r params$q4_lookback_days`-day lookback isn't fair to lesion events found in the first few weeks after go-live — there simply wasn't `r params$q4_lookback_days` days of NEDLAME history yet for NEDAP to have caught anything in. The lookback window here is **rolling**: for each lesion event it's capped at however many days NEDLAME had actually been running by that date (never more than `r params$q4_lookback_days`, but shorter for events soon after go-live), so early cases are judged against the detection window that could actually have existed, not a fixed window reaching back before the system was on.

```{r}
#| label: q4-build

# Confirmed directly with the farm: the DairyComp FTDAT gate tracks days
# since the most recent LAME diagnosis OR trim - not trims alone. A cow
# with frequent LAME diagnoses but no formal trim in between is just as
# ineligible for a new alert as a recently-trimmed one.
trims_for_suppression <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, trim_date = date_event) |>
  distinct()

lame_lesion_recent_all <- lame_data |>
  filter(event == "LAME", lesion == 1, date_event >= ymd(params$history_after_date)) |>
  left_join(animal_lactations |> select(id_animal, lact_number, date_fresh, date_dry), by = c("id_animal", "lact_number"))

# Days since this cow's most recent PRIOR trim, for each lesion case - computed
# so that a case with no prior trim at all still gets exactly one row (Inf),
# rather than a naive filter-then-min silently dropping it (see lessons-learned.md
# in mcp_server/knowledge/ for why that pattern is dangerous).
recent_trim_check <- lame_lesion_recent_all |>
  select(id_animal, lact_number, date_event) |>
  left_join(trims_for_suppression, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(days_since_trim = if_else(!is.na(trim_date) & trim_date < date_event, as.numeric(date_event - trim_date), Inf)) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(days_since_last_trim = min(days_since_trim), .groups = "drop")

lame_lesion_recent_all <- lame_lesion_recent_all |>
  left_join(recent_trim_check, by = c("id_animal", "lact_number", "date_event"))

# exclude lesions diagnosed on a heifer who never freshened (lact_number ==
# 0, no date_fresh on record - confirmed this is exactly how this herd's
# data represents "hasn't calved yet" - a real cowcard cross-check (cow
# 31377) caught this: she's lact_number 0 and was being counted as a
# "miss" despite never having been eligible for NEDAP monitoring at all),
# after the cow was already dried off, or diagnosed while NEDAP would
# have been structurally suppressed by a too-recent trim
n_excluded_prefresh <- sum(is.na(lame_lesion_recent_all$date_fresh))
n_excluded_dry <- sum(!is.na(lame_lesion_recent_all$date_fresh) & !is.na(lame_lesion_recent_all$date_dry) & lame_lesion_recent_all$date_event > lame_lesion_recent_all$date_dry)
n_excluded_recent_trim <- sum(
  !is.na(lame_lesion_recent_all$date_fresh) &
    lame_lesion_recent_all$days_since_last_trim <= params$recent_trim_suppression_days &
    (is.na(lame_lesion_recent_all$date_dry) | lame_lesion_recent_all$date_event <= lame_lesion_recent_all$date_dry)
)

lame_lesion_recent <- lame_lesion_recent_all |>
  filter(!is.na(date_fresh)) |>
  filter(is.na(date_dry) | date_event <= date_dry) |>
  filter(days_since_last_trim > params$recent_trim_suppression_days) |>
  select(-date_fresh, -date_dry, -days_since_last_trim) |>
  # first qualifying lesion case per cow-lactation only
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup()

nedlame_all <- events_all |>
  filter(Event == "NEDLAME") |>
  select(id_animal, lact_number, ned_date = date_event) |>
  distinct()

q4 <- lame_lesion_recent |>
  mutate(
    lookback_days_available = pmax(as.numeric(date_event - nedlame_start_date), 0),
    lookback_days_used = pmin(params$q4_lookback_days, lookback_days_available)
  ) |>
  left_join(nedlame_all, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(lookback_hit = !is.na(ned_date) & ned_date <= date_event & ned_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event, lookback_days_used) |>
  summarize(caught_by_nedap = any(lookback_hit), .groups = "drop") |>
  left_join(cohort |> select(id_animal, lact_number, tx_group), by = c("id_animal", "lact_number")) |>
  mutate(tx_group = if_else(is.na(tx_group), "Never Alerted (Not in Pilot Cohort)", tx_group))

n_truncated_lookback <- sum(q4$lookback_days_used < params$q4_lookback_days)
```

`r nrow(q4)` qualifying lesion cases on/after `r params$history_after_date` (one per cow, first diagnosis only; `r n_excluded_prefresh` additional cases were excluded as diagnosed on a heifer who hadn't freshened yet, `r n_excluded_dry` more as diagnosed after dry-off, and `r n_excluded_recent_trim` more as diagnosed within `r params$recent_trim_suppression_days` days of a prior trim, when NEDAP couldn't have alerted regardless), of which `r n_truncated_lookback` had a lookback window shorter than the full `r params$q4_lookback_days` days (because they fell within `r params$q4_lookback_days` days of NEDLAME go-live).

Note: `r params$recent_trim_suppression_days` days is the more permissive of the two trim-suppression gates in the DairyComp setup (`Strong Decline`/`Low Decline` enrollment both require it; `Low` enrollment requires 90 days instead, "to not have chronic cows flagged all the time"). Since a `Strong Decline` or `Low Decline` alert could still have fired past `r params$recent_trim_suppression_days` days even if a `Low` alert specifically couldn't, using the shorter window here is the conservative choice — it likely slightly *understates* how many cases were truly NEDAP-ineligible.

```{r}
#| label: q4-table

q4 |>
  group_by(tx_group) |>
  summarize(n_lesion_events = n(), pct_missed = mean(!caught_by_nedap) * 100, .groups = "drop") |>
  bind_rows(
    q4 |> summarize(tx_group = "Overall", n_lesion_events = n(), pct_missed = mean(!caught_by_nedap) * 100)
  ) |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_missed, decimals = 1) |>
  cols_label(n_lesion_events = "N Lesion Cases", pct_missed = "% Not Caught by NEDAP First") |>
  tab_header(title = paste("Lesion Cases On/After", params$history_after_date)) |>
  tab_source_note(source_note = "\"Never Alerted (Not in Pilot Cohort)\" = cows with a lesion case here whose alert history has no *valid* NEDLAME in `cohort` - almost all never received any NEDLAME at all, but as of the post-trim-artifact exclusion above, this bucket can also include a small number of cows whose only NEDLAME was itself excluded from `cohort` as a likely post-trim artifact (not a true zero-alert cow). Not broken out separately here since it doesn't change the overall miss rate, but worth knowing if the TX/Control-vs-Never-Alerted split specifically is being scrutinized closely.")
```

## What Lesions Is NEDAP Missing?

Lesion types found at these first-diagnosis cases, broken out by NEDAP detection status:

- **Never Alerted (Not in Pilot Cohort)** — this cow never received a NEDLAME alert at all, in the whole dataset. The most complete kind of miss.
- **Alerted, Missed** — this cow did get a NEDLAME alert at some point (she's in the TX or Control cohort), but not within the lookback window before *this* lesion was found. NEDAP was watching her, just not in time for this case.
- **Alerted, Caught in Time** — a NEDLAME alert fired within the lookback window before this lesion was diagnosed; NEDAP flagged it first.

If a cow had more than one distinct lesion type diagnosed on the same day (e.g. white line disease *and* digital dermatitis at the same exam), each distinct type is counted separately here; the same lesion type coded more than once on one day only counts once.

```{r}
#| label: q4-lesion-types

lesion_type_cols <- c("dd", "footrot", "wld", "sole_ulcer", "injury", "cork",
                       "hemorrhage", "sole_fracture", "toe_ulcer", "thin", "other")

q4_group <- q4 |>
  mutate(detection_group = if_else(tx_group == "Never Alerted (Not in Pilot Cohort)",
                                    "Never Alerted (Not in Pilot Cohort)",
                                    if_else(caught_by_nedap, "Alerted, Caught in Time", "Alerted, Missed")))

q4_lesion_types <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(q4_group |> select(id_animal, lact_number, date_event, detection_group),
             by = c("id_animal", "lact_number", "date_event")) |>
  pivot_longer(cols = all_of(lesion_type_cols), names_to = "lesion_type", values_to = "has_lesion") |>
  filter(has_lesion == 1) |>
  count(detection_group, lesion_type)

ggplot(q4_lesion_types, aes(x = fct_reorder(lesion_type, n, .fun = sum), y = n, fill = detection_group)) +
  geom_col() +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "Number of Lesion Cases", fill = "NEDAP Detection Status") +
  theme(legend.position = "bottom")
```

## Sensor-Level Check: True Misses vs. Pipeline Losses

The table above can only answer "did a `NEDLAME` event land in DairyComp in time" — it can't tell a true sensor miss apart from a case where the sensor fired but the flag never reached DC (the pipeline-loss pattern the callout above describes, first found via manual cowcard/Pulse cross-checks). The raw Nedap sensor export lets this report check that directly, for lesion cases within its coverage window (`r format(nedap_coverage_start)` to `r format(nedap_coverage_end)`): among cases NEDAP didn't catch in DC, did the sensor itself ever fire in the same rolling lookback window, whether or not that flag made it into DairyComp as a `NEDLAME` event?

```{r}
#| label: q4-sensor-build

# Coverage is capped at nedap_coverage_end - the raw export is a one-time
# pull, not a live feed. The lookback window only looks BACKWARD from
# date_event, so a case is fully checkable as long as date_event itself
# is on/before the coverage end date; cases after that are flagged as
# "Not Checked" rather than silently miscounted as true misses just
# because we don't have sensor data that far out yet.
q4_sensor <- q4_group |>
  mutate(in_coverage = date_event <= nedap_coverage_end)

# Same pre-filter-before-join pattern used throughout this report:
# restrict candidates to the checkable subset FIRST, then join - a
# left_join on the full q4_sensor table followed by a filter would risk
# silently dropping the "Not Checked" rows instead of correctly keeping
# and labeling them.
sensor_hits <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> select(id_animal, lact_number, attention_date),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(sensor_hit = !is.na(attention_date) &
           attention_date <= date_event &
           attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(sensor_fired_in_window = any(sensor_hit), .groups = "drop")

q4_sensor <- q4_sensor |>
  left_join(sensor_hits, by = c("id_animal", "lact_number", "date_event")) |>
  mutate(
    detection_group_3way = factor(case_when(
      !in_coverage ~ "Not Checked (Outside Raw Sensor Data Coverage)",
      caught_by_nedap ~ "Alerted, Caught in Time",
      sensor_fired_in_window ~ "Sensor Fired, Lost in Pipeline",
      TRUE ~ "Sensor Never Fired (True Miss)"
    ), levels = c(
      "Alerted, Caught in Time", "Sensor Fired, Lost in Pipeline",
      "Sensor Never Fired (True Miss)", "Not Checked (Outside Raw Sensor Data Coverage)"
    ))
  )

misses_in_coverage <- q4_sensor |> filter(in_coverage, !caught_by_nedap)
n_pipeline_loss <- sum(misses_in_coverage$detection_group_3way == "Sensor Fired, Lost in Pipeline")
n_true_miss <- sum(misses_in_coverage$detection_group_3way == "Sensor Never Fired (True Miss)")
pct_pipeline_loss <- round(100 * n_pipeline_loss / nrow(misses_in_coverage), 1)
n_outside_coverage <- sum(!q4_sensor$in_coverage)
```

```{r}
#| label: q4-sensor-plot
#| fig-height: 4

q4_sensor |>
  count(tx_group, detection_group_3way) |>
  ggplot(aes(x = tx_group, y = n, fill = detection_group_3way)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "Share of Lesion Cases", fill = "Sensor-Level Detection Status") +
  theme(legend.position = "bottom")
```

Of the `r nrow(misses_in_coverage)` DC-recorded misses that fall within the raw sensor's coverage window (`r n_outside_coverage` lesion cases fell after `r format(nedap_coverage_end)` and can't be checked this way with the current export), `r n_pipeline_loss` (`r pct_pipeline_loss`%) actually had the sensor fire within the lookback window — the flag simply never reached DairyComp as a `NEDLAME` event — and the remaining `r n_true_miss` had no sensor firing at all in that window, a true miss at the sensor level. **Within this coverage window, the technology-level miss rate is measurably lower than the DC-recorded miss rate above suggests**, though this gap shouldn't be extrapolated to cases outside the coverage window without a fresher sensor export — see the plan doc for how big the currently-checkable window is relative to the full analysis period.

## Of the Misses, Did Farm Staff Catch Them Anyway?

For the cases NEDAP didn't catch, this checks for a `CHKLAME` event (farm staff independently flagging a lame cow — see Q3) in the 7 days before, or on the same day as, the lesion diagnosis. A "yes" here means staff still caught the cow on their own even though NEDAP missed her; a "no" means neither NEDAP nor staff caught her before she was formally diagnosed.

```{r}
#| label: q4-staff-catch-build

missed_cases <- q4 |> filter(!caught_by_nedap)

staff_catch <- missed_cases |>
  left_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(staff_flagged = !is.na(date_event.y) & date_event.y <= date_event.x & date_event.y >= date_event.x - 7) |>
  group_by(id_animal, lact_number, date_event = date_event.x, tx_group) |>
  summarize(staff_flagged = any(staff_flagged), .groups = "drop")
```

```{r}
#| label: q4-staff-catch-table

staff_catch |>
  group_by(tx_group) |>
  summarize(n_missed = n(), n_staff_flagged = sum(staff_flagged), pct_staff_flagged = mean(staff_flagged) * 100, .groups = "drop") |>
  bind_rows(
    staff_catch |> summarize(tx_group = "Overall", n_missed = n(), n_staff_flagged = sum(staff_flagged), pct_staff_flagged = mean(staff_flagged) * 100)
  ) |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_staff_flagged, decimals = 1) |>
  cols_label(n_missed = "N Missed by NEDAP", n_staff_flagged = "Also Staff-Flagged (CHKLAME)", pct_staff_flagged = "% Also Staff-Flagged") |>
  tab_header(title = "Of NEDAP's Misses, How Many Did Staff Catch Independently?")
```

# Q5: Influence of Lameness History

Each cow's lameness-history label (`New` / `Repeat` / `Chronic`, from the same chronicity classifier `report_explore_lame_new.qmd` uses) reflects her **full lifetime record** as of her first NEDLAME alert — `lame_data` goes back to 2021, and this deliberately isn't restricted to her current lactation, since chronicity is a lifetime concept. A cow with no LAME/TRIM/FOOTRIM diagnosis anywhere in her life before her alert (still about a third of the cohort even with 5 years of history to draw on) is folded into `New`, the closest existing category to "no established chronic pattern" — there's no separate "never diagnosed" bucket. With only `r nrow(cohort)` cows split across 2 treatment arms x 3 history categories, cells here are thin (some under 20) — treat this section as descriptive/exploratory, not a confirmatory subgroup test. The milk-trend and time-to-event plots below are faceted by **alert type** rather than history group, since that split turned out more informative; history's own distribution is shown just below.

```{r}
#| label: q5-balance-plot
#| fig-height: 4

cohort |>
  count(tx_group, history_group) |>
  group_by(tx_group) |>
  mutate(pct = n / sum(n) * 100) |>
  ungroup() |>
  mutate(history_group = fct_relabel(history_group, \(x) str_wrap(x, width = 14))) |>
  ggplot(aes(x = history_group, y = n, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(n, " (", round(pct, 0), "%)")),
            position = position_dodge(width = 0.7), vjust = -0.3, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal() +
  labs(x = NULL, y = "N Cows", fill = "Treatment Group",
       title = "Lameness-History Distribution at Enrollment") +
  theme(legend.position = "bottom",
        axis.text.x = element_text(lineheight = 0.9),
        plot.margin = margin(10, 15, 10, 15))
```

## Milk Trend by Alert Type

```{r}
#| label: q5-milk-plot
#| fig-height: 6

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

## Time to Next Trim/LAME Event by Alert Type

```{r}
#| label: q5-km-plot
#| fig-height: 6

alert_levels <- c("LOW", "LOW DECLINE", "STRONG DECLINE")

# ggsurvfit only exposes a single combined `strata` column, so a KM plot
# faceted by alert_type is built manually: fit one survfit2() per
# alert level (skipping any level without both treatment arms
# represented) and stack the tidy() estimates before plotting.
fittable_alert_levels <- alert_levels |>
  keep(\(a) {
    d <- q2_any |> filter(alert_type == a)
    n_distinct(d$tx_group) == 2 && nrow(d) >= 10
  })

q5_km_data <- fittable_alert_levels |>
  map(\(a) {
    d <- q2_any |> filter(alert_type == a)
    fit <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = d)
    broom::tidy(fit) |>
      mutate(
        tx_group = str_remove(strata, "^tx_group="),
        alert_type = a,
        risk = 1 - estimate
      )
  }) |>
  list_rbind() |>
  mutate(alert_type = factor(alert_type, levels = alert_levels))

skipped_alert_levels <- setdiff(alert_levels, fittable_alert_levels)

ggplot(q5_km_data, aes(x = time, y = risk, color = tx_group)) +
  geom_step(linewidth = 1) +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

`r if (length(skipped_alert_levels) > 0) paste0("*Alert type(s) omitted for insufficient data in one or both arms: ", paste(skipped_alert_levels, collapse = ", "), ".*")`

# Q6: Time to Culling After NEDAP Enrollment

Same culling-outcome pattern as `report_explore_lame_new.qmd`, anchored on the first NEDLAME date instead of a lesion-event date.

```{r}
#| label: q6-build

q6 <- cohort |>
  left_join(animals |> select(id_animal, date_sold, date_died, date_left), by = "id_animal") |>
  mutate(
    culled_ever = as.integer(!is.na(date_left)),
    days_to_censor_cull = if_else(
      culled_ever == 1,
      as.numeric(date_left - first_nedlame_date),
      as.numeric(date_max_pull - first_nedlame_date)
    )
  ) |>
  filter(days_to_censor_cull >= 0)

km_cull <- survfit2(Surv(days_to_censor_cull, culled_ever) ~ tx_group, data = q6)
```

```{r}
#| label: q6-plot
#| fig-height: 5

km_cull |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% Culled") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q6-table

km_cull |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% Culled After First NEDLAME Alert")
```

```{r}
#| label: mnfrs-caveat-stat

# Pulled directly from Q2's own km_any object (not a separately hand-computed
# figure) so this can't drift out of sync with the rest of the report again -
# an earlier draft of this caveat cited a hand-typed 81.5%/42.9% figure that
# didn't reproduce against the current pipeline once the join-order bugs
# above were fixed.
km_any_tidy <- broom::tidy(km_any) |>
  mutate(tx_group_label = str_remove(strata, "^tx_group="), risk_pct = (1 - estimate) * 100)

pct_trimmed_14d_tx <- km_any_tidy |> filter(tx_group_label == "TX (Trim)", time <= 14) |> slice_max(time, n = 1) |> pull(risk_pct)
pct_trimmed_14d_control <- km_any_tidy |> filter(tx_group_label == "Control", time <= 14) |> slice_max(time, n = 1) |> pull(risk_pct)
```

# Caveats

- **MNFRS coding is now confirmed from the DairyComp setup documentation**, not just inferred: the `NDTLIST` command explicitly enrolls `MNFRS = 1` cows onto the trim list ("put nedap cows on trim list"), matching the empirical pattern found here (per Q2's own KM estimate, `r round(pct_trimmed_14d_tx, 0)`% of `MNFRS = 1` cows have a subsequent LAME/FOOTRIM/TRIM event within 14 days of their alert vs `r round(pct_trimmed_14d_control, 0)`% for `MNFRS = 2`).
- **`Chronic` here just means "not her first-ever diagnosis, and more than a day since the last one"** — the external chronicity classifier's `Repeat` vs. `Chronic` cutoff is a 1-day gap, so almost any cow with 2+ lifetime lesion diagnoses lands in `Chronic` regardless of whether those diagnoses were weeks or years apart (`r sum(cohort$history_group == "Chronic")` of `r nrow(cohort)` cows are `Chronic` here, only `r sum(cohort$history_group == "Repeat")` are `Repeat`). Don't read `Chronic` as "actively/currently recurring" — it doesn't distinguish a cow with one old resolved case 4 years ago from one with three flare-ups this year.
- **Two data-processing bugs were found and fixed while building this report**, both of which materially changed earlier draft results:
  1. The externally-fetched `fxn_collapse_lesions()`/`fxn_trim_vars()`/`fxn_dz_status()` functions (pulled live from GitHub) come bundled with their *own* `fxn_code_lesions()`, whose `trimonly` logic only checks the `protocols` field for "NONE" - this farm's routine/no-lesion trims are coded "NONE" in the `remark` field instead, which that version doesn't check. The GitHub function was silently overwriting this repo's own (correctly-calibrated) `functions/fxn_code_lesions.R`, so *every* LAME/FOOTRIM/TRIM row was being misclassified as having a lesion. Fixed by sourcing the local file last so it wins. This is why Q2's "any trim" and "with a lesion" curves were identical in an earlier draft.
  2. Q2's time-to-next-event join was filtering candidate events to the valid post-NEDLAME window *after* joining to the cohort rather than before, which silently dropped a cow from the analysis entirely (instead of correctly counting her as censored) whenever she had trim/LAME events somewhere in her history but none inside the valid window. This was quietly excluding ~45% of cows from the "any event" curve.
- **Short history**: `NEDLAME` only spans `r as.numeric(max(first_nedlame$first_nedlame_date) - min(first_nedlame$first_nedlame_date))` days of data (go-live `r format(nedlame_start_date)`), capping every time-to-event and milk-trend window above — results should be treated as an early read, not a mature-pilot conclusion. Q4's lookback window is rolling for this same reason (see Q4).
- **Small sample for Q5**: 2 treatment arms x 3 history categories over `r nrow(cohort)` cows produces thin cells (`Repeat` has only `r sum(cohort$history_group == "Repeat")` cows total); the two arms are not perfectly balanced on prior lameness history at enrollment either (see the Q5 balance chart), which should temper causal interpretation of any history-stratified differences.
- **Milk join resolution**: `r round(100 * n_milk_cows / nrow(cohort), 1)`% of cohort cows had a usable milk baseline in the analysis window; the remainder lacked resolvable milk records (e.g. dried off, sold, or outside the pulled milk-file date range).
- **Post-trim alert exclusion (Q1-Q3, Q5, Q6 only)**: `r n_excluded_post_trim` cows whose first `NEDLAME` fired within `r params$post_trim_exclusion_days` days after a recorded trim were excluded from the TX-vs-Control cohort as likely post-treatment gait artifacts rather than independent catches — see the "Excluded" section above for the full reasoning (tied to `NEDLAME`'s 5am batch-load schedule and trims commonly being entered a day or two late). This is a judgment call, not a certainty: it's possible a small number of these really were new, coincidentally-timed problems. Q4 is unaffected, since it doesn't draw from `cohort`.
