---
title: "NEDLAME Alert Pilot: TX (Trim) vs Control"
editor: source
format:
 html:
   embed-resources: true
   toc: true
   toc-location: left
   code-fold: true
execute:
    echo: false
    message: false
    warning: false
params:
  milk_window_pre: 14
  milk_window_post: 30
  milk_baseline_days: 7
  q4_lookback_days: 21
  history_after_date: "2026-06-01"
---

```{r}
#| label: setup

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  arrow,
  lubridate,
  gt,
  gtsummary,
  survival,
  survminer,
  ggsurvfit,
  scales,
  viridis
)

source(file.path('functions/fxn_code_lesions.R'))
source(file.path('functions/fxn_read_milk_folder.R'))

# GC's lame functions (fxn_collapse_lesions, fxn_trim_vars, fxn_dz_status) -
# fetched live from GitHub, same mechanism report_explore_lame_new.qmd uses
source(file.path('functions/fxn_load_os_fxns.R'))
fxn_load_os_fxns()

events_all <- read_parquet('data/intermediate_files/events_all_columns.parquet')
events_formatted <- read_parquet('data/intermediate_files/events_formatted.parquet')
animals <- read_parquet('data/intermediate_files/animals.parquet')
animal_lactations <- read_parquet('data/intermediate_files/animal_lactations.parquet')

date_max_pull <- max(events_all$date_event, na.rm = TRUE)
farm_name <- unique(events_all$location_event)
```

```{r}
#| label: cohort

# One row per cow-lactation that ever got a NEDLAME alert. Treatment group
# is read off the MNFRS value on the FIRST NEDLAME event for that cow -
# MNFRS is stable across a cow's NEDLAME rows even though it drifts over
# her broader event history (0 before enrollment, 1/2 once enrolled).
first_nedlame <- events_all |>
  filter(Event == "NEDLAME") |>
  group_by(id_animal) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  transmute(
    id_animal, id, lact_number,
    first_nedlame_date = date_event,
    alert_type = Remark,
    tx_group = case_when(
      MNFRS == "1" ~ "TX (Trim)",
      MNFRS == "2" ~ "Control",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(tx_group))

# Attach lactation-stage context (date_fresh/date_dry) so every downstream
# join can be scoped to id_animal + lact_number, never crossing into a
# different lactation for the same cow.
cohort <- first_nedlame |>
  left_join(
    animal_lactations |>
      select(id_animal, lact_number, date_fresh, date_dry, date_next_fresh),
    by = c("id_animal", "lact_number")
  ) |>
  mutate(
    dim_at_nedlame = as.numeric(first_nedlame_date - date_fresh),
    date_lact_end = pmin(
      coalesce(date_dry, date_next_fresh, date_max_pull),
      date_max_pull,
      na.rm = TRUE
    )
  )

n_tx <- sum(cohort$tx_group == "TX (Trim)")
n_control <- sum(cohort$tx_group == "Control")
```

```{r}
#| label: lesion-history-dataset

# Same lesion-coding / chronicity pipeline as report_explore_lame_new.qmd,
# just kept at the lactation grain (lact_number carried through) so the
# "history at enrollment" lookup below never crosses lactations.
all_lesions <- c("dd", "footrot", 'wld', "sole_ulcer", "injury",
                  "cork", "other", "hemorrhage",
                  "sole_fracture", "toe_ulcer", "thin",
                  "inf", "noninf", "toe", "axial", "lesion")
diseases <- c("inf", "noninf", "lesion")

lame_data <- events_formatted |>
  select(id_animal, lact_number, date_event, event,
         remark_letters1, protocols_letters1, location_event) |>
  filter(event %in% c("LAME", "TRIM", "FOOTRIM")) |>
  fxn_code_lesions(protocol_var = protocols_letters1, remark_var = remark_letters1) |>
  filter(!is.na(trimonly)) |>
  mutate(
    noninf = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 1,
      .default = noninf),
    other = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 0,
      .default = other)
  ) |>
  fxn_collapse_lesions(lesions = all_lesions) |>
  fxn_trim_vars(location_event, id_animal, date_var = date_event, trimonly_var = trimonly) |>
  fxn_dz_status(disease_cols = diseases, event_filter = trimonly == 0)

# Each NEDLAME cow's lameness-history label as of (at/before) her first
# NEDLAME date, restricted to the same lactation.
history_lookup <- lame_data |>
  select(id_animal, lact_number, date_event, status_lesion) |>
  filter(!is.na(status_lesion))

cohort <- cohort |>
  left_join(history_lookup, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(is.na(date_event) | date_event <= first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_max(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
  ungroup() |>
  mutate(history_group = if_else(is.na(status_lesion), "No Prior Lameness This Lactation", status_lesion)) |>
  select(-date_event, -status_lesion)

history_levels <- c("No Prior Lameness This Lactation", "New", "Repeat", "Chronic")
cohort <- cohort |>
  mutate(history_group = factor(history_group, levels = history_levels))
```

```{r}
#| label: milk-data

# data/milk/ is not wired into any existing pipeline script. Raw milk rows
# only carry a raw `id` (not the composite id_animal), so they're resolved
# to a specific animal-lactation by nearest-preceding date_fresh (see
# fxn_read_milk_folder.R for why a naive date_dry window isn't safe here).
milk_long <- fxn_read_milk_folder()
milk_resolved <- fxn_resolve_milk_to_lactation(milk_long, animal_lactations)
```

# Overview

`r n_tx` cows were enrolled in the **TX (Trim)** arm (`MNFRS = 1`, automatically trimmed after alert) and `r n_control` in the **Control** arm (`MNFRS = 2`, staff decide whether/when to act), based on `r nrow(first_nedlame)` cows that ever received a `NEDLAME` alert between `r format(min(first_nedlame$first_nedlame_date))` and `r format(max(first_nedlame$first_nedlame_date))`. This is a short (~5 week) history since NEDLAME is a newly-added event stream, which caps every time-window analysis below.

```{r}
#| label: overview-table

cohort |>
  count(tx_group, alert_type) |>
  pivot_wider(names_from = alert_type, values_from = n, values_fill = 0) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Cows per Treatment Group by Alert Type") |>
  tab_stubhead(label = "Treatment Group")
```

```{r}
#| label: dim-balance-table

cohort |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_dim = round(mean(dim_at_nedlame, na.rm = TRUE), 0),
    median_dim = round(median(dim_at_nedlame, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Lactation Stage at Enrollment",
    subtitle = "Days in milk (DIM) when first NEDLAME alert fired"
  ) |>
  cols_label(n = "N Cows", mean_dim = "Mean DIM", median_dim = "Median DIM")
```

# Q1: Milk Production Trend After First NEDLAME Alert

Cows are at very different lactation stages when flagged, so raw milk pounds aren't directly comparable across cows. Each cow's series is normalized to **% of her own pre-alert baseline** (mean production over the `r params$milk_baseline_days` days before the alert) before averaging by day and treatment group — this isolates the *shape* of the trend/recovery from each cow's absolute production level.

```{r}
#| label: q1-build

milk_cohort <- cohort |>
  inner_join(
    milk_resolved |> select(id_animal, lact_number, milk_date, dmlk1),
    by = c("id_animal", "lact_number"),
    relationship = "many-to-many"
  ) |>
  mutate(day_rel = as.numeric(milk_date - first_nedlame_date)) |>
  filter(day_rel >= -params$milk_window_pre, day_rel <= params$milk_window_post)

baseline <- milk_cohort |>
  filter(day_rel >= -params$milk_baseline_days, day_rel <= -1) |>
  group_by(id_animal, lact_number) |>
  summarize(baseline_dmlk1 = mean(dmlk1, na.rm = TRUE), .groups = "drop")

milk_norm <- milk_cohort |>
  inner_join(baseline, by = c("id_animal", "lact_number")) |>
  filter(baseline_dmlk1 >= 5) |> # drop the rare near-zero baseline (unreliable % denominator)
  mutate(pct_of_baseline = dmlk1 / baseline_dmlk1 * 100)

n_milk_cows <- n_distinct(milk_norm$id_animal)

trend <- milk_norm |>
  group_by(tx_group, day_rel) |>
  summarize(
    mean_pct = mean(pct_of_baseline, na.rm = TRUE),
    se_pct = sd(pct_of_baseline, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )
```

`r n_milk_cows` of the `r nrow(cohort)` cohort cows had at least one resolvable milk record in this window and a usable pre-alert baseline.

```{r}
#| label: q1-plot
#| fig-height: 5

ggplot(trend, aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    x = "Days Relative to First NEDLAME Alert",
    y = "Milk Production (% of Own Pre-Alert Baseline)",
    color = "Treatment Group", fill = "Treatment Group"
  )
```

## Lactation-Stage Balance

```{r}
#| label: q1-dim-check

cohort |>
  filter(id_animal %in% milk_norm$id_animal) |>
  ggplot(aes(x = tx_group, y = dim_at_nedlame, fill = tx_group)) +
  geom_boxplot() +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "Days in Milk at First NEDLAME Alert")
```

# Q2: Time to Next LAME/FOOTRIM/TRIM Event

For each cow, we look for the first `LAME`, `FOOTRIM`, or `TRIM` event strictly after her first NEDLAME alert, **within the same lactation** — censored at that lactation's dry-off date or the data pull date, whichever comes first. Because MNFRS=1 cows are automatically trimmed, "time to any trim" mechanically favors TX; a second version restricts to events where a lesion was actually identified, to separate "got trimmed on schedule" from "a problem was actually found."

```{r}
#| label: q2-build

next_events <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event, event) |>
  distinct()

lesion_lookup <- lame_data |> select(id_animal, lact_number, date_event, event, lesion)

fxn_time_to_next <- function(candidate_events) {
  cohort |>
    left_join(candidate_events, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
    filter(is.na(date_event) | (date_event > first_nedlame_date & date_event <= date_lact_end)) |>
    group_by(id_animal, lact_number) |>
    slice_min(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
    ungroup() |>
    mutate(
      event_occurred = as.integer(!is.na(date_event)),
      days_to_event = if_else(
        event_occurred == 1,
        as.numeric(date_event - first_nedlame_date),
        as.numeric(date_lact_end - first_nedlame_date)
      )
    ) |>
    filter(days_to_event >= 0)
}

q2_any <- fxn_time_to_next(next_events)

next_events_lesion <- next_events |>
  left_join(lesion_lookup, by = c("id_animal", "lact_number", "date_event", "event")) |>
  filter(!is.na(lesion) & lesion == 1)

q2_lesion <- fxn_time_to_next(next_events_lesion)
```

::: panel-tabset
## Time to Any Trim/LAME Event

```{r}
#| label: q2-any-plot
#| fig-height: 5

km_any <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_any)

km_any |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-any-table

km_any |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Trim/LAME Event")
```

## Time to a Trim/LAME Event *With a Lesion*

```{r}
#| label: q2-lesion-plot
#| fig-height: 5

km_lesion <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_lesion)

km_lesion |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Lesion Found") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-lesion-table

km_lesion |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Lesion Found")
```
:::

# Q3: % of Cows Independently Identified by Farm Staff

`CHKLAME` is the event that represents farm staff independently flagging a lame cow (distinct from the automated `NEDLAME` alert). A cow counts as **staff-identified** here if she has a `CHKLAME` event, in the same lactation, either (a) in the `r 7` days before her first NEDLAME alert, or (b) any time after it — excluding cows whose post-alert `CHKLAME` came at/after she'd already been trimmed, since a trim-triggered exam isn't independent detection.

```{r}
#| label: q3-build

chklame <- events_formatted |>
  filter(event == "CHKLAME") |>
  select(id_animal, lact_number, date_event) |>
  distinct()

trims <- events_formatted |>
  filter(event %in% c("FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event) |>
  distinct()

pre_week <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event >= first_nedlame_date - 7, date_event < first_nedlame_date) |>
  distinct(id_animal, lact_number)

first_post_chklame <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  rename(chklame_date = date_event)

first_post_trim <- cohort |>
  inner_join(trims, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(id_animal, lact_number, trim_date = date_event)

post_flag <- first_post_chklame |>
  left_join(first_post_trim, by = c("id_animal", "lact_number")) |>
  filter(is.na(trim_date) | chklame_date < trim_date) |>
  distinct(id_animal, lact_number)

staff_id <- cohort |>
  mutate(
    staff_identified = as.integer(
      paste(id_animal, lact_number) %in% paste(pre_week$id_animal, pre_week$lact_number) |
      paste(id_animal, lact_number) %in% paste(post_flag$id_animal, post_flag$lact_number)
    )
  )
```

```{r}
#| label: q3-table

q3_summary <- staff_id |>
  group_by(tx_group) |>
  summarize(n = n(), n_staff_id = sum(staff_identified), pct_staff_id = mean(staff_identified) * 100, .groups = "drop")

q3_test <- fisher.test(table(staff_id$tx_group, staff_id$staff_identified))

q3_summary |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_staff_id, decimals = 1) |>
  cols_label(n = "N Cows", n_staff_id = "Staff-Identified", pct_staff_id = "% Staff-Identified") |>
  tab_header(title = "Independent Staff Identification (CHKLAME)") |>
  tab_source_note(source_note = paste0("Fisher's exact test p = ", signif(q3_test$p.value, 3)))
```

# Q4: % of Post-June-1 Lesion Events NEDAP Didn't Catch First

Among `LAME` events with an identified lesion on or after `r params$history_after_date`, this checks whether that cow had a `NEDLAME` alert in the preceding `r params$q4_lookback_days` days (same lactation). A "miss" means staff/routine trimming found a lesion that NEDAP hadn't already flagged.

```{r}
#| label: q4-build

lame_lesion_recent <- lame_data |>
  filter(event == "LAME", lesion == 1, date_event >= ymd(params$history_after_date))

nedlame_all <- events_all |>
  filter(Event == "NEDLAME") |>
  select(id_animal, lact_number, ned_date = date_event) |>
  distinct()

q4 <- lame_lesion_recent |>
  left_join(nedlame_all, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(lookback_hit = !is.na(ned_date) & ned_date <= date_event & ned_date >= date_event - params$q4_lookback_days) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(caught_by_nedap = any(lookback_hit), .groups = "drop") |>
  left_join(cohort |> select(id_animal, lact_number, tx_group), by = c("id_animal", "lact_number")) |>
  mutate(tx_group = if_else(is.na(tx_group), "Never Alerted / No Group", tx_group))
```

```{r}
#| label: q4-table

q4 |>
  group_by(tx_group) |>
  summarize(n_lesion_events = n(), pct_missed = mean(!caught_by_nedap) * 100, .groups = "drop") |>
  bind_rows(
    q4 |> summarize(tx_group = "Overall", n_lesion_events = n(), pct_missed = mean(!caught_by_nedap) * 100)
  ) |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_missed, decimals = 1) |>
  cols_label(n_lesion_events = "N Lesion Events", pct_missed = "% Not Caught by NEDAP First") |>
  tab_header(title = paste("Lesion Events On/After", params$history_after_date))
```

# Q5: Influence of Lameness History

Each cow's lameness-history label (`New` / `Repeat` / `Chronic`, from the same chronicity classifier `report_explore_lame_new.qmd` uses, or "No Prior Lameness This Lactation" if she has no prior LAME/TRIM/FOOTRIM event this lactation) is taken as of her first NEDLAME alert. With only `r nrow(cohort)` cows split across 2 treatment arms x 4 history categories, cells here are thin (some under 20) — treat this section as descriptive/exploratory, not a confirmatory subgroup test.

```{r}
#| label: q5-balance-table

cohort |>
  count(tx_group, history_group) |>
  pivot_wider(names_from = history_group, values_from = n, values_fill = 0) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Lameness-History Distribution at Enrollment")
```

## Milk Trend by History Group

```{r}
#| label: q5-milk-plot
#| fig-height: 6

milk_norm |>
  group_by(tx_group, history_group, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~history_group) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

## Time to Next Trim/LAME Event by History Group

```{r}
#| label: q5-km-plot
#| fig-height: 6

# q2_any already carries history_group (inherited from cohort via
# fxn_time_to_next()), so no re-join is needed here.
q2_any_hist <- q2_any

# ggsurvfit only exposes a single combined `strata` column, so a KM plot
# faceted by history_group is built manually: fit one survfit2() per
# history level (skipping any level without both treatment arms
# represented) and stack the tidy() estimates before plotting.
fittable_levels <- history_levels |>
  keep(\(h) {
    d <- q2_any_hist |> filter(history_group == h)
    n_distinct(d$tx_group) == 2 && nrow(d) >= 10
  })

q5_km_data <- fittable_levels |>
  map(\(h) {
    d <- q2_any_hist |> filter(history_group == h)
    fit <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = d)
    broom::tidy(fit) |>
      mutate(
        tx_group = str_remove(strata, "^tx_group="),
        history_group = h,
        risk = 1 - estimate
      )
  }) |>
  list_rbind()

skipped_levels <- setdiff(as.character(history_levels), fittable_levels)

ggplot(q5_km_data, aes(x = time, y = risk, color = tx_group)) +
  geom_step(linewidth = 1) +
  facet_wrap(~history_group) +
  scale_color_viridis_d() +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

`r if (length(skipped_levels) > 0) paste0("*History group(s) omitted for insufficient data in one or both arms: ", paste(skipped_levels, collapse = ", "), ".*")`

## Staff-ID and NEDAP-Miss Rate by History Group

```{r}
#| label: q5-q3-q4-table

q3_by_hist <- staff_id |>
  group_by(tx_group, history_group) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop")

q3_by_hist |>
  gt(groupname_col = "tx_group") |>
  fmt_number(columns = pct_staff_id, decimals = 1) |>
  cols_label(history_group = "History Group", n = "N Cows", pct_staff_id = "% Staff-Identified") |>
  tab_header(title = "Q3 Staff-Identification Rate by Lameness History")
```

# Q6: Time to Culling After NEDAP Enrollment

Same culling-outcome pattern as `report_explore_lame_new.qmd`, anchored on the first NEDLAME date instead of a lesion-event date.

```{r}
#| label: q6-build

q6 <- cohort |>
  left_join(animals |> select(id_animal, date_sold, date_died, date_left), by = "id_animal") |>
  mutate(
    culled_ever = as.integer(!is.na(date_left)),
    days_to_censor_cull = if_else(
      culled_ever == 1,
      as.numeric(date_left - first_nedlame_date),
      as.numeric(date_max_pull - first_nedlame_date)
    )
  ) |>
  filter(days_to_censor_cull >= 0)

km_cull <- survfit2(Surv(days_to_censor_cull, culled_ever) ~ tx_group, data = q6)
```

```{r}
#| label: q6-plot
#| fig-height: 5

km_cull |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% Culled") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q6-table

km_cull |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% Culled After First NEDLAME Alert")
```

# Caveats

- **MNFRS coding was validated empirically, not from documentation**: cows with `MNFRS = 1` get trimmed within 14 days of their alert 81.5% of the time vs 42.9% for `MNFRS = 2`, consistent with `1 = TX (Trim)` / `2 = Control`. Confirm against the DairyComp/NEDLAME PDF if it defines this differently.
- **Short history**: `NEDLAME` only spans `r as.numeric(max(first_nedlame$first_nedlame_date) - min(first_nedlame$first_nedlame_date))` days of data, capping every time-to-event and milk-trend window above — results should be treated as an early read, not a mature-pilot conclusion.
- **Small sample for Q5**: 2 treatment arms x 4 history categories over `r nrow(cohort)` cows produces thin cells; the two arms are not perfectly balanced on prior lameness history at enrollment either (see the Q5 balance table), which should temper causal interpretation of any history-stratified differences.
- **Milk join resolution**: `r round(100 * n_milk_cows / nrow(cohort), 1)`% of cohort cows had a usable milk baseline in the analysis window; the remainder lacked resolvable milk records (e.g. dried off, sold, or outside the pulled milk-file date range).
