---
title: "NEDLAME Alert Pilot: TX (Trim) vs Control"
editor: source
format:
 html:
   embed-resources: true
   toc: true
   toc-location: left
   code-fold: true
execute:
    echo: false
    message: false
    warning: false
params:
  milk_window_pre: 14
  milk_window_post: 30
  milk_baseline_days: 7
  q4_lookback_days: 21
  history_after_date: "2026-06-01"
  recent_trim_suppression_days: 28
  post_trim_exclusion_days: 7
  analysis_end_date: "2026-08-27"
---

```{r}
#| label: setup

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  arrow,
  lubridate,
  gt,
  gtsummary,
  survival,
  survminer,
  ggsurvfit,
  scales,
  viridis
)

source(file.path('functions/fxn_read_milk_folder.R'))
source(file.path('functions/fxn_read_nedap_attentions.R'))

# GC's lame functions (fxn_collapse_lesions, fxn_trim_vars, fxn_dz_status) -
# fetched live from GitHub, same mechanism report_explore_lame_new.qmd uses.
# Two local overrides are sourced AFTER fxn_load_os_fxns() specifically so
# they win over the GitHub versions in the global environment:
#  - functions/fxn_code_lesions.R: the GitHub version's `trimonly` logic
#    only checks the *protocols* field for "NONE" - this farm's
#    routine/no-lesion trims are coded "NONE" in the *remark* field
#    instead, which the GitHub version doesn't check, so it silently
#    misclassifies every routine trim as having a lesion.
#  - functions/fxn_collapse_lesions.R: the GitHub version reassigns its
#    own `data` variable to an already-one-row-per-date subset before
#    computing the max() union across that date's lesion columns, so
#    when a cow has multiple event rows on the same date coding
#    DIFFERENT lesions, only whichever row survives the earlier
#    reduction keeps its lesion flags - the rest are silently dropped
#    (confirmed against a real cow with 4 same-day LAME rows: only 1 of
#    her 3 distinct lesion types survived under the GitHub version).
source(file.path('functions/fxn_load_os_fxns.R'))
fxn_load_os_fxns()
source(file.path('functions/fxn_code_lesions.R'))
source(file.path('functions/fxn_collapse_lesions.R'))

# Shared analysis functions: the cohort, exclusions and Q1-Q6 builds, so this
# report and the farm summary apply the same rules from one definition.
source(file.path('functions/fxn_nedlame_analysis.R'))

events_all <- read_parquet('data/intermediate_files/events_all_columns.parquet')
events_formatted <- read_parquet('data/intermediate_files/events_formatted.parquet')
animals <- read_parquet('data/intermediate_files/animals.parquet')
animal_lactations <- read_parquet('data/intermediate_files/animal_lactations.parquet')

date_max_pull <- max(events_all$date_event, na.rm = TRUE)
farm_name <- unique(events_all$location_event)

# NEDLAME go-live date, per the DairyComp setup: enrollment protocols were
# only turned on starting ~June 1 2026, so no NEDLAME alert can exist
# before this regardless of how "missed" a case looks.
nedlame_start_date <- min(events_all$date_event[events_all$Event == "NEDLAME"], na.rm = TRUE)
```

```{r}
#| label: cohort

# Cohort, exclusions and the analysis cut-off all live in
# functions/fxn_nedlame_analysis.R so every report applies the same rules.
invisible(list2env(fxn_build_cohort(events_all, animal_lactations, animals,
                                    date_max_pull, params), environment()))
```

```{r}
#| label: lesion-history-dataset

# Returns an augmented cohort (adds history_group), so it is reassigned here.
invisible(list2env(fxn_build_lame_history(events_formatted, cohort), environment()))
```

```{r}
#| label: milk-data

# data/milk/ is not wired into any existing pipeline script. Raw milk rows
# only carry a raw `id` (not the composite id_animal), so they're resolved
# to a specific animal-lactation by nearest-preceding date_fresh (see
# fxn_read_milk_folder.R for why a naive date_dry window isn't safe here).
# Only the window Q1 actually uses is read. The folder holds every milk day
# back to 2023, but the curves only span milk_window_pre days before an alert
# to milk_window_post days after, so reading the lot wastes ~26 s and then
# hands ~10x more rows than needed to the resolver, whose join costs far more
# than the read. Derived from the data rather than hard-coded, so it follows
# the alert window and the params automatically.
milk_from <- min(cohort$first_nedlame_date, na.rm = TRUE) - params$milk_window_pre
milk_to   <- max(cohort$first_nedlame_date, na.rm = TRUE) + params$milk_window_post

milk_long <- fxn_read_milk_folder(from = milk_from, to = milk_to)
milk_resolved <- fxn_resolve_milk_to_lactation(milk_long, animal_lactations)
```

```{r}
#| label: nedap-attentions-data

# Raw Nedap sensor export - what the SENSOR actually flagged, independent
# of whether that flag ever reached DairyComp as a NEDLAME event. Used in
# Q4 to split "NEDAP missed it" cases into true sensor misses vs pipeline
# losses (a real cowcard/Pulse cross-check first surfaced this pattern -
# see the plan doc's Round 8/10). This is a one-time export, not a live
# feed, so its coverage window is whatever was current when it was
# pulled - captured here explicitly rather than assumed, since it can
# lag behind the DairyComp/milk data read in above.
attentions_raw <- fxn_read_nedap_attentions()
attentions_resolved <- fxn_resolve_nedap_to_lactation(attentions_raw, animal_lactations)

nedap_coverage_start <- min(attentions_resolved$attention_date, na.rm = TRUE)
nedap_coverage_end <- max(attentions_resolved$attention_date, na.rm = TRUE)
```

```{r}
#| label: platform-enrollment-data


# Nedap runs two separate platform instances at this farm - "Activity"
# (used by farm staff daily) and "SmartSight" (mobility monitoring,
# generates NEDLAME) - kept separate at the farm's request to avoid
# confusion when checking the Nedap platform. A cow absent from
# SmartSight's animal list was never being monitored for lameness at
# all, regardless of what the raw attentions export's date coverage
# shows - this check is animal-level enrollment, not date-bound, so it
# can resolve cases the date-bound attentions export above can't (see
# the Platform Enrollment Check subsection in Q4).
cows_activity <- read_csv("data/cows_activity.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

cows_smartsight <- read_csv("data/cows_smartsight.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

# Renumber-aware enrollment - see fxn_build_enrolled_ids for why the current
# animal number alone is not enough.
invisible(list2env(fxn_build_enrolled_ids(events_all, cows_smartsight), environment()))
```

# Overview

## Setup in DairyComp

For this study the follow things were created in DairyComp to handle the file data from NEDAP daily.

- A command to clear all filein items before new data was brought in as a workaround to deal with historical flag issues

- New NEDAP lame event with 3 protocols to flag cows and criteria to create them

  - Low (score \<30 today, no alarm in the last 90 days, not trimmed in last 90 days, not a cow with history of upper leg lameness)

    - ENTER EC=70\\P215 FOR LFTYP=30102 LMFLG=1 LMSV=1-30 DSNLM=90-1 FTDAT\<-90 UPLAM\<\>1

  - Low Declining (no alarm in the last 7 days, so a cow already alerted is not re-flagged, and not trimmed in last 28 days)

    - ENTER EC=70\\P214 FOR LFTYP=30101 LMFLG=1 DSNLM=7-1 FTDAT\<-28

  - Strong Decline (no alarm in the last 7 days, not trimmed in last 28 days); the second OR statement lets a cow who already has a Low alert still raise a Strong Decline

    - ENTER EC=70\\P213 FOR LFTYP=30100 LMFLG=1 (DSNLM=7-1 FTDAT\<-28)(FTDAT\<-28 %73.NDLAME.101\>LOW)

- Cows were added to trim list with a command that uses an existing item (MNFRS) that randomizes cows by lactation group.

  - Enter FTDAT=1.1.1 for MNFRS=1 %.70NEDLAME.101=TODAY

Above commands were all run daily around 5 am after NEDAP file in occurred.

## Results

- `r nrow(first_nedlame)` cows received a `NEDLAME` alert between `r format(min(first_nedlame$first_nedlame_date))` and `r format(max(first_nedlame$first_nedlame_date))`

- The analysis stops at `r format(analysis_end)`, the last day the raw Nedap sensor export covers. DairyComp holds alerts past that date, but they cannot be checked against what the camera actually saw, so they are left out of every section rather than being partly checkable. Follow-up is not capped: a cow who entered on or before the cut-off keeps her later trims, culls and milk records.

- `r n_excluded_post_trim` were excluded as alerts occurred after a recent hoof trimming. These were still coded in DairyComp as the farm imports hoof trimming data a few days after the actual hoof trimming date. See @tbl-post-trim-exclusion-table.

This results in

- `r n_tx` cows enrolled in the **TX (Trim)** arm (`MNFRS = 1`, automatically trimmed after alert)

- `r n_control` in the **Control** arm (`MNFRS = 2`, staff decide whether/when to act),

**NOTE**: no cow is counted more than once in any table below. Where this report looks for "the" trim, LAME diagnosis, staff check, or NEDLAME alert for a cow, it either uses her **first** qualifying occurrence in the relevant window or asks only whether *at least one* occurred. Therefore a cow flagged or diagnosed repeatedly isn't counted more than once. When a single trimming date has more than one distinct lesion type coded (e.g. white line disease *and* digital dermatitis found at the same trimming), each distinct type is kept; the same lesion type coded more than once on one day only counts once.

### Description of Cows

The tables below describe the cows enrolled by treatment group

## By Alert Type

::: panel-tabset
```{r}
#| label: overview-table

cohort |>
  count(tx_group, alert_type) |>
  pivot_wider(names_from = alert_type, values_from = n, values_fill = 0) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Most cows were flagged as Low Declines") |>
  tab_stubhead(label = "Treatment Group")
```

## By Lactation Number

```{r}
#| label: lact-group-table

cohort |>
  mutate(lact_group_display = case_when(
    lact_number == 1 ~ "1",
    lact_number == 2 ~ "2",
    lact_number >= 3 ~ "3+",
    TRUE ~ NA_character_
  )) |>
  count(tx_group, lact_group_display) |>
  pivot_wider(names_from = lact_group_display, values_from = n, values_fill = 0) |>
  select(tx_group, `1`, `2`, `3+`) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Most cows enrolled were Lact 3+",
    ) |>
  tab_stubhead(label = "Treatment Group") |>
  cols_label(`3+` = "3+")
```

## By Lactation Stage (DIM)

```{r}
#| label: dim-balance-table

cohort |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_dim = round(mean(dim_at_nedlame, na.rm = TRUE), 0),
    median_dim = round(median(dim_at_nedlame, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Some skewness in DIM",
    ) |>
  cols_label(n = "N Cows", mean_dim = "Mean DIM", median_dim = "Median DIM")
```

```{r}
#| label: q1-dim-check
#| fig-height: 4

cohort |>
  ggplot(aes(x = tx_group, y = dim_at_nedlame, fill = tx_group)) +
  geom_violin(alpha = 0.5, color = NA, trim = FALSE) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.4, size = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "DIM at 1st NEDLAME Alert")
```
:::

### Details about excluded cows with alerts right 1-7 days after trimming

```{r}
#| label: tbl-post-trim-exclusion-table

cohort_excluded_post_trim |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_days_after_trim = round(mean(days_since_nearest_trim), 1),
    .groups = "drop"
  ) |>
  bind_rows(
    cohort_excluded_post_trim |> summarize(tx_group = "Overall", n = n(), mean_days_after_trim = round(mean(days_since_nearest_trim), 1))
  ) |>
  gt(rowname_col = "tx_group") |>
  cols_label(n = "N Excluded", mean_days_after_trim = "Mean Days After Trim") |>
  tab_header(title = "Systems flags cows recenlty trimmed quickly")
```

```{r}
#| label: post-trim-exclusion-plot
#| fig-height: 4

cohort_excluded_post_trim |>
  count(alert_type, days_since_nearest_trim) |>
  ggplot(aes(x = factor(days_since_nearest_trim), y = n, fill = alert_type)) +
  geom_col() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = "Days Between Trim and NEDLAME Alert", y = "N Cows Excluded", fill = "Alert Type") +
  theme(legend.position = "bottom")
```

```{r}
#| label: post-trim-followup-check

# The real evidence for the artifact reading: did anything ever come of
# these alerts? Computed rather than asserted - an earlier draft described
# this as "a separate check (not shown here)" and instead leaned on the
# share of decline-type alerts, which turns out to be just the cohort's
# base rate (see prose below) and therefore argues nothing either way.
excluded_followup <- cohort_excluded_post_trim |>
  select(id_animal, lact_number, first_nedlame_date, date_lact_end) |>
  left_join(
    events_formatted |>
      filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
      select(id_animal, lact_number, followup_date = date_event) |>
      distinct(),
    by = c("id_animal", "lact_number"), relationship = "many-to-many"
  ) |>
  mutate(is_followup = !is.na(followup_date) &
           followup_date > first_nedlame_date &
           followup_date <= date_lact_end) |>
  group_by(id_animal, lact_number) |>
  summarize(had_followup = any(is_followup), .groups = "drop")

n_excluded_no_followup <- sum(!excluded_followup$had_followup)
pct_excluded_no_followup <- round(100 * n_excluded_no_followup / nrow(excluded_followup), 0)

# Base-rate comparison, so the decline-share figure can be reported honestly.
pct_decline_excluded <- round(100 * mean(cohort_excluded_post_trim$alert_type %in% c("LOW DECLINE", "STRONG DECLINE")), 1)
pct_decline_cohort <- round(100 * mean(cohort$alert_type %in% c("LOW DECLINE", "STRONG DECLINE")), 1)
```

# Milk Production After 1st NEDLAME Alert

```{r}
#| label: q1-build

invisible(list2env(fxn_build_q1(cohort, milk_resolved, params), environment()))
```

The figures below compare milk production between the treatment groups. Cows were assigned to TX or Control at enrollment, so this is a like-for-like comparison rather than a descriptive one - and because the groups were randomised, seeing no difference here is itself evidence against a large effect, not just a lack of adjustment.

- `r n_milk_cows` of the `r nrow(cohort)` cohort cows had at least one milk record after their alert and a usable pre-alert baseline.

- The milk extract runs to `r format(milk_coverage_end)`, behind the event data. A cow needs `r params$milk_window_post` days of records after her alert to appear at the right-hand end of the curves, so the number of cows contributing falls across the x-axis: **`r n_milk_day30` cows sit behind the day-`r params$milk_window_post` point, not `r n_milk_cows`.** `r n_no_milk_at_all` cohort cows were alerted after the milk data ends and contribute nothing. The day-`r params$milk_window_post` figures therefore describe the earlier-alerting cows, and will change as milk data catches up.

Due to the fact that cows were at very different lactation stages when flagged milk production is not directly comparable across cows. Each cow's production is normalized to **% of her own pre-alert baseline** (mean production over the `r params$milk_baseline_days` days before the alert) before averaging by day and treatment group. This allow us to isolates the *shape* of the trend/recovery from each cow's absolute production level.

::: panel-tabset
## By Treatment Group

```{r}
#| label: q1-plot
#| fig-height: 5

ggplot(trend, aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    x = "Days Relative to First NEDLAME Alert",
    y = "Milk Production (% of Own Pre-Alert Baseline)",
    color = "Treatment Group", fill = "Treatment Group"
  )
```

## By Alert Type

Does the milk trend differ by the type of the Nedap alert?

```{r}
#| label: q1-alert-type-plot
#| fig-height: 5

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Trimming

## Time to next trimming after alert

These curves measure **protocol compliance rather than alert accuracy**. For each cow we took the first trimming event after her first NEDLAME alert, censoring at end of lactation or culling. In the TX arm the trim is caused by the alert by protocol, so time-to-trim is a measure of how faithfully the protocol was executed, not of whether the alert was correct. The Control curve is doing a different job: with no protocol acting on the alert, it shows when those cows were trimmed anyway, by staff decision or by the routine schedule - the split between those two is in @tbl-trim-driver. Whether the alerts were *right* is addressed by the lesion-found table below and by the miss analysis later on.

The figures below show the time to the next trimming event split by treatment group and for various lesion or alerts types.

- It is clear that most of the trim cows got trimmed within 7-10 days.

```{r}
#| label: q2-build

invisible(list2env(fxn_build_q2(cohort, events_formatted, lame_data), environment()))
```

::: panel-tabset
## Time to Any Trim/LAME Event

```{r}
#| label: q2-any-plot
#| fig-height: 5

km_any <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_any)

km_any |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-any-table

km_any |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Trim/LAME Event")
```

## Time to a Trim/LAME Event *With a Lesion*

```{r}
#| label: q2-lesion-plot
#| fig-height: 5

km_lesion <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_lesion)

km_lesion |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Lesion Found") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-lesion-table

km_lesion |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Lesion Found")
```

## By Alert Type

Decline cows mostly all get trimmed

```{r}
#| label: q2-alert-type-plot
#| fig-height: 6

alert_levels <- c("LOW", "LOW DECLINE", "STRONG DECLINE")

# ggsurvfit only exposes a single combined `strata` column, so a KM plot
# faceted by alert_type is built manually: fit one survfit2() per
# alert level (skipping any level without both treatment arms
# represented) and stack the tidy() estimates before plotting.
fittable_alert_levels <- alert_levels |>
  keep(\(a) {
    d <- q2_any |> filter(alert_type == a)
    n_distinct(d$tx_group) == 2 && nrow(d) >= 10
  })

q2_alert_type_km_data <- fittable_alert_levels |>
  map(\(a) {
    d <- q2_any |> filter(alert_type == a)
    fit <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = d)
    broom::tidy(fit) |>
      mutate(
        tx_group = str_remove(strata, "^tx_group="),
        alert_type = a,
        risk = 1 - estimate,
        # survival's CI bounds flip when converted to risk (1 - survival):
        # the upper bound on survival becomes the LOWER bound on risk.
        risk_low = 1 - conf.high,
        risk_high = 1 - conf.low
      )
  }) |>
  list_rbind() |>
  mutate(alert_type = factor(alert_type, levels = alert_levels))

skipped_alert_levels <- setdiff(alert_levels, fittable_alert_levels)

ggplot(q2_alert_type_km_data, aes(x = time, y = risk, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = risk_low, ymax = risk_high), alpha = 0.15, color = NA) +
  geom_step(linewidth = 1) +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event", color = "Treatment Group", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```

`r if (length(skipped_alert_levels) > 0) paste0("*Alert type(s) omitted for insufficient data in one or both arms: ", paste(skipped_alert_levels, collapse = ", "), ".*")`
:::

## What actually brought the Control cows in?

In the Control arm a cow is only trimmed if someone decides to act, so the trims above are a mixture of two different things: staff spotting her and putting her on the list, or her simply coming up on the routine schedule. A `CHKLAME` recorded between her alert and her trim marks the first.

```{r}
#| label: q2-chklame-build

invisible(list2env(fxn_build_control_trim_drivers(
  next_events, cohort, events_formatted, lame_data), environment()))
```

```{r}
#| label: tbl-trim-driver

control_trims |>
  group_by(Driver) |>
  summarize(`Cows` = n(),
            `% of trims` = round(100 * n() / n_ctl_trimmed, 1),
            `% finding a lesion` = round(100 * mean(found_lesion), 0),
            `Median days alert to trim` = median(days_to_trim),
            .groups = "drop") |>
  gt() |>
  tab_header(title = "Most Control trims were routine, and most of those found nothing") |>
  tab_source_note("Control cows only, counting each cow once at her first trim after the alert.")
```

- Of the `r n_ctl_trimmed` Control cows trimmed after their alert, **`r n_ctl_staff` (`r pct_ctl_staff`%) had a `CHKLAME` first**; the other `r n_ctl_routine` were trimmed with no staff check recorded.

- **The two groups are not doing the same job.** Staff-flagged trims found a lesion `r pct_lesion_staff`% of the time against `r pct_lesion_routine`% for routine ones, and happened at a median `r med_days_staff` days after the alert rather than `r med_days_routine`.

- So the Control curve above is mostly the routine trim schedule catching up with these cows, not staff responding to them. That is worth holding onto when comparing the arms: TX is being compared against a mixture, and only about a third of it is an active decision.

*This is a description of the trims that happened, not a third survival curve. Whether a cow gets a `CHKLAME` is decided after her alert, during follow-up, so splitting the earlier curves on it would compare groups defined by something that happens after the clock starts and would overstate the difference.*

## How many trims with no lesions?

The table below shows the number of cows who had a first NEDAP alert and where trimmed **but** did not have a lesions.

**This table's population is only the number of cows in the `r nrow(cohort)`-cow NEDLAME cohort that were trimmed.**

```{r}
#| label: q2-trimonly-summary

q2_first_event_detail <- q2_any |>
  filter(event_occurred == 1) |>
  left_join(lame_data |> select(id_animal, lact_number, date_event, event, trimonly),
            by = c("id_animal", "lact_number", "date_event"))

q2_trimonly_tab <- q2_first_event_detail |>
  group_by(tx_group) |>
  summarize(n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100, .groups = 'drop') |>
  bind_rows(
    q2_first_event_detail |> summarize(tx_group = 'Overall', n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100)
  )

# Title built from the cells beneath it. It used to be typed by hand and had
# already gone stale against a rebuild, quoting a range the table contradicted.
# level is 'TX (Trim)', not 'TX' - matching on 'TX' silently gave numeric(0)
tx_lvl <- setdiff(q2_trimonly_tab$tx_group, c('Control', 'Overall'))
stopifnot(length(tx_lvl) == 1)
pct_trimonly_tx  <- round(q2_trimonly_tab$pct_trimonly[q2_trimonly_tab$tx_group == tx_lvl], 1)
pct_trimonly_ctl <- round(q2_trimonly_tab$pct_trimonly[q2_trimonly_tab$tx_group == 'Control'], 1)
trimonly_title <- sprintf('%d-%d%% of Trims do not have a lesion',
                          floor(min(q2_trimonly_tab$pct_trimonly)), ceiling(max(q2_trimonly_tab$pct_trimonly)))

q2_trimonly_tab |>
  gt(rowname_col = 'tx_group') |>
  fmt_number(columns = pct_trimonly, decimals = 1) |>
  cols_label(n = 'N Cows with Trim', n_trimonly = 'No Lesion', pct_trimonly = '% Trim-Only') |>
  tab_header(title = trimonly_title)
```

# How does the Farm staff do?

The dairy uses a `CHKLAME` event to flag cows in the need of trimming. The tables below compares what % of cows with a NEDAP alert were flagged by farms staff \<=7 days of the NEDAP alert.

```{r}
#| label: q3-build

invisible(list2env(fxn_build_q3(cohort, events_formatted), environment()))
```

::: panel-tabset
## By Treatment Group

```{r}
#| label: q3-treatment-group-plot
#| fig-height: 4

staff_id |>
  group_by(tx_group) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = tx_group, y = pct_staff_id, fill = tx_group)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "% Staff-Identified (CHKLAME)")
```

## By Alert Type

Staff catches a bigger % of Low and Strong Decline alerts

```{r}
#| label: q3-alert-type-by-group-plot
#| fig-height: 4

staff_id |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = alert_type, y = pct_staff_id, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")),
            position = position_dodge(width = 0.7), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  labs(x = "Alert Type", y = "% Staff-Identified (CHKLAME)", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Influence of Lameness History

Lameness history can play an important role in camera performance. @fig-chronic shows the breakdown of new vs chronic cows that were enrolled. @fig-chronic-milk compares milk production across the lesion history types. Unlike the treatment groups, these are not randomised - a cow's history is a property she arrived with, along with her age, yield and everything else that goes with it - so read these curves as describing the groups, not as what a lesion history does to a cow.

Each cow's lameness-history label (`New` / `Chronic`) reflects her **lifetime record** as of her first NEDLAME alert. The `lame_data` goes back to 2021, and this deliberately isn't restricted to her current lactation, since chronicity for lameness should be considered a lifetime event.

```{r}
#| label: fig-chronic
#| fig-height: 4

cohort |>
  count(tx_group, history_group) |>
  group_by(tx_group) |>
  mutate(pct = n / sum(n) * 100) |>
  ungroup() |>
  mutate(history_group = fct_relabel(history_group, \(x) str_wrap(x, width = 14))) |>
  ggplot(aes(x = history_group, y = n, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(n, " (", round(pct, 0), "%)")),
            position = position_dodge(width = 0.7), vjust = -0.3, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal() +
  labs(x = NULL, y = "N Cows", fill = "Treatment Group",
       title = "Lameness-History Distribution at Enrollment") +
  theme(legend.position = "bottom",
        axis.text.x = element_text(lineheight = 0.9),
        plot.margin = margin(10, 15, 10, 15))
```

::: callout-note
`Repeat` is left out of the plot below but kept in the chart above. All `r sum(cohort$history_group == "Repeat")` `Repeat` cows got that label the same way: their most recent lesion before enrollment landed exactly one day after an earlier one. That is a timing artifact, not a clinical category.
:::

## Milk Trend by History Group

```{r}
#| label: fig-chronic-milk
#| fig-height: 6

milk_norm |>
  filter(history_group %in% c("New", "Chronic")) |>
  group_by(tx_group, history_group, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~history_group) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

# Time to Culling After NEDAP Enrollment

```{r}
#| label: q6-build

invisible(list2env(fxn_build_q6(cohort, date_max_pull), environment()))

# Quoted in the caveat below. The curves used to be summarised as 'no real
# impact', which claims a demonstrated absence of effect - with this few
# events the honest statement is that the study cannot tell either way.
n_culls <- sum(q6$culled_ever)
n_culls_tx  <- sum(q6$culled_ever[q6$tx_group != 'Control'])
n_culls_ctl <- sum(q6$culled_ever[q6$tx_group == 'Control'])
```

```{r}
#| label: q6-plot
#| fig-height: 5

km_cull |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% Culled") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q6-table

km_cull |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% Culled After First NEDLAME Alert")
```

Time to culling is compared between the treatment groups below. The curves sit close together, but only `r n_culls` cows were culled in total (`r n_culls_tx` TX, `r n_culls_ctl` Control). That is far too few to tell whether trimming off the alert changes culling - this is not evidence that it makes no difference, it is a question the study is too small and too short to answer.

# What cows were missed by the system?

This section looks at cows with lesion regardless of treatment group. Among cows with a `LAME`-diagnosed lesion on or after `r params$history_after_date`, this checks whether that cow had a `NEDLAME` alert in the preceding lookback window (same lactation). A "miss" means staff/routine trimming found a lesion that NEDAP hadn't already flagged. Each cow is counted once here, at her **first** qualifying lesion diagnosis on/after that date (consistent with using the first event throughout this report).

**Note** *Lame events that were in heifers, dried off cows or cows trimmed within 28 days of another trim were excluded as the system does not monitor those cows or they were automatically excluded*.

```{r}
#| label: q4-build

invisible(list2env(fxn_build_q4(
  cohort = cohort, all_nedlame_cows = all_nedlame_cows, lame_data = lame_data,
  events_formatted = events_formatted, events_all = events_all,
  animal_lactations = animal_lactations, enrolled_ids = enrolled_ids,
  nedlame_start_date = nedlame_start_date, params = params), environment()))
```

A total of `r nrow(q4)` cows had a lesion diagnosed after the project start date (`r params$history_after_date`). This excludes `r n_excluded_recent_trim` cows diagnosed within `r params$recent_trim_suppression_days` days of a prior trim, when NEDAP couldn't have alerted a cow due to the set up in DairyComp. Of the `r nrow(q4)` cows, `r n_truncated_lookback` had a lookback window shorter than the full `r params$q4_lookback_days` days (because they fell within `r params$q4_lookback_days` days of NEDLAME go-live).

Of those `r nrow(q4)` cows, `r sum(!q4$caught_by_nedap)` (`r round(100 * mean(!q4$caught_by_nedap), 1)`%) had no NEDAP alert entered in DairyComp in the `r params$q4_lookback_days` days before their lesion was diagnosed.

## Alert Check: True Misses vs. Pipeline Losses

The figure above can only answer "did a `NEDLAME` event land in DairyComp in time". However, it can't tell a true alert miss apart from a case where the cow was flagged but the flag never reached DairyComp. The impact of this was checked with raw flag data from Nedap for the following time period (`r format(nedap_coverage_start)` to `r format(nedap_coverage_end)`).

The figure below compares among lesions what the reasons for not being flagged were. Some pipeline issues exist (likely low alerts with a score \>30) but in the never alerted section the majority are ones that where never flagged.

```{r}
#| label: q4-sensor-build

# Coverage is capped at nedap_coverage_end - the raw export is a one-time
# pull, not a live feed. The lookback window only looks BACKWARD from
# date_event, so a case is fully checkable as long as date_event itself
# is on/before the coverage end date; cases after that are flagged as
# "Not Checked" rather than silently miscounted as true misses just
# because we don't have sensor data that far out yet.
q4_sensor <- q4_group |>
  mutate(in_coverage = date_event <= nedap_coverage_end)

# Same pre-filter-before-join pattern used throughout this report:
# restrict candidates to the checkable subset FIRST, then join - a
# left_join on the full q4_sensor table followed by a filter would risk
# silently dropping the "Not Checked" rows instead of correctly keeping
# and labeling them.
sensor_hits <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> select(id_animal, lact_number, attention_date),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(sensor_hit = !is.na(attention_date) &
           attention_date < date_event &
           attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(sensor_fired_in_window = any(sensor_hit), .groups = "drop")

q4_sensor <- q4_sensor |>
  left_join(sensor_hits, by = c("id_animal", "lact_number", "date_event")) |>
  mutate(
    # The strings in case_when() and in levels= must match EXACTLY or the
    # non-matching ones become NA and vanish from every count and plot.
    detection_group_3way = factor(case_when(
      !in_coverage ~ "No Data (Outside Coverage)",
      caught_by_nedap ~ "Flagged, Caught in Time",
      sensor_fired_in_window ~ "Flagged, but no alert in time",
      TRUE ~ "Never Flagged (True Miss)"
    ), levels = c(
      "Flagged, Caught in Time", "Flagged, but no alert in time",
      "Never Flagged (True Miss)", "No Data (Outside Coverage)"
    ))
  )

misses_in_coverage <- q4_sensor |> filter(in_coverage, !caught_by_nedap)
n_pipeline_loss <- sum(misses_in_coverage$detection_group_3way == "Flagged, but no alert in time")
n_true_miss <- sum(misses_in_coverage$detection_group_3way == "Never Flagged (True Miss)")
pct_pipeline_loss <- round(100 * n_pipeline_loss / nrow(misses_in_coverage), 1)
n_outside_coverage <- sum(!q4_sensor$in_coverage)
# Misses specifically that fall outside coverage - the figure that belongs
# next to nrow(misses_in_coverage). n_outside_coverage counts ALL lesion
# cases outside coverage (caught ones included), so quoting it alongside a
# miss-denominated number mixes two bases.
n_misses_outside_coverage <- sum(!q4_sensor$in_coverage & !q4_sensor$caught_by_nedap)

# Sensitivity on is_completed. The raw export marks some attention episodes
# as not completed, and the loader does not filter them out, so an incomplete
# episode currently counts as the camera having flagged her. Rather than make
# that choice silently, re-run the same window test using only completed
# episodes and report how many pipeline losses depend on the incomplete ones.
sensor_hits_completed <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> filter(is_completed) |> select(id_animal, lact_number, attention_date),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(sensor_hit = !is.na(attention_date) &
           attention_date < date_event &
           attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(fired_completed = any(sensor_hit), .groups = "drop")

n_pipeline_loss_completed_only <- misses_in_coverage |>
  filter(detection_group_3way == "Flagged, but no alert in time") |>
  left_join(sensor_hits_completed, by = c("id_animal", "lact_number", "date_event")) |>
  summarize(n = sum(coalesce(fired_completed, FALSE))) |>
  pull(n)
n_pipeline_loss_incomplete_only <- n_pipeline_loss - n_pipeline_loss_completed_only
n_attentions_incomplete <- sum(!attentions_resolved$is_completed, na.rm = TRUE)
```

```{r}
#| label: q4-sensor-plot
#| fig-height: 4

q4_sensor |>
  count(tx_group, detection_group_3way) |>
  ggplot(aes(x = tx_group, y = n, fill = detection_group_3way)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_viridis_d() +
  guides(fill = guide_legend(nrow = 2)) +
  theme_minimal() +
  labs(x = NULL, y = "% of Lesion Cases", fill = "Alert Status") +
  theme(legend.position = "bottom")
```

Further investigation of the missed cows showed:

- Of the `r nrow(misses_in_coverage)` DC-recorded misses that fall within the raw sensor's coverage window, `r n_pipeline_loss` (`r pct_pipeline_loss`%) actually had an alert within the lookback window.

  - The flag never got into DairyComp as a `NEDLAME` event (potential pipeline issue)

- The remaining `r n_true_miss` had no flag at all in that window, a true miss.

`r if (n_misses_outside_coverage > 0) paste0("- The other ", n_misses_outside_coverage, " misses were diagnosed after ", format(nedap_coverage_end), ", which the current export can't see. These are left unclassified rather than assumed.") else paste0("Every miss here is checkable against the raw sensor feed: the analysis stops at ", format(analysis_end), ", the last day the Nedap export covers.")`

One judgement call sits behind these numbers. `r n_attentions_incomplete` of the resolved flags are marked not completed in the raw export, and they are counted here as the camera having flagged the cow. `r n_pipeline_loss_incomplete_only` of the `r n_pipeline_loss` pipeline losses rest only on such a flag; counting completed flags alone would give `r n_pipeline_loss_completed_only` instead. The choice matters at the margin, so it is stated rather than assumed.

## What Lesions are being missed?

The figure below shows the lesion types found at these first-diagnosis cases, broken out by NEDAP detection status:

- **Flagged, Caught in Time** — an alert landed in DairyComp inside the window before the lesion.
- **Flagged, but no alert in time** — the camera did flag her, but no alert reached DairyComp in time.
- **Never Flagged (True Miss)** — the camera never saw a problem at all.

Where a cow had two different lesions on the same day, each type is counted. The same type coded twice on one day counts once.

From the figure below it is clear that DD and other are the main categories being missed. Other on this farm is typically Corkscrews.

```{r}
#| label: q4-lesion-types

invisible(list2env(fxn_build_q4_groups(q4_sensor, lame_lesion_recent), environment()))

ggplot(q4_lesion_types, aes(x = lesion_type, y = pct, fill = detection_group_3way)) +
  geom_col(width = 0.7) +
  geom_text(data = lesion_totals, inherit.aes = FALSE,
            aes(x = lesion_type, y = 101, label = paste0("n=", total_cases)),
            hjust = 0, size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 100, 25), labels = function(x) paste0(x, "%"),
                     limits = c(0, 118), expand = expansion(mult = c(0, 0))) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases with that lesion", fill = NULL) +
  theme(legend.position = "bottom")
```

## Which leg is being missed?

The same cases split by which foot the lesion was found on. A cow can have lesions on more than one foot, so the bars add to more than `r nrow(q4_group)` cases.

```{r}
#| label: fig-missed-by-leg
#| fig-height: 4

leg_names <- c(LF = "Left Front", RF = "Right Front", LR = "Left Rear", RR = "Right Rear")

# locate_lesion packs one or more two-character foot codes into one string
# ("RRLR" = right rear + left rear, "RRLRRFLF" = all four). It must be parsed
# in 2-character chunks, NOT with substring matching: str_detect("LRRF", "RR")
# is TRUE, but "LRRF" is LR + RF, so substring matching would invent a
# right-rear lesion that was never recorded. Checked every value in the data -
# all are even length and every chunk is one of LF/RF/LR/RR.
leg_long <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM"), !is.na(locate_lesion), locate_lesion != "") |>
  select(id_animal, lact_number, date_event, locate_lesion) |>
  # locate_lesion packs one or more two-character foot codes into one string
  # ("RRLR" = right rear + left rear, "RRLRRFLF" = all four), so it must be
  # read in 2-character chunks, never by substring matching: str_detect("LRRF",
  # "RR") is TRUE but "LRRF" is LR + RF.
  #
  # It is built in functions/fxn_location.R, which used to search the whole
  # remark and so invented a phantom left front on XNLF* codes - XNL
  # (treatment) + F (foot rot) + RF (leg) is ONE foot, but "LF" spans the
  # treatment's trailing L and the lesion's F. Fixed at source, so nothing is
  # needed here.
  mutate(leg = str_extract_all(locate_lesion, ".{2}")) |>
  unnest(leg) |>
  filter(leg %in% names(leg_names)) |>
  distinct(id_animal, lact_number, date_event, leg) |>
  # Map to readable names HERE, once, so everything downstream sees
  # "Left Rear" rather than "LR". Leaving the raw codes in place and
  # renaming later silently broke a front/rear split that tested
  # str_detect(leg, "Rear") - it matched nothing and called every foot Front.
  mutate(leg = unname(leg_names[leg]))

q4_legs <- leg_long |>
  inner_join(q4_sensor |> select(id_animal, lact_number, date_event, detection_group_3way),
             by = c("id_animal", "lact_number", "date_event")) |>
  count(detection_group_3way, leg) |>
  # Same reasoning as the lesion-type figure above: a share within each foot,
  # so the four feet are comparable regardless of how many lesions each carries.
  group_by(leg) |>
  mutate(total_cases = sum(n), pct = 100 * n / total_cases) |>
  ungroup() |>
  mutate(leg = fct_reorder(leg, total_cases))

leg_totals <- q4_legs |> distinct(leg, total_cases)

# leg is a factor now (ordered for the plot), and str_detect() rejects factors
n_legs_rear <- sum(q4_legs$n[str_detect(as.character(q4_legs$leg), "Rear")])
n_legs_front <- sum(q4_legs$n[str_detect(as.character(q4_legs$leg), "Front")])

ggplot(q4_legs, aes(x = leg, y = pct, fill = detection_group_3way)) +
  geom_col(width = 0.7) +
  geom_text(data = leg_totals, inherit.aes = FALSE,
            aes(x = leg, y = 101, label = paste0("n=", total_cases)),
            hjust = 0, size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 100, 25), labels = function(x) paste0(x, "%"),
                     limits = c(0, 118), expand = expansion(mult = c(0, 0))) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases on that foot", fill = NULL) +
  theme(legend.position = "bottom")
```

Rear feet account for `r n_legs_rear` of the `r n_legs_rear + n_legs_front` foot-level lesions, front feet `r n_legs_front`. That is where lesions occur, not where the camera fails - which segment of a stacked bar is tallest says nothing about detection, because every segment scales with how common lesions are on that foot. Whether the camera misses front feet more than hind feet is asked directly below.

## What is the camera actually missing?

```{r}
#| label: true-miss-build

checkable <- q4_sensor |>
  filter(in_coverage) |>
  mutate(camera = if_else(detection_group_3way == "Never Flagged (True Miss)",
                          "Camera never flagged", "Camera flagged her"))

n_cam_miss <- sum(checkable$camera == "Camera never flagged")
n_cam_flag <- sum(checkable$camera == "Camera flagged her")

lesion_by_camera <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(checkable |> select(id_animal, lact_number, date_event, camera),
             by = c("id_animal", "lact_number", "date_event")) |>
  pivot_longer(all_of(lesion_type_cols), names_to = "lesion_type", values_to = "has_lesion") |>
  filter(has_lesion == 1) |>
  count(camera, lesion_type) |>
  mutate(pct = if_else(camera == "Camera never flagged", 100 * n / n_cam_miss, 100 * n / n_cam_flag))

pct_dd_miss  <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "dd"  & lesion_by_camera$camera == "Camera never flagged"], 0)
pct_dd_flag  <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "dd"  & lesion_by_camera$camera == "Camera flagged her"], 0)
pct_wld_miss <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "wld" & lesion_by_camera$camera == "Camera never flagged"], 0)
pct_wld_flag <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "wld" & lesion_by_camera$camera == "Camera flagged her"], 0)

# Front/rear share within each group, computed here rather than inline so the
# prose stays readable.
leg_end <- leg_long |>
  mutate(end = if_else(str_detect(leg, "Rear"), "Rear", "Front")) |>
  distinct(id_animal, lact_number, date_event, end)

leg_by_camera <- leg_end |>
  inner_join(checkable |> select(id_animal, lact_number, date_event, camera),
             by = c("id_animal", "lact_number", "date_event")) |>
  count(camera, end) |>
  group_by(camera) |>
  mutate(pct = 100 * n / sum(n)) |>
  ungroup()

pct_rear_miss <- round(leg_by_camera$pct[leg_by_camera$camera == "Camera never flagged" & leg_by_camera$end == "Rear"], 0)
pct_rear_flag <- round(leg_by_camera$pct[leg_by_camera$camera == "Camera flagged her"   & leg_by_camera$end == "Rear"], 0)
```


The `r n_cam_miss` cases the camera never flagged are compared below against the `r n_cam_flag` it did flag. Both sit inside the sensor coverage window, so both have raw flag data behind them. Percentages are within each group, and a cow can have more than one lesion type and more than one affected foot, so they add to more than 100%.

```{r}
#| label: fig-true-miss-lesion
#| fig-height: 4

lesion_by_camera |>
  ggplot(aes(x = fct_reorder(lesion_type, pct, .fun = max), y = pct, fill = camera)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases in that group", fill = NULL) +
  theme(legend.position = "bottom")
```

- **Digital dermatitis is over-represented**: `r pct_dd_miss`% of the cases the camera never flagged, against `r pct_dd_flag`% of the ones it did flag.
- **White line is the opposite**: `r pct_wld_miss`% of misses vs `r pct_wld_flag`% of flagged cases. The camera picks white line up well.


- **Front versus hind is not where the difference is.** Rear feet are `r pct_rear_miss`% of the missed cases against `r pct_rear_flag`% of the ones the camera did flag. Hind-foot lesions dominate both groups in much the same proportion, so there is no sign the camera is specifically blind to front feet.

So the difference is in the lesion, not the limb. That is not really a surprise: locomotion scoring is known to pick digital dermatitis up poorly, and the camera is scoring locomotion, so a lesion that produces little gait change is exactly the lesion it will tend to miss. What this analysis adds is the size of the gap on this farm, and that it is specific rather than general - white line, which does change how a cow walks, is caught well.

Since we are running a separate SmartSight instance, I wanted to make sure all cows were visible on it. A cow absent from SmartSight is not being watched for lameness at all, so a lesion found on her cannot be a camera miss.

```{r}
#| label: platform-enrollment-build

q4_sensor <- q4_sensor |>
  left_join(events_all |> select(id_animal, id) |> distinct(), by = "id_animal") |>
  mutate(
    # alias-aware, matching the Q4 eligibility rule exactly - a cow enrolled
    # under a previous animal number is enrolled (see enrolled_ids in setup)
    not_in_smartsight = !(id %in% enrolled_ids),
    # Checked, not assumed: the point of this section is that a cow can be
    # current on the Activity platform (which staff use daily) while absent
    # from SmartSight. Earlier drafts asserted "even though she's on
    # Activity" in prose without computing it - now computed so the claim
    # can't drift away from the data.
    in_activity = id %in% cows_activity$id
  )

# Unenrolled cows are excluded upstream in Q4, so this should now be empty.
# Kept as an assertion rather than deleted: if it ever stops being zero, the
# exclusion has silently stopped working.
n_residual_not_in_smartsight <- sum(!q4_sensor$caught_by_nedap & q4_sensor$not_in_smartsight)
n_total_misses <- sum(!q4_sensor$caught_by_nedap)

# Herd-level enrollment check. The misses above are only the unenrolled cows
# who also happened to get a lesion, so they understate the gap. This asks the
# direct question instead: of every first-lactation cow that freshened in the
# study window, what share is on the SmartSight list at all? Compared against
# the Activity list, which is the control - if both dropped together it would
# point at the export; if only SmartSight drops it points at enrollment.
ss_ids  <- unique(cows_smartsight$id)
act_ids <- unique(cows_activity$id)

l1_fresh <- animal_lactations |>
  filter(lact_number == 1, !is.na(date_fresh), date_fresh >= ymd(params$history_after_date) - 150,
         date_fresh <= analysis_end) |>
  distinct(id_animal, id, date_fresh) |>
  mutate(id = as.character(id),
         on_smartsight = id %in% ss_ids,
         on_activity   = id %in% act_ids,
         fresh_month   = format(date_fresh, "%Y-%m"))

enroll_by_month <- l1_fresh |>
  group_by(`Freshened` = fresh_month) |>
  summarize(`Cows` = n(),
            `% on SmartSight` = round(100 * mean(on_smartsight), 1),
            `% on Activity`   = round(100 * mean(on_activity), 1),
            .groups = "drop")

# the hard break: the most recent freshening among enrolled first-lactation cows
last_enrolled_fresh <- max(l1_fresh$date_fresh[l1_fresh$on_smartsight], na.rm = TRUE)
l1_after  <- l1_fresh |> filter(date_fresh > last_enrolled_fresh)
l1_before <- l1_fresh |> filter(date_fresh <= last_enrolled_fresh)
n_l1_after_unenrolled <- sum(!l1_after$on_smartsight)
pct_l1_before_enrolled <- round(100 * mean(l1_before$on_smartsight), 1)
n_l1_after_on_activity <- sum(l1_after$on_activity)

```


`r n_excluded_not_enrolled` lesion cases were excluded on this basis, and they are no longer counted as misses anywhere above.

Those are only the unenrolled cows who also happened to get a lesion, so they understate the problem. Asking the question directly - of every first-lactation cow that freshened in the window, how many are on each platform:

```{r}
#| label: tbl-enrollment-by-month

enroll_by_month |>
  gt() |>
  tab_header(title = "SmartSight enrollment of first-lactation cows stops; Activity does not") |>
  tab_source_note("Activity is the control here. Both lists were pulled on the same day, so a drop in one and not the other cannot be an export problem.")
```

- The most recent freshening among first-lactation cows who *are* on SmartSight is **`r format(last_enrolled_fresh)`**. Before that date `r pct_l1_before_enrolled`% were enrolled; after it, **`r n_l1_after_unenrolled` first-lactation cows are absent**, and `r n_l1_after_on_activity` of them are on Activity.

- Both lists were pulled on the same day, well after the cut-off, so this is not a stale-export artifact. A drop in SmartSight with Activity unaffected points at enrollment having stopped, not at data transfer.

- **Animal numbers get reassigned, and that had to be handled before this check could be trusted.** DairyComp logs an ID change as an `XID` event holding the old number, and `r nrow(id_aliases)` such changes appear in this extract. A cow renumbered after she was enrolled looks absent under her current number while being present under her old one - cow 23143, renumbered from 14347, is enrolled all along as 14347. Enrollment is therefore checked against a cow's current *and* prior numbers; `r n_rescued_by_alias` animals are only found that way.

## Why Did No Alert Arrive in Time?

Two rules limit which cows can become a `Low` alert: a score under 31, and no history of upper-leg lameness (`Leg-AboveFoot` in the trim `protocols`).

Score-at-flag isn't available in this data, so only the injury/`Leg-AboveFoot` half of this can be checked. It is shown in the figure below.

```{r}
#| label: pipeline-loss-injury-build

# All four by-design rules - FTDAT, DSNLM, injury history, and the late
# arrival - live in functions/fxn_nedlame_analysis.R so the standalone
# pipeline-loss workbook reaches the identical verdict. They used to be
# written here and again in that script, and had already drifted: DSNLM was
# implemented in the workbook and not in the report.
invisible(list2env(
  fxn_build_pipeline_by_design(q4_sensor, attentions_resolved, events_formatted,
                               events_all, lame_data),
  environment()))
```

```{r}
#| label: pipeline-loss-injury-plot
#| fig-height: 4

# `reason` assigns ONE cause per case, so these bars sum to the total and
# match the counts in the prose. They used to be computed separately here.
pipeline_loss_cases |>
  count(category = reason) |>
  ggplot(aes(x = fct_reorder(category, n), y = n, fill = category)) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "N Cases") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

**Every cow in this section was still missed.** She is counted in the `r sum(!q4_sensor$caught_by_nedap)` above either way — the question here is only whether the camera's signal went astray on the way to DairyComp, or whether the system was correctly declining to raise an alert.

Four rules can explain this without anything having gone missing. Each case is counted once, under the first rule that applies, so the figures below add to `r nrow(pipeline_loss_cases)`. Of those cases:

- `r n_only_ftdat` had every flag land inside its own `FTDAT` window (90 days for `Low`, 28 for the declines). DairyComp was right to skip them.

- `r n_only_injury` were `LOW`-only flags on cows with upper-leg lameness history, which blocks the `Low` route.

- `r n_only_dsnlm` had every flag suppressed by `DSNLM`: she had already been alerted inside the window (90 days on the `Low` route, 7 on the declines), so a repeat alert is deliberately withheld to keep the cowcard clean.

- `r n_only_late` did get an alert into DairyComp — the day after the lesion, too late to act on. The daily batch runs at ~5am and stamps an alert with the day it was loaded, so an attention after ~5am on day D is logged as D+1, landing just outside a lookback that only looks backward. Two readings fit and the data cannot separate them: either that flag arrived late, or the camera made a fresh observation of her post-trim gait and *that* is what came through. Under the second reading the original flag really was lost. Either way she was missed, and either way DairyComp was receiving something from the camera around that date — which is why these are not counted among the confirmed losses below.

That accounts for **`r n_pipeline_loss_by_design` of the `r nrow(pipeline_loss_cases)`**. The remaining `r n_pipeline_loss_unexplained` are not all failures either, and it matters to separate them:

- **`r n_pipeline_loss_unknowable` cannot be judged at all.** Only a `LOW` flag fired, and the `Low` route requires a score of 1-30. Score-at-flag is not in this data, so a cow scoring 31-69 who was correctly declined looks exactly like a lost flag. These are unresolved, not unexplained.

- **`r n_pipeline_loss_genuine` are genuine pipeline losses.** A decline route fired, and the decline routes carry no score gate — so there is no rule left that would explain the silence. These are the cases worth taking to Nedap, and cow-level detail for them is in `pipeline_losses.xlsx`.

So the integration problem is **`r n_pipeline_loss_genuine` confirmed cases, not `r nrow(pipeline_loss_cases)`** — with an upper bound of `r n_pipeline_loss_unexplained` if every unjudgeable case turned out to be a failure, which is unlikely.

The by-design count is still a floor: the trim lookup only covers the current lactation while the real `FTDAT` probably looks back further, so some of the `r n_pipeline_loss_unknowable` will be correct suppressions too.

## Of the Misses, Did Farm Staff Catch Them Anyway?

```{r}
#| label: q4-staff-catch-build

invisible(list2env(fxn_build_staff_catch(q4, chklame), environment()))
```

Of the `r nrow(staff_catch)` cows whose lesion had no `NEDLAME` alert recorded in DairyComp beforehand (they were trimmed on the routine round or because staff picked them out),

- `r sum(staff_catch$staff_flagged)` (`r round(100 * mean(staff_catch$staff_flagged), 1)`%) were flagged by the farm staff in the 7 days before diagnosis.

## Summarizing the Misses

Each check above narrows a different group of cows, so the percentages don't share a denominator. The table below ties them together. Indented rows are splits of the row above, and the "% of" column always refers to the base named beside it.

```{r}
#| label: q4-reconciliation

n_caught_total <- sum(q4$caught_by_nedap)
n_staff_caught <- sum(staff_catch$staff_flagged)

q4_recon <- tibble::tribble(
  ~step, ~n, ~base, ~base_n,
  "Qualifying lesion cases (after eligibility exclusions)", nrow(q4), "—", NA_integer_,
  " Caught by NEDAP in time", n_caught_total, "qualifying cases", nrow(q4),
  " Not caught — \"misses\"", n_total_misses, "qualifying cases", nrow(q4),
  "  Within raw sensor coverage (checkable)", nrow(misses_in_coverage), "misses", n_total_misses,
  "  Outside coverage (not checkable yet)", n_misses_outside_coverage, "misses", n_total_misses,
  "   Flag fired anyway — candidate pipeline loss", n_pipeline_loss, "checkable misses", nrow(misses_in_coverage),
  "   Never flagged — true miss", n_true_miss, "checkable misses", nrow(misses_in_coverage),
  "    Of candidates: blocked by its own FTDAT window", n_pipeline_loss_ftdat_blocked, "candidates", n_pipeline_loss,
  "    Of candidates: LOW-only + injury/upper-leg history", n_pipeline_loss_injury_explained, "candidates", n_pipeline_loss,
  "    Of candidates: by design overall (either rule)", n_pipeline_loss_by_design, "candidates", n_pipeline_loss,
  "    Of candidates: cannot tell (LOW only, no score)", n_pipeline_loss_unknowable, "candidates", n_pipeline_loss,
  "    Of candidates: GENUINE pipeline loss", n_pipeline_loss_genuine, "candidates", n_pipeline_loss
) |>
  mutate(pct = if_else(is.na(base_n), NA_real_, 100 * n / base_n))

q4_recon |>
  select(step, n, pct, base) |>
  gt() |>
  fmt_number(columns = pct, decimals = 1) |>
  sub_missing(columns = pct, missing_text = "") |>
  cols_label(step = "", n = "N", pct = "% of base", base = "…of what") |>
  cols_align(align = "left", columns = step) |>
  tab_header(
    title = "Q4 Reconciliation: How the Populations Nest",
    subtitle = "Indented rows split the row above them"
  ) |>
  tab_source_note(source_note = paste0(
    "Cows never enrolled in SmartSight are excluded before this chain starts (",
    n_excluded_not_enrolled, " lesion cases), on the same basis as dry cows and unfreshened heifers - ",
    "so they are not a row here and the residual inside the chain is ", n_residual_not_in_smartsight, ". ",
    "One check does cut ACROSS the chain rather than extending it: ", n_staff_caught, " of the ", n_total_misses,
    " misses were independently flagged by staff, which can apply to any row."))
```

**The bottom line of this:**

- of the `r n_total_misses` cases DairyComp records as missed,

- only `r nrow(misses_in_coverage)` can currently be checked against the raw sensor feed at all.

  Within that checkable slice,

- `r n_true_miss` are genuine missing alerts

- `r n_pipeline_loss` are cases where the camera did flag her

  - `r n_pipeline_loss_by_design` were correct by-design suppressions (`FTDAT` window or injury/upper-leg rule), not integration failures,

  - `r n_pipeline_loss_unknowable` cannot be judged, because only a `LOW` flag fired and score-at-flag is not in the data,

  - leaving **`r n_pipeline_loss_genuine` genuine pipeline losses** — `r round(100 * n_pipeline_loss_genuine / nrow(misses_in_coverage), 1)`% of the checkable misses. That, not the headline miss rate, is the size of the integration problem.
