---
title: "NEDLAME Alert Pilot: TX (Trim) vs Control"
editor: source
format:
 html:
   embed-resources: true
   toc: true
   toc-location: left
   code-fold: true
execute:
    echo: false
    message: false
    warning: false
params:
  milk_window_pre: 14
  milk_window_post: 30
  milk_baseline_days: 7
  q4_lookback_days: 21
  history_after_date: "2026-06-01"
  recent_trim_suppression_days: 28
  post_trim_exclusion_days: 7
  analysis_end_date: "2026-08-27"
---

```{r}
#| label: setup

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  arrow,
  lubridate,
  gt,
  gtsummary,
  survival,
  survminer,
  ggsurvfit,
  scales,
  viridis,
  patchwork
)

source(file.path('functions/fxn_read_milk_folder.R'))
source(file.path('functions/fxn_read_nedap_attentions.R'))

# GC's lame functions (fxn_collapse_lesions, fxn_trim_vars, fxn_dz_status) -
# fetched live from GitHub, same mechanism report_explore_lame_new.qmd uses.
# Two local overrides are sourced AFTER fxn_load_os_fxns() specifically so
# they win over the GitHub versions in the global environment:
#  - functions/fxn_code_lesions.R: the GitHub version's `trimonly` logic
#    only checks the *protocols* field for "NONE" - this farm's
#    routine/no-lesion trims are coded "NONE" in the *remark* field
#    instead, which the GitHub version doesn't check, so it silently
#    misclassifies every routine trim as having a lesion.
#  - functions/fxn_collapse_lesions.R: the GitHub version reassigns its
#    own `data` variable to an already-one-row-per-date subset before
#    computing the max() union across that date's lesion columns, so
#    when a cow has multiple event rows on the same date coding
#    DIFFERENT lesions, only whichever row survives the earlier
#    reduction keeps its lesion flags - the rest are silently dropped
#    (confirmed against a real cow with 4 same-day LAME rows: only 1 of
#    her 3 distinct lesion types survived under the GitHub version).
source(file.path('functions/fxn_load_os_fxns.R'))
fxn_load_os_fxns()
source(file.path('functions/fxn_code_lesions.R'))
source(file.path('functions/fxn_collapse_lesions.R'))

# Shared analysis functions: the cohort, exclusions and Q1-Q6 builds, so this
# report and the farm summary apply the same rules from one definition.
source(file.path('functions/fxn_nedlame_analysis.R'))

events_all <- read_parquet('data/intermediate_files/events_all_columns.parquet')
events_formatted <- read_parquet('data/intermediate_files/events_formatted.parquet')
animals <- read_parquet('data/intermediate_files/animals.parquet')
animal_lactations <- read_parquet('data/intermediate_files/animal_lactations.parquet')

date_max_pull <- max(events_all$date_event, na.rm = TRUE)
farm_name <- unique(events_all$location_event)

# NEDLAME go-live date, per the DairyComp setup: enrollment protocols were
# only turned on starting ~June 1 2026, so no NEDLAME alert can exist
# before this regardless of how "missed" a case looks.
nedlame_start_date <- min(events_all$date_event[events_all$Event == "NEDLAME"], na.rm = TRUE)
```

```{r}
#| label: cohort

# Cohort, exclusions and the analysis cut-off all live in
# functions/fxn_nedlame_analysis.R so every report applies the same rules.
invisible(list2env(fxn_build_cohort(events_all, animal_lactations, animals,
                                    date_max_pull, params), environment()))
```

```{r}
#| label: lesion-history-dataset

# Returns an augmented cohort (adds history_group), so it is reassigned here.
invisible(list2env(fxn_build_lame_history(events_formatted, cohort), environment()))
```

```{r}
#| label: milk-data

# data/milk/ is not wired into any existing pipeline script. Raw milk rows
# only carry a raw `id` (not the composite id_animal), so they're resolved
# to a specific animal-lactation by nearest-preceding date_fresh (see
# fxn_read_milk_folder.R for why a naive date_dry window isn't safe here).
# Only the window Q1 actually uses is read. The folder holds every milk day
# back to 2023, but the curves only span milk_window_pre days before an alert
# to milk_window_post days after, so reading the lot wastes ~26 s and then
# hands ~10x more rows than needed to the resolver, whose join costs far more
# than the read. Derived from the data rather than hard-coded, so it follows
# the alert window and the params automatically.
milk_from <- min(cohort$first_nedlame_date, na.rm = TRUE) - params$milk_window_pre
milk_to   <- max(cohort$first_nedlame_date, na.rm = TRUE) + params$milk_window_post

milk_long <- fxn_read_milk_folder(from = milk_from, to = milk_to)
milk_resolved <- fxn_resolve_milk_to_lactation(milk_long, animal_lactations)
```

```{r}
#| label: nedap-attentions-data

# Raw Nedap sensor export - what the SENSOR actually flagged, independent
# of whether that flag ever reached DairyComp as a NEDLAME event. Used in
# Q4 to split "NEDAP missed it" cases into true sensor misses vs pipeline
# losses (a real cowcard/Pulse cross-check first surfaced this pattern -
# see the plan doc's Round 8/10). This is a one-time export, not a live
# feed, so its coverage window is whatever was current when it was
# pulled - captured here explicitly rather than assumed, since it can
# lag behind the DairyComp/milk data read in above.
attentions_raw <- fxn_read_nedap_attentions()
attentions_resolved <- fxn_resolve_nedap_to_lactation(attentions_raw, animal_lactations)

nedap_coverage_start <- min(attentions_resolved$attention_date, na.rm = TRUE)
nedap_coverage_end <- max(attentions_resolved$attention_date, na.rm = TRUE)
```

```{r}
#| label: platform-enrollment-data


# Nedap runs two separate platform instances at this farm - "Activity"
# (used by farm staff daily) and "SmartSight" (mobility monitoring,
# generates NEDLAME) - kept separate at the farm's request to avoid
# confusion when checking the Nedap platform. A cow absent from
# SmartSight's animal list was never being monitored for lameness at
# all, regardless of what the raw attentions export's date coverage
# shows - this check is animal-level enrollment, not date-bound, so it
# can resolve cases the date-bound attentions export above can't (see
# the Platform Enrollment Check subsection in Q4).
cows_activity <- read_csv("data/cows_activity.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

cows_smartsight <- read_csv("data/cows_smartsight.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

# Renumber-aware enrollment - see fxn_build_enrolled_ids for why the current
# animal number alone is not enough.
invisible(list2env(fxn_build_enrolled_ids(events_all, cows_smartsight), environment()))
```

```{=html}
<!-- Everything below the bottom line is presentation. The computation is
     hoisted here because the bottom line quotes figures that sections much
     further down produce, and Quarto evaluates inline code in document
     order. Hoisting the existing chunks rather than recomputing avoids a
     second construction of detection_group_3way, which is the divergence
     this project has been bitten by repeatedly. -->
```

```{r}
#| label: q2-build

invisible(list2env(fxn_build_q2(cohort, events_formatted, lame_data), environment()))
```

```{r}
#| label: q4-build

invisible(list2env(fxn_build_q4(
  cohort = cohort, all_nedlame_cows = all_nedlame_cows, lame_data = lame_data,
  events_formatted = events_formatted, events_all = events_all,
  animal_lactations = animal_lactations, enrolled_ids = enrolled_ids,
  nedlame_start_date = nedlame_start_date, params = params), environment()))
```

```{r}
#| label: q4-sensor-build

# Coverage is capped at nedap_coverage_end - the raw export is a one-time
# pull, not a live feed. The lookback window only looks BACKWARD from
# date_event, so a case is fully checkable as long as date_event itself
# is on/before the coverage end date; cases after that are flagged as
# "Not Checked" rather than silently miscounted as true misses just
# because we don't have sensor data that far out yet.
q4_sensor <- q4_group |>
  mutate(in_coverage = date_event <= nedap_coverage_end)

# Same pre-filter-before-join pattern used throughout this report:
# restrict candidates to the checkable subset FIRST, then join - a
# left_join on the full q4_sensor table followed by a filter would risk
# silently dropping the "Not Checked" rows instead of correctly keeping
# and labeling them.
sensor_hits <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> select(id_animal, lact_number, attention_date),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(sensor_hit = !is.na(attention_date) &
           attention_date < date_event &
           attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(sensor_fired_in_window = any(sensor_hit), .groups = "drop")

q4_sensor <- q4_sensor |>
  left_join(sensor_hits, by = c("id_animal", "lact_number", "date_event")) |>
  mutate(
    # The strings in case_when() and in levels= must match EXACTLY or the
    # non-matching ones become NA and vanish from every count and plot.
    detection_group_3way = factor(case_when(
      !in_coverage ~ "No Data (Outside Coverage)",
      caught_by_nedap ~ "Flagged, Caught in Time",
      warned_not_acted ~ "Alerted earlier, no action taken",
      sensor_fired_in_window ~ "Flagged, but no alert in time",
      TRUE ~ "Never Flagged (True Miss)"
    ), levels = c(
      "Flagged, Caught in Time", "Alerted earlier, no action taken", "Flagged, but no alert in time",
      "Never Flagged (True Miss)", "No Data (Outside Coverage)"
    ))
  )

misses_in_coverage <- q4_sensor |> filter(in_coverage, is_miss)
n_pipeline_loss <- sum(misses_in_coverage$detection_group_3way == "Flagged, but no alert in time")
n_true_miss <- sum(misses_in_coverage$detection_group_3way == "Never Flagged (True Miss)")
# These two split the misses on a string. Rename a level anywhere and the
# comparison quietly matches nothing: the counts go to zero, the page still
# renders, and the numbers just come out wrong. Happened in the farm report.
stopifnot(n_pipeline_loss + n_true_miss == nrow(misses_in_coverage))
pct_pipeline_loss <- round(100 * n_pipeline_loss / nrow(misses_in_coverage), 1)
n_outside_coverage <- sum(!q4_sensor$in_coverage)
# Misses specifically that fall outside coverage - the figure that belongs
# next to nrow(misses_in_coverage). n_outside_coverage counts ALL lesion
# cases outside coverage (caught ones included), so quoting it alongside a
# miss-denominated number mixes two bases.
n_misses_outside_coverage <- sum(!q4_sensor$in_coverage & q4_sensor$is_miss)

# Sensitivity on is_completed. The raw export marks some attention episodes
# as not completed, and the loader does not filter them out, so an incomplete
# episode currently counts as the camera having flagged her. Rather than make
# that choice silently, re-run the same window test using only completed
# episodes and report how many pipeline losses depend on the incomplete ones.
sensor_hits_completed <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> filter(is_completed) |> select(id_animal, lact_number, attention_date),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(sensor_hit = !is.na(attention_date) &
           attention_date < date_event &
           attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(fired_completed = any(sensor_hit), .groups = "drop")

n_pipeline_loss_completed_only <- misses_in_coverage |>
  filter(detection_group_3way == "Flagged, but no alert in time") |>
  left_join(sensor_hits_completed, by = c("id_animal", "lact_number", "date_event")) |>
  summarize(n = sum(coalesce(fired_completed, FALSE))) |>
  pull(n)
n_pipeline_loss_incomplete_only <- n_pipeline_loss - n_pipeline_loss_completed_only
n_attentions_incomplete <- sum(!attentions_resolved$is_completed, na.rm = TRUE)
```

```{r}
#| label: pipeline-loss-injury-build

# All four by-design rules - FTDAT, DSNLM, injury history, and the late
# arrival - live in functions/fxn_nedlame_analysis.R so the standalone
# pipeline-loss workbook reaches the identical verdict. They used to be
# written here and again in that script, and had already drifted: DSNLM was
# implemented in the workbook and not in the report.
invisible(list2env(
  fxn_build_pipeline_by_design(q4_sensor, attentions_resolved, events_formatted,
                               events_all, lame_data),
  environment()))
```

```{r}
#| label: q6-build

invisible(list2env(fxn_build_q6(cohort, date_max_pull), environment()))

# Quoted in the caveat below. The curves used to be summarised as 'no real
# impact', which claims a demonstrated absence of effect - with this few
# events the honest statement is that the study cannot tell either way.
n_culls <- sum(q6$culled_ever)
n_culls_tx  <- sum(q6$culled_ever[q6$tx_group != 'Control'])
n_culls_ctl <- sum(q6$culled_ever[q6$tx_group == 'Control'])
```

```{r}
#| label: lesion-yield-build

invisible(list2env(fxn_build_lesion_yield(q2_lesion, q2_any), environment()))
```

# **The Bottom Line**

- The camera never flagged **`r n_true_miss` of the `r nrow(q4)` cows** found with a lesion.

  - Digital dermatitis is over-represented among them and white line under-represented

    - This is not surprising considering the model being trained on locomotion scoring.

- A further **`r n_pipeline_loss` were flagged but no alert reached DairyComp in time.**

  - Most of that is the enrollment rules working as designed.

    - **`r n_pipeline_loss_genuine` are confirmed losses** between the camera and DairyComp.

    - This excludes the first lactation animals that stopped being entered in the SmartSight instance in August.

- Trimming off the alert (Trim Treatment (TX) group) found a lesion in **`r n_extra_lesion_cows` more cows** than the control approach.

  - `r pct_trimmed_tx`% of TX cows were trimmed against `r pct_trimmed_ctl`% of Control. Since the groups were randmomized the TX lesions were we can infer that that they were found earlier.

- **No clear difference in milk or culling over 30 days.** The milk curves slightly favour TX, but cow-to-cow variation dwarfs the group difference, and with `r n_culls` culls in total the study cannot answer the culling question either way.

- More longer term follow up is required to measure other outcomes such as repeat lesions and longer-run production.

- The pilot gave valuable insight for the proposed follow up study and details of that proposal will follow.

Each of these is developed below, with the checks behind it.

# Setup in DairyComp

For this study the follow things were created in DairyComp to handle the file data from NEDAP daily.

- A command to clear all filein items before new data was brought in as a workaround to deal with historical flag issues in the NEDAP files

- New NEDAP lame event with 3 protocols to flag cows and criteria to create them

  - Low (score \<30 today, no alarm in the last 90 days, not trimmed in last 90 days, not a cow with history of upper leg lameness)

    - ENTER EC=70\\P215 FOR LFTYP=30102 LMFLG=1 LMSV=1-30 DSNLM=90-1 FTDAT\<-90 UPLAM\<\>1
    - The exclusions were added to prevent excessive alarms for cows with know locomotion issues.

  - Low Declining (no alarm in the last 7 days, so a cow already alerted is not re-flagged, and not trimmed in last 28 days)

    - ENTER EC=70\\P214 FOR LFTYP=30101 LMFLG=1 DSNLM=7-1 FTDAT\<-28

  - Strong Decline (no alarm in the last 7 days, not trimmed in last 28 days); the second OR statement lets a cow who already has a Low alert still raise a Strong Decline

    - ENTER EC=70\\P213 FOR LFTYP=30100 LMFLG=1 (DSNLM=7-1 FTDAT\<-28)(FTDAT\<-28 %73.NDLAME.101\>LOW)

- Cows were added to trim list with a command that uses an existing item (MNFRS) that randomizes cows by lactation group.

  - Enter FTDAT=1.1.1 for MNFRS=1 %.70NEDLAME.101=TODAY

Above commands were all run daily around 5 am after NEDAP file in occurred.

# Results

- `r nrow(first_nedlame)` cows received a `NEDLAME` alert between `r format(min(first_nedlame$first_nedlame_date))` and `r format(max(first_nedlame$first_nedlame_date))`

- The analysis stops at `r format(analysis_end)`, the last day the raw Nedap sensor export covers.

  - Follow-up extends past this data to allow following of trims, culls and milk records.

- `r n_excluded_post_trim` alerts on cows were excluded as alerts occurred after a recent hoof trimming. These were still coded in DairyComp as the farm imports hoof trimming data a few days after the actual hoof trimming date.

  - See @tbl-post-trim-exclusion-table. for details on these cows

With above exclusions and conditions it resulted in

- `r n_tx` cows enrolled in the **TX (Trim)** arm (`MNFRS = 1`, automatically added to trim list after an alert)

- `r n_control` in the **Control** arm (`MNFRS = 2`, staff or routine trimming would have added her to the trim list.

**NOTE**: *No cow is counted more than once in any table below. Where this report looks for "the" trim, LAME diagnosis, staff check, or NEDLAME alert for a cow, it either uses her **first** qualifying occurrence in the relevant window or asks only whether at least one occurred. Therefore a cow flagged or diagnosed repeatedly isn't counted more than once. When a single trimming date has more than one distinct lesion type coded (e.g. white line disease and digital dermatitis found at the same trimming), each distinct type is kept; the same lesion type coded more than once on one day only counts once.*

## Description of Cows

The tables below describe the cows enrolled by treatment group to ensure groups were similar.

- The high number of low declines is not surprising considering we move the low threshold to allow more scores to get into DairyComp.

::: panel-tabset
## By Alert Type

```{r}
#| label: tbl-overview-table

cohort |>
  count(tx_group, alert_type) |>
  pivot_wider(names_from = alert_type, values_from = n, values_fill = 0) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Most cows were flagged as Low Declines") |>
  tab_stubhead(label = "Treatment Group")
```

## By Lactation Number

```{r}
#| label: tbl-lact-group-table

cohort |>
  mutate(lact_group_display = case_when(
    lact_number == 1 ~ "1",
    lact_number == 2 ~ "2",
    lact_number >= 3 ~ "3+",
    TRUE ~ NA_character_
  )) |>
  count(tx_group, lact_group_display) |>
  pivot_wider(names_from = lact_group_display, values_from = n, values_fill = 0) |>
  select(tx_group, `1`, `2`, `3+`) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Lact 3+ was the largest group enrolled",
    ) |>
  tab_stubhead(label = "Treatment Group") |>
  cols_label(`3+` = "3+")
```

## By Lactation Stage (DIM)

```{r}
#| label: tbl-dim-balance-table

cohort |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_dim = round(mean(dim_at_nedlame, na.rm = TRUE), 0),
    median_dim = round(median(dim_at_nedlame, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Some skewness in DIM",
    ) |>
  cols_label(n = "N Cows", mean_dim = "Mean DIM", median_dim = "Median DIM")
```

```{r}
#| label: fig-q1-dim-check
#| fig-height: 4

cohort |>
  ggplot(aes(x = tx_group, y = dim_at_nedlame, fill = tx_group)) +
  geom_violin(alpha = 0.5, color = NA, trim = FALSE) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.4, size = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "DIM at 1st NEDLAME Alert")
```
:::

### Details about excluded cows with alerts 1-7 days after trimming

```{r}
#| label: tbl-post-trim-exclusion-table

cohort_excluded_post_trim |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_days_after_trim = round(mean(days_since_nearest_trim), 1),
    .groups = "drop"
  ) |>
  bind_rows(
    cohort_excluded_post_trim |> summarize(tx_group = "Overall", n = n(), mean_days_after_trim = round(mean(days_since_nearest_trim), 1))
  ) |>
  gt(rowname_col = "tx_group") |>
  cols_label(n = "N Excluded", mean_days_after_trim = "Mean Days After Trim") |>
  tab_header(title = "System flags recently trimmed cows quickly")
```

```{r}
#| label: fig-post-trim-exclusion-plot
#| fig-height: 4

cohort_excluded_post_trim |>
  count(alert_type, days_since_nearest_trim) |>
  ggplot(aes(x = factor(days_since_nearest_trim), y = n, fill = alert_type)) +
  geom_col() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = "Days Between Trim and NEDLAME Alert", y = "N Cows Excluded", fill = "Alert Type") +
  theme(legend.position = "bottom")
```

```{r}
#| label: post-trim-followup-check

# The real evidence for the artifact reading: did anything ever come of
# these alerts? Computed rather than asserted - an earlier draft described
# this as "a separate check (not shown here)" and instead leaned on the
# share of decline-type alerts, which turns out to be just the cohort's
# base rate (see prose below) and therefore argues nothing either way.
excluded_followup <- cohort_excluded_post_trim |>
  select(id_animal, lact_number, first_nedlame_date, date_lact_end) |>
  left_join(
    events_formatted |>
      filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
      select(id_animal, lact_number, followup_date = date_event) |>
      distinct(),
    by = c("id_animal", "lact_number"), relationship = "many-to-many"
  ) |>
  mutate(is_followup = !is.na(followup_date) &
           followup_date > first_nedlame_date &
           followup_date <= date_lact_end) |>
  group_by(id_animal, lact_number) |>
  summarize(had_followup = any(is_followup), .groups = "drop")

n_excluded_no_followup <- sum(!excluded_followup$had_followup)
pct_excluded_no_followup <- round(100 * n_excluded_no_followup / nrow(excluded_followup), 0)

# Base-rate comparison, so the decline-share figure can be reported honestly.
pct_decline_excluded <- round(100 * mean(cohort_excluded_post_trim$alert_type %in% c("LOW DECLINE", "STRONG DECLINE")), 1)
pct_decline_cohort <- round(100 * mean(cohort$alert_type %in% c("LOW DECLINE", "STRONG DECLINE")), 1)
```

## Trimming Data

### Time to next trimming after alert

These time to event curves measure **protocol compliance**.

- For each cow we took the first trim after her first NEDLAME alert, censored at end of lactation or culling.

  - It is clear that most of the TX cows got trimmed within 7-10 days as designed.
  - The difference between the time to event with a lesion (@fig-q2-lesion-plot) illustrates how the TX group finds more cows with lesions.

In TX the trim is caused by the alert, so the curve shows how well the protocol was followed.

The Control curve does a different job: nothing there acts on the alert, so it shows when those cows were trimmed anyway due to either staff flagging them or a routine scheduled trimming.

- A breakdown of the staff-versus-routine split is in @tbl-trim-driver.

::: panel-tabset
## Time to Any Trim/LAME Event

```{r}
#| label: fig-q2-any-plot
#| fig-height: 5

km_any <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_any)

km_any |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: tbl-q2-any-table

km_any |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Trim/LAME Event")
```

## Time to a Trim/LAME Event *With a Lesion*

```{r}
#| label: fig-q2-lesion-plot
#| fig-height: 5

km_lesion <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_lesion)

km_lesion |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Lesion Found") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: tbl-q2-lesion-table

km_lesion |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Lesion Found")
```

## By Alert Type

Decline cows mostly all get trimmed

```{r}
#| label: fig-q2-alert-type-plot
#| fig-height: 6

alert_levels <- c("LOW", "LOW DECLINE", "STRONG DECLINE")

# ggsurvfit only exposes a single combined `strata` column, so a KM plot
# faceted by alert_type is built manually: fit one survfit2() per
# alert level (skipping any level without both treatment arms
# represented) and stack the tidy() estimates before plotting.
fittable_alert_levels <- alert_levels |>
  keep(\(a) {
    d <- q2_any |> filter(alert_type == a)
    n_distinct(d$tx_group) == 2 && nrow(d) >= 10
  })

q2_alert_type_km_data <- fittable_alert_levels |>
  map(\(a) {
    d <- q2_any |> filter(alert_type == a)
    fit <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = d)
    broom::tidy(fit) |>
      mutate(
        tx_group = str_remove(strata, "^tx_group="),
        alert_type = a,
        risk = 1 - estimate,
        # survival's CI bounds flip when converted to risk (1 - survival):
        # the upper bound on survival becomes the LOWER bound on risk.
        risk_low = 1 - conf.high,
        risk_high = 1 - conf.low
      )
  }) |>
  list_rbind() |>
  mutate(alert_type = factor(alert_type, levels = alert_levels))

skipped_alert_levels <- setdiff(alert_levels, fittable_alert_levels)

ggplot(q2_alert_type_km_data, aes(x = time, y = risk, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = risk_low, ymax = risk_high), alpha = 0.15, color = NA) +
  geom_step(linewidth = 1) +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event",
       color = "Treatment Group", fill = "Treatment Group",
       caption = "Shaded band: 95% confidence interval from the Kaplan-Meier fit.") +
  theme(legend.position = "bottom")
```

`r if (length(skipped_alert_levels) > 0) paste0("*Alert type(s) omitted for insufficient data in one or both arms: ", paste(skipped_alert_levels, collapse = ", "), ".*")`
:::

### How many trims with no lesions?

The table below shows the number of cows who had a first NEDAP alert and were trimmed **but** did not have a lesion.

**Note:** *This table's population is only the number of cows in the `r nrow(cohort)`-cow NEDLAME cohort that were trimmed.*

```{r}
#| label: tbl-q2-trimonly-summary

q2_first_event_detail <- q2_any |>
  filter(event_occurred == 1) |>
  left_join(lame_data |> select(id_animal, lact_number, date_event, event, trimonly),
            by = c("id_animal", "lact_number", "date_event"))

q2_trimonly_tab <- q2_first_event_detail |>
  group_by(tx_group) |>
  summarize(n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100, .groups = 'drop') |>
  bind_rows(
    q2_first_event_detail |> summarize(tx_group = 'Overall', n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100)
  )

# Title built from the cells beneath it. It used to be typed by hand and had
# already gone stale against a rebuild, quoting a range the table contradicted.
# level is 'TX (Trim)', not 'TX' - matching on 'TX' silently gave numeric(0)
tx_lvl <- setdiff(q2_trimonly_tab$tx_group, c('Control', 'Overall'))
stopifnot(length(tx_lvl) == 1)
pct_trimonly_tx  <- round(q2_trimonly_tab$pct_trimonly[q2_trimonly_tab$tx_group == tx_lvl], 1)
pct_trimonly_ctl <- round(q2_trimonly_tab$pct_trimonly[q2_trimonly_tab$tx_group == 'Control'], 1)
trimonly_title <- sprintf('%d-%d%% of Trims do not have a lesion',
                          floor(min(q2_trimonly_tab$pct_trimonly)), ceiling(max(q2_trimonly_tab$pct_trimonly)))

q2_trimonly_tab |>
  gt(rowname_col = 'tx_group') |>
  fmt_number(columns = pct_trimonly, decimals = 1) |>
  cols_label(n = 'N Cows with Trim', n_trimonly = 'No Lesion', pct_trimonly = '% Trim-Only') |>
  tab_header(title = trimonly_title)
```

## What actually brought the Control cows in?

```{r}
#| label: q2-chklame-build

invisible(list2env(fxn_build_control_trim_drivers(
  next_events, cohort, events_formatted, lame_data), environment()))
```

In the Control arm a cow is trimmed if she is flagged by farms staff (CHKLAME) or she is due for her routine trimming. The table below shows that the majority of trims in the Control group were due to routine trimming as only **`r n_ctl_staff` (`r pct_ctl_staff`%)** had a `CHKLAME` event.

- The rate of lesions an time of the CHKLAME vs routine trimming differs as staff-flagged trims found a lesion `r pct_lesion_staff`% of the time against `r pct_lesion_routine`% for routine ones, and staff flag trims happened at a median `r med_days_staff` days after the alert rather than `r med_days_routine` for routine trimmings.

```{r}
#| label: tbl-trim-driver

control_trims |>
  group_by(Driver) |>
  summarize(`Cows` = n(),
            `% of trims` = round(100 * n() / n_ctl_trimmed, 1),
            `% finding a lesion` = round(100 * mean(found_lesion), 0),
            `Median days alert to trim` = median(days_to_trim),
            .groups = "drop") |>
  gt() |>
  tab_header(title = "Most Control trims were routine, and most of those found nothing") |>
  tab_source_note("Control cows only, counting each cow once at her first trim after the alert.")
```

## How many extra lesions did the Trim Treatment group find?

From the time to lesion curve above ( @fig-q2-lesion-plot) it is clear that TX found more cows.

- A lesion was found in `r n_lesion_tx` of the `r n_cows_tx` TX cows (`r pct_lesion_arm_tx`%)

- In Control cows that was `r n_lesion_ctl` of `r n_cows_ctl` (`r pct_lesion_arm_ctl`%).

Rate-adjusted for the unequal arms, that is **`r n_extra_lesion_cows` more cows with a lesion identified**.

**Note:** *the denominator: this is the **full arm** (`r n_cows_tx` and `r n_cows_ctl`), not the trimmed subset the no-lesion table uses (`r n_trimmed_tx` and `r n_trimmed_ctl`).*

### How does the Farm staff do?

The dairy uses a `CHKLAME` event to flag cows in the need of trimming.

The tables below compares what % of cows with a NEDAP alert were flagged by farms staff \<=7 days of the NEDAP alert in both groups.

Overall farm staff flags about 20% of cows with an alert. M

- This differs by alert type with more cows with low scores or strong declines being flagged.

```{r}
#| label: q3-build

invisible(list2env(fxn_build_q3(cohort, events_formatted), environment()))
```

::: panel-tabset
## By Treatment Group

```{r}
#| label: fig-q3-treatment-group-plot
#| fig-height: 4

staff_id |>
  group_by(tx_group) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = tx_group, y = pct_staff_id, fill = tx_group)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "% Staff-Identified (CHKLAME)")
```

## By Alert Type

Staff catches a bigger % of Low and Strong Decline alerts

```{r}
#| label: fig-q3-alert-type-by-group-plot
#| fig-height: 4

staff_id |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = alert_type, y = pct_staff_id, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")),
            position = position_dodge(width = 0.7), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  labs(x = "Alert Type", y = "% Staff-Identified (CHKLAME)", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Milk Production After 1st NEDLAME Alert

```{r}
#| label: q1-build

invisible(list2env(fxn_build_q1(cohort, milk_resolved, params), environment()))
```

The figures below compare milk production between the treatment groups. Cows were assigned to TX or Control at enrollment.

- `r n_milk_cows` of the `r nrow(cohort)` cohort cows had at least one milk record after their alert and a usable pre-alert baseline.

- Fewer cows sit behind the right-hand end of these curves than the left

  - a cow needs `r params$milk_window_post` days of milk records after her alert to appear at day `r params$milk_window_post`, and only `r n_milk_day30` have that so far.

    - This means the numbers can still change as the remaining cows get to 30 days after alert.

Overall numerically there is an advantage to the trimmed group in milk production after an alert but there is a lot of variation (the shaded part around the solid line) so this is not something to put a lot of trust in.

**NOTE:** *Due to the fact that cows were at very different lactation stages when flagged milk production is not directly comparable across cows. Each cow's production is normalized to **% of her own pre-alert baseline** (mean production over the `r params$milk_baseline_days` days before the alert) before averaging by day and treatment group. This allow us to isolates the shape of the trend/recovery from each cow's absolute production level.*

::: panel-tabset
## By Treatment Group

```{r}
#| label: fig-q1-plot
#| fig-height: 5

ggplot(trend, aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    x = "Days Relative to First NEDLAME Alert",
    y = "Milk Production (% of Own Pre-Alert Baseline)",
    color = "Treatment Group", fill = "Treatment Group",
    caption = "Shaded band: ± 1 standard error of the mean (≈ 68%). Not a test of the difference between arms."
  )
```

## By Alert Type

Does the milk trend differ by the type of the Nedap alert?

```{r}
#| label: fig-q1-alert-type-plot
#| fig-height: 5

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE),
            se_pct = sd(pct_of_baseline, na.rm = TRUE) / sqrt(n()),
            n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)",
       color = "Treatment Group", fill = "Treatment Group",
       caption = "Shaded band: ± 1 standard error of the mean (≈ 68%). Not a test of the difference between arms.") +
  theme(legend.position = "bottom")
```
:::

## Influence of Lameness History

Lifetime lameness history is important confounder in lameness work and this section considers the impact of that in this pilot. Sample size becomes and issue in this and this is presented mainly to consider for possible future work.

### Lameness History Distribution at Enrollment

@fig-chronic shows the breakdown of **New** vs **Chronic** cows that were enrolled. Chronicity is based on a cow's **lifetime record** at her first NEDLAME alert. Data going back to 2021 was used to establish chronicity.

- More New cases were enrolled

```{r}
#| label: fig-chronic
#| fig-height: 4

cohort |>
  count(tx_group, history_group) |>
  group_by(tx_group) |>
  filter(history_group %in% c("New", "Chronic")) |>
  mutate(pct = n / sum(n) * 100) |>
  ungroup() |>
  mutate(history_group = fct_relabel(history_group, \(x) str_wrap(x, width = 14))) |>
  ggplot(aes(x = history_group, y = n, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(n, " (", round(pct, 0), "%)")),
            position = position_dodge(width = 0.7), vjust = -0.3, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal() +
  labs(x = NULL, y = "N Cows", fill = "Treatment Group",
       ) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(lineheight = 0.9),
        plot.margin = margin(10, 15, 10, 15))
```

### Milk Trend by History Group

@fig-chronic-milk compares milk production across the lesion history types.

- Unlike the treatment groups, these are not randomised and care should be taken interpreting them and should considered descriptive.

  - This does suggest the numerical difference in milk production are driven by cows in the NEW category.

```{r}
#| label: fig-chronic-milk
#| fig-height: 6

milk_norm |>
  filter(history_group %in% c("New", "Chronic")) |>
  group_by(tx_group, history_group, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE),
            se_pct = sd(pct_of_baseline, na.rm = TRUE) / sqrt(n()),
            n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~history_group) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)",
       color = "Treatment Group", fill = "Treatment Group",
       caption = "Shaded band: ± 1 standard error of the mean (≈ 68%). Not a test of the difference between arms.") +
  theme(legend.position = "bottom")
```

## Time to Culling After NEDAP Enrollment

Time to culling is compared below. The curves sit close together, but only `r n_culls` cows were culled (`r n_culls_tx` TX, `r n_culls_ctl` Control).

That is too few to tell whether trimming off the alert changes culling, more cows and time is needed to answer that question.

```{r}
#| label: fig-q6-plot
#| fig-height: 5

km_cull |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% Culled") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: tbl-q6-table

km_cull |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% Culled After First NEDLAME Alert")
```

## What cows were missed by the system?

This section looks at every cow with a lesion, regardless of treatment group.

Among cows with a `LAME`-diagnosed lesion on or after `r params$history_after_date`, this checks whether a `NEDLAME` alert had reached DairyComp in the lookback window before it.

A "miss" means **no useful warning reached DairyComp before the lesion was found** - either the camera never flagged her, or it flagged her and the alert never arrived.

- A total of `r nrow(q4)` lesion cases were diagnosed after the project start date (`r params$history_after_date`), across `r n_distinct(q4$id_animal)` cows.

  - **`r n_misses` (`r round(100 * n_misses / nrow(q4), 1)`%) got no useful warning from the camera**

Those `r nrow(q4)` cases divide three ways:

1.  **`r sum(q4$caught_by_nedap)` caught in time** - an alert reached DairyComp inside the window before the lesion was found.

2.  **`r n_warned_not_acted` alerted earlier, with no action taken** - the camera warned about her further ahead and she was never examined in between. Counted separately rather than as a miss: the warning arrived, nobody acted on it.

3.  **`r n_misses` missed.** These divide two further ways:

    1.  **`r n_pipeline_loss` the camera flagged, but no alert reached DairyComp in time.** A pipeline problem - the filters we set up, or the daily file-in - rather than a limit of the camera.

    2.  **`r n_true_miss` the camera never flagged at all.**

Telling those last two apart needs the raw Nedap flag export rather than DairyComp alone, which covers `r format(nedap_coverage_start)` to `r format(nedap_coverage_end)`. DairyComp on its own can only answer "did a `NEDLAME` land in time"; it cannot tell a cow the camera never saw from one it flagged whose alert never arrived.

A breakdown of these cases by treatment group:

Each bar is one group of lesion cases, scaled to 100% so the four outcomes can be compared; the case count is printed above each bar, because the groups are very different sizes. The third group is **not a treatment arm** - it is every lesion case in a cow the pilot never enrolled, and it is the largest of the three.

```{r}
#| label: fig-q4-sensor-plot
#| fig-height: 4.4

n_by_arm <- q4_sensor |> count(tx_group, name = "n_cases")

# Quoted below, so the prose cannot drift from the figure.
arm_caught <- q4_sensor |>
  group_by(tx_group) |>
  summarize(pct = round(100 * mean(detection_group_3way == "Flagged, Caught in Time")),
            .groups = "drop")
pct_caught_arm <- function(a) arm_caught$pct[grepl(a, arm_caught$tx_group, fixed = TRUE)]
n_arm <- function(a) n_by_arm$n_cases[grepl(a, n_by_arm$tx_group, fixed = TRUE)]
arm_split <- q4_sensor |>
  group_by(tx_group) |>
  summarize(pct_warned = round(100 * mean(detection_group_3way == "Alerted earlier, no action taken")),
            pct_any    = round(100 * mean(detection_group_3way %in%
                          c("Flagged, Caught in Time", "Alerted earlier, no action taken"))),
            .groups = "drop")
pct_warned_arm <- function(a) arm_split$pct_warned[grepl(a, arm_split$tx_group, fixed = TRUE)]
pct_any_arm    <- function(a) arm_split$pct_any[grepl(a, arm_split$tx_group, fixed = TRUE)]

# Cows in the "not in pilot cohort" bar that nonetheless carry a NEDLAME. They
# were alerted and then dropped from the cohort by the post-trim rule, so the
# bucket name is true of ENROLMENT, not of the camera. Named in the prose so a
# reader who spots the sliver is not left guessing.
n_unenrolled_alerted <- sum(grepl("Never Alerted", q4_sensor$tx_group) &
                              (q4_sensor$caught_by_nedap | q4_sensor$warned_not_acted))

q4_sensor |>
  count(tx_group, detection_group_3way) |>
  ggplot(aes(x = tx_group, y = n, fill = detection_group_3way)) +
  geom_col(position = "fill") +
  geom_text(data = n_by_arm, inherit.aes = FALSE,
            aes(x = tx_group, y = 1.04, label = paste0("n=", n_cases)),
            size = 3, colour = "grey30") +
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.25)) +
  coord_cartesian(ylim = c(0, 1.08), clip = "off") +
  scale_fill_viridis_d() +
  guides(fill = guide_legend(nrow = 2)) +
  theme_minimal() +
  labs(x = NULL, y = "% of lesion cases in that group", fill = "Alert Status") +
  theme(legend.position = "bottom")
```

Two of the four outcomes come from DairyComp alone - *Flagged, Caught in Time* and *Alerted earlier, no action taken*. The other two can only be told apart using the raw Nedap export: without it there is no way to distinguish a flag that never arrived from a cow the camera never saw.

Within the pilot, the two arms differ far less than the bars suggest. A `NEDLAME` reached DairyComp before the lesion in `r pct_any_arm("Control")`% of Control cases and `r pct_any_arm("TX")`% of TX cases - effectively the same. **Read those two figures as a comparison, not as a sensitivity.** A cow enters either arm by having received a `NEDLAME` at some point, so a high prior-alert rate is partly built into how the cohort is defined; what that selection does not explain is the difference *between* the arms, since both are selected the same way. **And what differs is whether anyone acted.** `r pct_caught_arm("TX")`% of TX cases were caught in time against `r pct_caught_arm("Control")`% of Control, because a Control alert nobody acted on falls into *alerted earlier, no action taken* - `r pct_warned_arm("Control")`% of Control cases against `r pct_warned_arm("TX")`% of TX. That gap is protocol compliance, not detection, which is the same distinction the trimming section draws. The third bar is different in kind rather than degree: almost all of it is *never flagged* or *flagged, but no alert in time*, because a cow outside the cohort has no `NEDLAME` by construction. **That bar is therefore not a pipeline failure rate** - those are cows the camera did see, in a study that never enrolled them.

One wrinkle, since it is visible as a thin sliver on that third bar: `r n_unenrolled_alerted` of the `r n_arm("Never Alerted")` unenrolled cases do carry a `NEDLAME`. They were alerted and then dropped from the cohort by the post-trim rule, so "not in pilot cohort" describes enrolment, not the camera.

### How solid are those two numbers?

`r if (n_misses_outside_coverage > 0) paste0("- The other ", n_misses_outside_coverage, " misses were diagnosed after ", format(nedap_coverage_end), ", which the current export can't see. These are left unclassified rather than assumed.") else paste0("Every miss here is checkable against the raw sensor feed: the analysis stops at ", format(analysis_end), ", the last day the Nedap export covers.")`

One judgement call sits behind these numbers. `r n_attentions_incomplete` of the resolved flags are marked not completed in the raw export, and are counted here as the camera having flagged the cow. `r n_pipeline_loss_incomplete_only` of the `r n_pipeline_loss` pipeline losses rest on such a flag alone; counting only completed flags would give `r n_pipeline_loss_completed_only`. Stated rather than assumed.

## What Lesions are being missed?

The figure below shows the lesion types found at these first-diagnosis cases, broken out by NEDAP detection status:

- **Flagged, Caught in Time**
  - an alert landed in DairyComp inside the window before the lesion.
- **Flagged, but no alert in time**
  - the camera did flag her, but no alert reached DairyComp in time.
- **Never Flagged (True Miss)**
  - the camera never saw a problem at all.
- **Alerted earlier but no action taken**.
  - camera flagged the cow but for unknow reasons she was not trimmed.

Where a cow had two different lesions on the same day, each type is counted. The same type coded twice on one day counts once.

Two views of the same cases, because either one alone invites a wrong reading. **Panel A** asks how often each lesion is caught. **Panel B** asks what each outcome tends to be, with *Flagged, Caught in Time* as the comparison column - without it, a common lesion looks like a blind spot simply by being common.

The four-way split is kept deliberately. The farm report collapses it to alerted / not alerted, which is what the farm needs; Nedap needs *flagged, but no alert in time* (an integration problem) held apart from *never flagged* (a detection problem).

Lesion types with fewer than 20 cases are combined into one bar - separately they are too few to carry a percentage.

```{r}
#| label: fig-q4-lesion-types
#| fig-height: 9

invisible(list2env(fxn_build_q4_groups(q4_sensor, lame_lesion_recent), environment()))

STATUS_LEV <- c("Flagged, Caught in Time", "Alerted earlier, no action taken",
                "Flagged, but no alert in time", "Never Flagged (True Miss)")

les_base <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, detection_group_3way) |>
  mutate(status = factor(as.character(detection_group_3way), levels = STATUS_LEV))
# Same trap as elsewhere: a renamed level silently becomes NA and vanishes from
# every count while the page still renders. Assert the mapping caught everything.
stopifnot(!any(is.na(les_base$status)), nrow(les_base) > 0)

les_long <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(les_base |> select(id_animal, lact_number, date_event, status),
             by = c("id_animal", "lact_number", "date_event")) |>
  pivot_longer(all_of(lesion_type_cols), names_to = "lesion_type", values_to = "has_lesion") |>
  filter(has_lesion == 1)

# Types under 20 cases cannot carry a percentage: toe ulcer is 2 cows, and under
# the previous per-type chart those 2 rendered as a full-width 100% bar.
les_min_n  <- 20
les_tot    <- les_long |> count(lesion_type, name = "n_total")
les_rare   <- sort(les_tot$lesion_type[les_tot$n_total < les_min_n])
n_les_rare <- sum(les_tot$n_total[les_tot$n_total < les_min_n])
les_long   <- les_long |>
  mutate(lt = if_else(lesion_type %in% les_rare, "rare types (combined)", lesion_type))

n_by_lesion    <- les_long |> count(lt, name = "n_cases")
rate_by_lesion <- les_long |> count(lt, status) |>
  left_join(n_by_lesion, by = "lt") |>
  mutate(pct = 100 * n / n_cases)

# Best to worst, left to right, by share caught in time.
les_ord <- rate_by_lesion |> filter(status == STATUS_LEV[1]) |>
  arrange(desc(pct)) |> pull(lt)
les_ord <- c(les_ord, setdiff(unique(les_long$lt), les_ord))
rate_by_lesion <- rate_by_lesion |> mutate(lt = factor(lt, levels = les_ord))
n_by_lesion    <- n_by_lesion    |> mutate(lt = factor(lt, levels = les_ord))

# Okabe-Ito, colourblind-safe, and deliberately NOT viridis: panel A uses viridis
# for STATUS, and two palettes both ending in yellow would make the lesion "dd"
# read as the status "Never Flagged".
les_pal <- setNames(c("#0072B2", "#009E73", "#CC79A7", "#56B4E9", "#D55E00",
                      "#E69F00", "#999999")[seq_along(les_ord)], les_ord)

caught_lab <- rate_by_lesion |> filter(status == STATUS_LEV[1]) |>
  mutate(lab = paste0(round(pct), "%"))

# Best at the TOP: coord_flip() puts the first factor level at the bottom, so
# the display order is reversed here rather than in the ordering logic above.
rate_by_lesion <- rate_by_lesion |> mutate(lt = factor(lt, levels = rev(les_ord)))
n_by_lesion    <- n_by_lesion    |> mutate(lt = factor(lt, levels = rev(les_ord)))
caught_lab     <- caught_lab     |> mutate(lt = factor(lt, levels = rev(les_ord)))

p_lesion_rate <- ggplot(rate_by_lesion, aes(x = lt, y = pct, fill = fct_rev(status))) +
  geom_col(width = 0.72) +
  geom_text(data = caught_lab, inherit.aes = FALSE,
            aes(x = lt, y = pct / 2, label = lab), colour = "white", size = 3.4) +
  geom_text(data = n_by_lesion, inherit.aes = FALSE,
            aes(x = lt, y = 102, label = paste0("n=", n_cases)),
            hjust = 0, size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 100, 25), labels = function(x) paste0(x, "%"),
                     limits = c(0, 116), expand = expansion(mult = c(0, 0))) +
  scale_fill_viridis_d(drop = FALSE, direction = -1,
                       guide = guide_legend(reverse = TRUE, nrow = 2)) +
  theme_minimal() +
  labs(x = NULL, y = "% of cases with that lesion", fill = NULL,
       title = "A. How often is each lesion caught?",
       subtitle = "Within each lesion type, best at the top. The figure inside each bar is the share caught in time.") +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold", size = 11),
        axis.text.y = element_text(colour = rev(les_pal[les_ord]), face = "bold"),
        panel.grid.major.y = element_blank())

n_by_status   <- les_base |> count(status, name = "n_cases")
mix_by_status <- les_long |> count(status, lt) |>
  left_join(n_by_status, by = "status") |>
  # rev() so that after coord_flip() the within-group order reads top-to-bottom
  # in the same order as the legend; guide_legend(reverse) keeps the two aligned.
  mutate(pct = 100 * n / n_cases, lt = factor(lt, levels = rev(les_ord)))

# Outcomes best-to-worst reading DOWN the axis. The previous vertical version
# claimed an order the reader could not see: the x-axis was ordered, but the eye
# went to the dodged bars inside each group, whose heights are not monotonic.
mix_by_status <- mix_by_status |>
  mutate(status = factor(as.character(status), levels = rev(STATUS_LEV)))
n_by_status   <- n_by_status   |>
  mutate(status = factor(as.character(status), levels = rev(STATUS_LEV)))

p_lesion_mix <- ggplot(mix_by_status, aes(x = status, y = pct, fill = lt)) +
  geom_col(width = 0.78, position = position_dodge(width = 0.85)) +
  geom_text(aes(label = round(pct)), position = position_dodge(width = 0.85),
            hjust = -0.25, size = 2.7, colour = "grey25") +
  geom_text(data = n_by_status, inherit.aes = FALSE,
            aes(x = status, y = -4.5, label = paste0("n=", n_cases)),
            size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = c(0.10, 0.10))) +
  scale_fill_manual(values = les_pal, guide = guide_legend(nrow = 1, reverse = TRUE)) +
  scale_x_discrete(labels = function(x) gsub(", ", ",\n", x)) +
  theme_minimal() +
  labs(x = NULL, y = "% of cases in that status", fill = NULL,
       title = "B. What does each outcome tend to be?",
       subtitle = "Outcomes best to worst reading down. Bars do not sum to 100% - a cow can carry more than one lesion.") +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold", size = 11),
        axis.text.y = element_text(size = 8), panel.grid.major.y = element_blank())

# Quoted in the prose below, so the text cannot drift from the figure.
pctl <- function(l, s) round(rate_by_lesion$pct[rate_by_lesion$lt == l & rate_by_lesion$status == s])
pctm <- function(l, s) round(mix_by_status$pct[mix_by_status$lt == l & mix_by_status$status == s])

p_lesion_rate / p_lesion_mix
```

Digital dermatitis is the clear gap. The camera caught **`r pctl('dd', STATUS_LEV[1])`%** of DD cases in time against **`r pctl('wld', STATUS_LEV[1])`%** of white line, and DD is **`r pctm('dd', STATUS_LEV[4])`%** of the cases it never flagged against **`r pctm('dd', STATUS_LEV[1])`%** of those it caught. White line runs the other way on both counts. Hemorrhage, injury and "other" - typically corkscrew claw on this farm - sit in between.

None of that is a surprise. Locomotion scoring is known to detect digital dermatitis poorly, and the camera is scoring locomotion, so a lesion that barely changes how a cow walks is the one it will tend to miss. What this adds is the size of the gap on this farm, and that it is specific rather than general: white line, which does change how a cow walks, is caught well.

One point for Nedap in particular: **DD is over-represented in both failure modes**, not only in detection. It is `r pctm('dd', STATUS_LEV[3])`% of the cases whose flag produced no `NEDLAME`, as well as `r pctm('dd', STATUS_LEV[4])`% of those never flagged at all. **The first of those is the surprising one** - it rests on only `r n_by_status$n_cases[n_by_status$status == STATUS_LEV[3]]` cases, so it may be noise, but it is worth putting to Nedap rather than leaving unstated.

`r length(les_rare)` lesion types with fewer than `r les_min_n` cases (`r paste(les_rare, collapse = ", ")`; `r n_les_rare` cases in total) are combined into a single bar.

## Which leg is being missed?

The same cases split by which foot the lesion was found on. A cow can have lesions on more than one foot, so the bars add to more than `r nrow(q4_group)` cases.

```{r}
#| label: fig-missed-by-leg
#| fig-height: 4

leg_names <- c(LF = "Left Front", RF = "Right Front", LR = "Left Rear", RR = "Right Rear")

# locate_lesion packs one or more two-character foot codes into one string
# ("RRLR" = right rear + left rear, "RRLRRFLF" = all four). It must be parsed
# in 2-character chunks, NOT with substring matching: str_detect("LRRF", "RR")
# is TRUE, but "LRRF" is LR + RF, so substring matching would invent a
# right-rear lesion that was never recorded. Checked every value in the data -
# all are even length and every chunk is one of LF/RF/LR/RR.
leg_long <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM"), !is.na(locate_lesion), locate_lesion != "") |>
  select(id_animal, lact_number, date_event, locate_lesion) |>
  # locate_lesion packs one or more two-character foot codes into one string
  # ("RRLR" = right rear + left rear, "RRLRRFLF" = all four), so it must be
  # read in 2-character chunks, never by substring matching: str_detect("LRRF",
  # "RR") is TRUE but "LRRF" is LR + RF.
  #
  # It is built in functions/fxn_location.R, which used to search the whole
  # remark and so invented a phantom left front on XNLF* codes - XNL
  # (treatment) + F (foot rot) + RF (leg) is ONE foot, but "LF" spans the
  # treatment's trailing L and the lesion's F. Fixed at source, so nothing is
  # needed here.
  mutate(leg = str_extract_all(locate_lesion, ".{2}")) |>
  unnest(leg) |>
  filter(leg %in% names(leg_names)) |>
  distinct(id_animal, lact_number, date_event, leg) |>
  # Map to readable names HERE, once, so everything downstream sees
  # "Left Rear" rather than "LR". Leaving the raw codes in place and
  # renaming later silently broke a front/rear split that tested
  # str_detect(leg, "Rear") - it matched nothing and called every foot Front.
  mutate(leg = unname(leg_names[leg]))

q4_legs <- leg_long |>
  inner_join(q4_sensor |> select(id_animal, lact_number, date_event, detection_group_3way),
             by = c("id_animal", "lact_number", "date_event")) |>
  count(detection_group_3way, leg) |>
  # Same reasoning as the lesion-type figure above: a share within each foot,
  # so the four feet are comparable regardless of how many lesions each carries.
  group_by(leg) |>
  mutate(total_cases = sum(n), pct = 100 * n / total_cases) |>
  ungroup() |>
  mutate(leg = fct_reorder(leg, total_cases))

leg_totals <- q4_legs |> distinct(leg, total_cases)

# leg is a factor now (ordered for the plot), and str_detect() rejects factors
n_legs_rear <- sum(q4_legs$n[str_detect(as.character(q4_legs$leg), "Rear")])
n_legs_front <- sum(q4_legs$n[str_detect(as.character(q4_legs$leg), "Front")])

ggplot(q4_legs, aes(x = leg, y = pct, fill = detection_group_3way)) +
  geom_col(width = 0.7) +
  geom_text(data = leg_totals, inherit.aes = FALSE,
            aes(x = leg, y = 101, label = paste0("n=", total_cases)),
            hjust = 0, size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 100, 25), labels = function(x) paste0(x, "%"),
                     limits = c(0, 118), expand = expansion(mult = c(0, 0))) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases on that foot", fill = NULL) +
  theme(legend.position = "bottom")
```

Rear feet account for `r n_legs_rear` of the `r n_legs_rear + n_legs_front` foot-level lesions, front feet `r n_legs_front`. That is where lesions occur, not where the camera fails - which segment of a stacked bar is tallest says nothing about detection, because every segment scales with how common lesions are on that foot. Whether the camera misses front feet more than hind feet is asked directly below.

## What is the camera actually missing?

```{r}
#| label: true-miss-build

checkable <- q4_sensor |>
  filter(in_coverage) |>
  mutate(camera = if_else(detection_group_3way == "Never Flagged (True Miss)",
                          "Camera never flagged", "Camera flagged her"))

n_cam_miss <- sum(checkable$camera == "Camera never flagged")
n_cam_flag <- sum(checkable$camera == "Camera flagged her")
# Same trap, same guard: both groups non-empty and accounting for every case.
stopifnot(n_cam_miss > 0, n_cam_flag > 0, n_cam_miss + n_cam_flag == nrow(checkable))

lesion_by_camera <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(checkable |> select(id_animal, lact_number, date_event, camera),
             by = c("id_animal", "lact_number", "date_event")) |>
  pivot_longer(all_of(lesion_type_cols), names_to = "lesion_type", values_to = "has_lesion") |>
  filter(has_lesion == 1) |>
  count(camera, lesion_type) |>
  mutate(pct = if_else(camera == "Camera never flagged", 100 * n / n_cam_miss, 100 * n / n_cam_flag))

pct_dd_miss  <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "dd"  & lesion_by_camera$camera == "Camera never flagged"], 0)
pct_dd_flag  <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "dd"  & lesion_by_camera$camera == "Camera flagged her"], 0)
pct_wld_miss <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "wld" & lesion_by_camera$camera == "Camera never flagged"], 0)
pct_wld_flag <- round(lesion_by_camera$pct[lesion_by_camera$lesion_type == "wld" & lesion_by_camera$camera == "Camera flagged her"], 0)

# Front/rear share within each group, computed here rather than inline so the
# prose stays readable.
leg_end <- leg_long |>
  mutate(end = if_else(str_detect(leg, "Rear"), "Rear", "Front")) |>
  distinct(id_animal, lact_number, date_event, end)

leg_by_camera <- leg_end |>
  inner_join(checkable |> select(id_animal, lact_number, date_event, camera),
             by = c("id_animal", "lact_number", "date_event")) |>
  count(camera, end) |>
  group_by(camera) |>
  mutate(pct = 100 * n / sum(n)) |>
  ungroup()

pct_rear_miss <- round(leg_by_camera$pct[leg_by_camera$camera == "Camera never flagged" & leg_by_camera$end == "Rear"], 0)
pct_rear_flag <- round(leg_by_camera$pct[leg_by_camera$camera == "Camera flagged her"   & leg_by_camera$end == "Rear"], 0)
```

The `r n_cam_miss` cases the camera never flagged are compared below against the `r n_cam_flag` it did. Both sit inside the sensor coverage window, so both have raw flag data behind them. Percentages are within each group, and a cow can have several lesion types and several affected feet, so they add to more than 100%.

```{r}
#| label: fig-true-miss-lesion
#| fig-height: 4

lesion_by_camera |>
  ggplot(aes(x = fct_reorder(lesion_type, pct, .fun = max), y = pct, fill = camera)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases in that group", fill = NULL) +
  theme(legend.position = "bottom")
```

- **Digital dermatitis is over-represented**: `r pct_dd_miss`% of the cases the camera never flagged, against `r pct_dd_flag`% of the ones it did flag.

- **White line is the opposite**: `r pct_wld_miss`% of misses vs `r pct_wld_flag`% of flagged cases. The camera picks white line up well.

- **Front versus hind is not where the difference is.** Rear feet are `r pct_rear_miss`% of the missed cases against `r pct_rear_flag`% of the ones the camera did flag. Hind-foot lesions dominate both groups in much the same proportion, so there is no sign the camera is specifically blind to front feet.

So the difference is in the lesion, not the limb. That is no surprise: locomotion scoring detects digital dermatitis poorly, and the camera scores locomotion, so a lesion that barely changes how a cow walks is the one it will miss. What this adds is the size of the gap here, and that it is specific rather than general - white line, which does change the gait, is caught well.

Since we are running a separate SmartSight instance, I wanted to make sure all cows were visible on it. A cow absent from SmartSight is not being watched for lameness at all, so a lesion found on her cannot be a camera miss.

```{r}
#| label: platform-enrollment-build

q4_sensor <- q4_sensor |>
  left_join(events_all |> select(id_animal, id) |> distinct(), by = "id_animal") |>
  mutate(
    # alias-aware, matching the Q4 eligibility rule exactly - a cow enrolled
    # under a previous animal number is enrolled (see enrolled_ids in setup)
    not_in_smartsight = !(id %in% enrolled_ids),
    # Checked, not assumed: the point of this section is that a cow can be
    # current on the Activity platform (which staff use daily) while absent
    # from SmartSight. Earlier drafts asserted "even though she's on
    # Activity" in prose without computing it - now computed so the claim
    # can't drift away from the data.
    in_activity = id %in% cows_activity$id
  )

# Unenrolled cows are excluded upstream in Q4, so this should now be empty.
# Kept as an assertion rather than deleted: if it ever stops being zero, the
# exclusion has silently stopped working.
n_residual_not_in_smartsight <- sum(q4_sensor$is_miss & q4_sensor$not_in_smartsight)
n_total_misses <- sum(q4_sensor$is_miss)

# Herd-level enrollment check. The misses above are only the unenrolled cows
# who also happened to get a lesion, so they understate the gap. This asks the
# direct question instead: of every first-lactation cow that freshened in the
# study window, what share is on the SmartSight list at all? Compared against
# the Activity list, which is the control - if both dropped together it would
# point at the export; if only SmartSight drops it points at enrollment.
ss_ids  <- unique(cows_smartsight$id)
act_ids <- unique(cows_activity$id)

l1_fresh <- animal_lactations |>
  filter(lact_number == 1, !is.na(date_fresh), date_fresh >= ymd(params$history_after_date) - 150,
         date_fresh <= analysis_end) |>
  distinct(id_animal, id, date_fresh) |>
  mutate(id = as.character(id),
         on_smartsight = id %in% ss_ids,
         on_activity   = id %in% act_ids,
         fresh_month   = format(date_fresh, "%Y-%m"))

enroll_by_month <- l1_fresh |>
  group_by(`Freshened` = fresh_month) |>
  summarize(`Cows` = n(),
            `% on SmartSight` = round(100 * mean(on_smartsight), 1),
            `% on Activity`   = round(100 * mean(on_activity), 1),
            .groups = "drop")

# the hard break: the most recent freshening among enrolled first-lactation cows
last_enrolled_fresh <- max(l1_fresh$date_fresh[l1_fresh$on_smartsight], na.rm = TRUE)
l1_after  <- l1_fresh |> filter(date_fresh > last_enrolled_fresh)
l1_before <- l1_fresh |> filter(date_fresh <= last_enrolled_fresh)
n_l1_after_unenrolled <- sum(!l1_after$on_smartsight)
pct_l1_before_enrolled <- round(100 * mean(l1_before$on_smartsight), 1)
n_l1_after_on_activity <- sum(l1_after$on_activity)

```

`r n_excluded_not_enrolled` lesion cases were excluded on this basis, and they are no longer counted as misses anywhere above.

Those are only the unenrolled cows who also happened to get a lesion, so they understate the problem. Asking the question directly - of every first-lactation cow that freshened in the window, how many are on each platform:

```{r}
#| label: tbl-enrollment-by-month

enroll_by_month |>
  gt() |>
  tab_header(title = "SmartSight enrollment of first-lactation cows stops; Activity does not") |>
  tab_source_note("Activity is the control here. Both lists were pulled on the same day, so a drop in one and not the other cannot be an export problem.")
```

- The most recent freshening among first-lactation cows who *are* on SmartSight is **`r format(last_enrolled_fresh)`**. Before that date `r pct_l1_before_enrolled`% were enrolled; after it, **`r n_l1_after_unenrolled` first-lactation cows are absent**, and `r n_l1_after_on_activity` of them are on Activity.

- Both lists were pulled on the same day, well after the cut-off, so this is not a stale-export artifact. A drop in SmartSight with Activity unaffected points at enrollment having stopped, not at data transfer.

- **Animal numbers get reassigned, and that had to be handled before this check could be trusted.** DairyComp logs an ID change as an `XID` event holding the old number, and `r nrow(id_aliases)` such changes appear in this extract. A cow renumbered after she was enrolled looks absent under her current number while being present under her old one - cow 23143, renumbered from 14347, is enrolled all along as 14347. Enrollment is therefore checked against a cow's current *and* prior numbers; `r n_rescued_by_alias` animals are only found that way.

## Why Did No Alert Arrive in Time?

Two rules limit which cows can become a `Low` alert: a score under 31, and no history of upper-leg lameness (`Leg-AboveFoot` in the trim `protocols`).

Score-at-flag isn't available in this data, so only the injury/`Leg-AboveFoot` half of this can be checked. It is shown in the figure below.

```{r}
#| label: fig-pipeline-loss-injury-plot
#| fig-height: 4

# `reason` assigns ONE cause per case, so these bars sum to the total and
# match the counts in the prose. They used to be computed separately here.
pipeline_loss_cases |>
  count(category = reason) |>
  ggplot(aes(x = fct_reorder(category, n), y = n, fill = category)) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "N Cases") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

**Every cow in this section was still missed.** She is counted in the `r sum(q4_sensor$is_miss)` above either way — the question here is only whether the camera's signal went astray on the way to DairyComp, or whether the system was correctly declining to raise an alert.

Four rules can explain this without anything having gone missing. Each case is counted once, under the first rule that applies, so the figures below add to `r nrow(pipeline_loss_cases)`. Of those cases:

- `r n_only_ftdat` had every flag land inside its own `FTDAT` window (90 days for `Low`, 28 for the declines). DairyComp was right to skip them.

- `r n_only_injury` were `LOW`-only flags on cows with upper-leg lameness history, which blocks the `Low` route.

`r if (n_only_dsnlm > 0) paste0('- ', n_only_dsnlm, ' had every flag suppressed by DSNLM: she had already been alerted inside the window (90 days on the Low route, 7 on the declines), so a repeat alert is deliberately withheld to keep the cowcard clean.') else paste0('- No case is now explained by DSNLM alone. That rule needs a prior NEDLAME before the flag, and a cow with an earlier alert has already left this population - either caught inside the lookback, or set aside as warned-and-not-acted-on. The two rules overlap almost entirely. It is not unreachable: a cow alerted, then trimmed, then flagged again inside the 7-day decline window would still land here, because the trim keeps her a miss. None currently does.')`

- `r n_only_late` did get an alert into DairyComp — the day after the lesion, too late to act on. The daily batch runs at \~5am and stamps an alert with the day it was loaded, so an attention after \~5am on day D is logged as D+1, landing just outside a lookback that only looks backward. Two readings fit and the data cannot separate them: either that flag arrived late, or the camera made a fresh observation of her post-trim gait and *that* is what came through. Under the second reading the original flag really was lost. Either way she was missed, and either way DairyComp was receiving something from the camera around that date — which is why these are not counted among the confirmed losses below.

That accounts for **`r n_pipeline_loss_by_design` of the `r nrow(pipeline_loss_cases)`**. The remaining `r n_pipeline_loss_unexplained` are not all failures either, and it matters to separate them:

- **`r n_pipeline_loss_unknowable` cannot be judged at all.** Only a `LOW` flag fired, and the `Low` route requires a score of 1-30. Score-at-flag is not in this data, so a cow scoring 31-69 who was correctly declined looks exactly like a lost flag. These are unresolved, not unexplained.

- **`r n_pipeline_loss_genuine` are genuine pipeline losses.** A decline route fired, and the decline routes carry no score gate — so there is no rule left that would explain the silence. These are the cases worth taking to Nedap, and cow-level detail for them is in `pipeline_losses.xlsx`.

So the integration problem is **`r n_pipeline_loss_genuine` confirmed cases, not `r nrow(pipeline_loss_cases)`** — with an upper bound of `r n_pipeline_loss_unexplained` if every unjudgeable case turned out to be a failure, which is unlikely.

One earlier caveat is retired. The `FTDAT` lookup is now cow-level, matching DairyComp, where it is a lifetime date that does not reset at freshening. It used to be scoped to the current lactation, which hid prior-lactation trims and under-counted the correct suppressions. The remaining `r n_pipeline_loss_unknowable` are unjudgeable for the score reason alone.

## Of the Misses, Did Farm Staff Catch Them Anyway?

```{r}
#| label: q4-staff-catch-build

invisible(list2env(fxn_build_staff_catch(q4, chklame), environment()))
```

Of the `r nrow(staff_catch)` cows whose lesion had no `NEDLAME` alert recorded in DairyComp beforehand (they were trimmed on the routine round or because staff picked them out),

- `r sum(staff_catch$staff_flagged)` (`r round(100 * mean(staff_catch$staff_flagged), 1)`%) were flagged by the farm staff in the 7 days before diagnosis.

## Summarizing the Misses

Each check above narrows a different group of cows, so the percentages don't share a denominator. The table below ties them together. Indented rows are splits of the row above, and the "% of" column always refers to the base named beside it.

```{r}
#| label: tbl-q4-reconciliation

n_caught_total <- sum(q4$caught_by_nedap)
n_staff_caught <- sum(staff_catch$staff_flagged)

q4_recon <- tibble::tribble(
  ~step, ~n, ~base, ~base_n,
  "Qualifying lesion cases (after eligibility exclusions)", nrow(q4), "—", NA_integer_,
  " Caught by NEDAP in time", n_caught_total, "qualifying cases", nrow(q4),
  " Not caught — \"misses\"", n_total_misses, "qualifying cases", nrow(q4),
  "  Within raw sensor coverage (checkable)", nrow(misses_in_coverage), "misses", n_total_misses,
  "  Outside coverage (not checkable yet)", n_misses_outside_coverage, "misses", n_total_misses,
  "   Flag fired anyway — candidate pipeline loss", n_pipeline_loss, "checkable misses", nrow(misses_in_coverage),
  "   Never flagged — true miss", n_true_miss, "checkable misses", nrow(misses_in_coverage),
  "    Of candidates: blocked by its own FTDAT window", n_pipeline_loss_ftdat_blocked, "candidates", n_pipeline_loss,
  "    Of candidates: LOW-only + injury/upper-leg history", n_pipeline_loss_injury_explained, "candidates", n_pipeline_loss,
  "    Of candidates: by design overall (either rule)", n_pipeline_loss_by_design, "candidates", n_pipeline_loss,
  "    Of candidates: cannot tell (LOW only, no score)", n_pipeline_loss_unknowable, "candidates", n_pipeline_loss,
  "    Of candidates: GENUINE pipeline loss", n_pipeline_loss_genuine, "candidates", n_pipeline_loss
) |>
  mutate(pct = if_else(is.na(base_n), NA_real_, 100 * n / base_n))

q4_recon |>
  select(step, n, pct, base) |>
  gt() |>
  fmt_number(columns = pct, decimals = 1) |>
  sub_missing(columns = pct, missing_text = "") |>
  cols_label(step = "", n = "N", pct = "% of base", base = "…of what") |>
  cols_align(align = "left", columns = step) |>
  tab_header(
    title = "Q4 Reconciliation: How the Populations Nest",
    subtitle = "Indented rows split the row above them"
  ) |>
  tab_source_note(source_note = paste0(
    "Cows never enrolled in SmartSight are excluded before this chain starts (",
    n_excluded_not_enrolled, " lesion cases), on the same basis as dry cows and unfreshened heifers - ",
    "so they are not a row here and the residual inside the chain is ", n_residual_not_in_smartsight, ". ",
    "One check does cut ACROSS the chain rather than extending it: ", n_staff_caught, " of the ", n_total_misses,
    " misses were independently flagged by staff, which can apply to any row."))
```

**The bottom line of this:**

- of the `r n_total_misses` cases DairyComp records as missed,

- only `r nrow(misses_in_coverage)` can currently be checked against the raw sensor feed at all.

  Within that checkable slice,

- `r n_true_miss` are genuine missing alerts

- `r n_pipeline_loss` are cases where the camera did flag her

  - `r n_pipeline_loss_by_design` were correct by-design suppressions (`FTDAT` window or injury/upper-leg rule), not integration failures,

  - `r n_pipeline_loss_unknowable` cannot be judged, because only a `LOW` flag fired and score-at-flag is not in the data,

  - leaving **`r n_pipeline_loss_genuine` genuine pipeline losses** — `r round(100 * n_pipeline_loss_genuine / nrow(misses_in_coverage), 1)`% of the checkable misses. That, not the headline miss rate, is the size of the integration problem.
