---
title: "NEDLAME Alert Pilot: TX (Trim) vs Control"
editor: source
format:
 html:
   embed-resources: true
   toc: true
   toc-location: left
   code-fold: true
execute:
    echo: false
    message: false
    warning: false
params:
  milk_window_pre: 14
  milk_window_post: 30
  milk_baseline_days: 7
  q4_lookback_days: 21
  history_after_date: "2026-06-01"
  recent_trim_suppression_days: 28
  post_trim_exclusion_days: 7
  analysis_end_date: "2026-08-27"
---

```{r}
#| label: setup

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  arrow,
  lubridate,
  gt,
  gtsummary,
  survival,
  survminer,
  ggsurvfit,
  scales,
  viridis
)

source(file.path('functions/fxn_read_milk_folder.R'))
source(file.path('functions/fxn_read_nedap_attentions.R'))

# GC's lame functions (fxn_collapse_lesions, fxn_trim_vars, fxn_dz_status) -
# fetched live from GitHub, same mechanism report_explore_lame_new.qmd uses.
# Two local overrides are sourced AFTER fxn_load_os_fxns() specifically so
# they win over the GitHub versions in the global environment:
#  - functions/fxn_code_lesions.R: the GitHub version's `trimonly` logic
#    only checks the *protocols* field for "NONE" - this farm's
#    routine/no-lesion trims are coded "NONE" in the *remark* field
#    instead, which the GitHub version doesn't check, so it silently
#    misclassifies every routine trim as having a lesion.
#  - functions/fxn_collapse_lesions.R: the GitHub version reassigns its
#    own `data` variable to an already-one-row-per-date subset before
#    computing the max() union across that date's lesion columns, so
#    when a cow has multiple event rows on the same date coding
#    DIFFERENT lesions, only whichever row survives the earlier
#    reduction keeps its lesion flags - the rest are silently dropped
#    (confirmed against a real cow with 4 same-day LAME rows: only 1 of
#    her 3 distinct lesion types survived under the GitHub version).
source(file.path('functions/fxn_load_os_fxns.R'))
fxn_load_os_fxns()
source(file.path('functions/fxn_code_lesions.R'))
source(file.path('functions/fxn_collapse_lesions.R'))

# Shared analysis functions: the cohort, exclusions and Q1-Q6 builds, so this
# report and the full report apply the same rules from one definition.
source(file.path('functions/fxn_nedlame_analysis.R'))

events_all <- read_parquet('data/intermediate_files/events_all_columns.parquet')
events_formatted <- read_parquet('data/intermediate_files/events_formatted.parquet')
animals <- read_parquet('data/intermediate_files/animals.parquet')
animal_lactations <- read_parquet('data/intermediate_files/animal_lactations.parquet')

date_max_pull <- max(events_all$date_event, na.rm = TRUE)
farm_name <- unique(events_all$location_event)

# NEDLAME go-live date, per the DairyComp setup: enrollment protocols were
# only turned on starting ~June 1 2026, so no NEDLAME alert can exist
# before this regardless of how "missed" a case looks.
nedlame_start_date <- min(events_all$date_event[events_all$Event == "NEDLAME"], na.rm = TRUE)
```

```{r}
#| label: cohort

# Cohort, exclusions and the analysis cut-off all live in
# functions/fxn_nedlame_analysis.R so every report applies the same rules.
invisible(list2env(fxn_build_cohort(events_all, animal_lactations, animals,
                                    date_max_pull, params), environment()))
```

```{r}
#| label: lesion-history-dataset

# Returns an augmented cohort (adds history_group), so it is reassigned here.
invisible(list2env(fxn_build_lame_history(events_formatted, cohort), environment()))
```

```{r}
#| label: milk-data

# data/milk/ is not wired into any existing pipeline script. Raw milk rows
# only carry a raw `id` (not the composite id_animal), so they're resolved
# to a specific animal-lactation by nearest-preceding date_fresh (see
# fxn_read_milk_folder.R for why a naive date_dry window isn't safe here).
# Only the window Q1 actually uses is read. The folder holds every milk day
# back to 2023, but the curves only span milk_window_pre days before an alert
# to milk_window_post days after, so reading the lot wastes ~26 s and then
# hands ~10x more rows than needed to the resolver, whose join costs far more
# than the read. Derived from the data rather than hard-coded, so it follows
# the alert window and the params automatically.
milk_from <- min(cohort$first_nedlame_date, na.rm = TRUE) - params$milk_window_pre
milk_to   <- max(cohort$first_nedlame_date, na.rm = TRUE) + params$milk_window_post

milk_long <- fxn_read_milk_folder(from = milk_from, to = milk_to)
milk_resolved <- fxn_resolve_milk_to_lactation(milk_long, animal_lactations)
```

```{r}
#| label: nedap-attentions-data

# Raw Nedap sensor export - what the SENSOR actually flagged, independent
# of whether that flag ever reached DairyComp as a NEDLAME event. Used in
# Q4 to split "NEDAP missed it" cases into true sensor misses vs pipeline
# losses (a real cowcard/Pulse cross-check first surfaced this pattern -
# see the plan doc's Round 8/10). This is a one-time export, not a live
# feed, so its coverage window is whatever was current when it was
# pulled - captured here explicitly rather than assumed, since it can
# lag behind the DairyComp/milk data read in above.
attentions_raw <- fxn_read_nedap_attentions()
attentions_resolved <- fxn_resolve_nedap_to_lactation(attentions_raw, animal_lactations)

nedap_coverage_start <- min(attentions_resolved$attention_date, na.rm = TRUE)
nedap_coverage_end <- max(attentions_resolved$attention_date, na.rm = TRUE)
```

```{r}
#| label: platform-enrollment-data


# Nedap runs two separate platform instances at this farm - "Activity"
# (used by farm staff daily) and "SmartSight" (mobility monitoring,
# generates NEDLAME) - kept separate at the farm's request to avoid
# confusion when checking the Nedap platform. A cow absent from
# SmartSight's animal list was never being monitored for lameness at
# all, regardless of what the raw attentions export's date coverage
# shows - this check is animal-level enrollment, not date-bound, so it
# can resolve cases the date-bound attentions export above can't (see
# the Platform Enrollment Check subsection in Q4).
cows_activity <- read_csv("data/cows_activity.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

cows_smartsight <- read_csv("data/cows_smartsight.csv", show_col_types = FALSE) |>
  transmute(id = as.character(Animal))

# Renumber-aware enrollment - see fxn_build_enrolled_ids for why the current
# animal number alone is not enough.
invisible(list2env(fxn_build_enrolled_ids(events_all, cows_smartsight), environment()))
```

# Overview

To determine if using the camera benefited cows we set up a study that trimmed 1/2 the cows that the camera flagged and left the other group to be trimmed if your farm staff found them or if they showed up on the trim list.

- Nedap flags cows in 3 ways and the setup in DairyComp section describes the flags and how this was handled in DairyComp

To compare the groups we looked at the following outcomes after being flagged:

- Milk Production

- Culling

- Cows found with a lesion - in either group or neither - where the camera had not flagged her first

## Setup in DairyComp

For this study the follow things were created in DairyComp to handle the file data from NEDAP daily.

- A command to clear all filein items before new data was brought in as a workaround to deal with historical flag issues

- New NEDAP lame event with 3 protocols to flag cows and criteria to create them

  - Low (score \<30 today, not trimmed or an alarm in last 90 days not a cow with history of upper leg lameness)

    - ENTER EC=70\\P215 FOR LFTYP=30102 LMFLG=1 LMSV=1-30 DSNLM=90-1 FTDAT\<-90 UPLAM\<\>1

  - Low Declining (less than 30 score, declining score and no alarm in last 7 days and not trimmed in last 28 days)

    - ENTER EC=70\\P214 FOR LFTYP=30101 LMFLG=1 DSNLM=7-1 FTDAT\<-28

  - Strong Decline (any score and a strong decline in scores with no alarm in last 7 days, not trimmed in last 28 days) second OR statement allows cows with a low alert already to get another alert

    - ENTER EC=70\\P213 FOR LFTYP=30100 LMFLG=1 (DSNLM=7-1 FTDAT\<-28)(FTDAT\<-28 %73.NDLAME.101\>LOW)

- Cows were added to trim list with a command that uses an existing item (MNFRS left over from Luciano's work) that randomizes cows by lactation group.

  - Enter FTDAT=1.1.1 for MNFRS=1 %.70NEDLAME.101=TODAY

Above commands were all run daily around 5 am after NEDAP file-in occurred.

## Results

- `r nrow(first_nedlame)` cows received a `NEDLAME` alert between `r format(min(first_nedlame$first_nedlame_date))` and `r format(max(first_nedlame$first_nedlame_date))`

- `r n_excluded_post_trim` were excluded as alerts occurred after a recent hoof trimming.

  - These were still coded in DairyComp due to the delay in importing hoof trimming data a few days after the actual hoof trimming date.

This results in

- `r n_tx` cows enrolled in the **TX (Trim)** arm (`MNFRS = 1`, automatically trimmed after alert)

- `r n_control` in the **Control** arm (`MNFRS = 2`, staff decide whether/when to act or trimmed because of regular trim list),

# Description of Cows

The tables below describe the cows enrolled by treatment group

## By Alert Type

::: panel-tabset
```{r}
#| label: overview-table

cohort |>
  count(tx_group, alert_type) |>
  pivot_wider(names_from = alert_type, values_from = n, values_fill = 0) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Most cows were flagged as Low Declines") |>
  tab_stubhead(label = "Treatment Group")
```

## By Lactation Number

```{r}
#| label: lact-group-table

cohort |>
  mutate(lact_group_display = case_when(
    lact_number == 1 ~ "1",
    lact_number == 2 ~ "2",
    lact_number >= 3 ~ "3+",
    TRUE ~ NA_character_
  )) |>
  count(tx_group, lact_group_display) |>
  pivot_wider(names_from = lact_group_display, values_from = n, values_fill = 0) |>
  select(tx_group, `1`, `2`, `3+`) |>
  mutate(Total = rowSums(across(where(is.numeric)))) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Lact 3+ was the largest group enrolled",
    ) |>
  tab_stubhead(label = "Treatment Group") |>
  cols_label(`3+` = "3+")
```

## By Lactation Stage (DIM)

```{r}
#| label: dim-balance-table

cohort |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_dim = round(mean(dim_at_nedlame, na.rm = TRUE), 0),
    median_dim = round(median(dim_at_nedlame, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Similar DIM between groups",
    ) |>
  cols_label(n = "N Cows", mean_dim = "Mean DIM", median_dim = "Median DIM")
```

```{r}
#| label: q1-dim-check
#| fig-height: 4

cohort |>
  ggplot(aes(x = tx_group, y = dim_at_nedlame, fill = tx_group)) +
  geom_violin(alpha = 0.5, color = NA, trim = FALSE) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.4, size = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "DIM at 1st NEDLAME Alert")
```
:::

# Trimming

## Time to next trimming after alert

These curves are about **protocol compliance**. They show how quickly a cow was trimmed after her first NEDLAME alert. 
In the TX group the trim is triggered by the alert by design, so the curve measures how well the protocol was followed. In the Control group nothing is triggered, so the same curve shows when those cows got trimmed anyway - whether staff picked them out or the routine round came round to them.

The figures below show the time to the next trimming event split by treatment group and alerts types.

- It is clear that most of the TRIM treatment cows got trimmed within 7-10 days and about 20% of the cows in the Control group get trimmed within that same time period.

```{r}
#| label: q2-build

invisible(list2env(fxn_build_q2(cohort, events_formatted, lame_data), environment()))

# Control-arm trim drivers, shared with the full report so both quote the same
# split between staff-flagged and routine trims.
invisible(list2env(fxn_build_control_trim_drivers(
  next_events, cohort, events_formatted, lame_data), environment()))
```

::: panel-tabset
## Time to Any Trim/LAME Event

```{r}
#| label: q2-any-plot
#| fig-height: 5

km_any <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_any)

km_any |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-any-table

km_any |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Trim/LAME Event")
```

## By Alert Type

Decline cows mostly all get trimmed

```{r}
#| label: q2-alert-type-plot
#| fig-height: 6

alert_levels <- c("LOW", "LOW DECLINE", "STRONG DECLINE")

# ggsurvfit only exposes a single combined `strata` column, so a KM plot
# faceted by alert_type is built manually: fit one survfit2() per
# alert level (skipping any level without both treatment arms
# represented) and stack the tidy() estimates before plotting.
fittable_alert_levels <- alert_levels |>
  keep(\(a) {
    d <- q2_any |> filter(alert_type == a)
    n_distinct(d$tx_group) == 2 && nrow(d) >= 10
  })

q2_alert_type_km_data <- fittable_alert_levels |>
  map(\(a) {
    d <- q2_any |> filter(alert_type == a)
    fit <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = d)
    broom::tidy(fit) |>
      mutate(
        tx_group = str_remove(strata, "^tx_group="),
        alert_type = a,
        risk = 1 - estimate,
        # survival's CI bounds flip when converted to risk (1 - survival):
        # the upper bound on survival becomes the LOWER bound on risk.
        risk_low = 1 - conf.high,
        risk_high = 1 - conf.low
      )
  }) |>
  list_rbind() |>
  mutate(alert_type = factor(alert_type, levels = alert_levels))

skipped_alert_levels <- setdiff(alert_levels, fittable_alert_levels)

ggplot(q2_alert_type_km_data, aes(x = time, y = risk, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = risk_low, ymax = risk_high), alpha = 0.15, color = NA) +
  geom_step(linewidth = 1) +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event", color = "Treatment Group", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```

`r if (length(skipped_alert_levels) > 0) paste0("*Alert type(s) omitted for insufficient data in one or both arms: ", paste(skipped_alert_levels, collapse = ", "), ".*")`
:::

- Only about half the Control cows (`r n_ctl_trimmed` of `r sum(cohort$tx_group == 'Control')`) were trimmed at all during the study.

### Why did the Control cows get trimmed?

A Control cow is only trimmed if someone decides to act - either staff spotted her and put her on the list, or she came up on the routine round. That split matters, because it decides what the camera is really being compared against.

```{r}
#| label: tbl-farm-trim-driver

control_trims |>
  group_by(Driver) |>
  summarize(`Cows` = n(),
            `% of trims` = round(100 * n() / n_ctl_trimmed, 1),
            `% finding a lesion` = round(100 * mean(found_lesion), 0),
            `Median days alert to trim` = median(days_to_trim),
            .groups = 'drop') |>
  gt() |>
  tab_header(title = 'Most Control trims were routine, and most of those found nothing')
```

- Only `r n_ctl_staff` of the `r n_ctl_trimmed` (`r pct_ctl_staff`%) had a staff `CHKLAME` before the trim. The other `r n_ctl_routine` were the routine round.
- The two are not doing the same job: staff-flagged trims found something `r pct_lesion_staff`% of the time against `r pct_lesion_routine`% for routine ones, and happened a median `r med_days_staff` days after the alert rather than `r med_days_routine`.
- So the Control group is mostly the routine schedule catching up, not staff acting on cows. Worth holding onto when reading the comparison: the camera is being measured against a mixture, and only about a third of it is an active decision.

## How many trims with no lesions?

The table below shows the number of cows who had a first NEDAP alert and were trimmed **but** did not have a lesion.

**This table's population is only the number of cows in the `r nrow(cohort)`-cow NEDLAME cohort that were trimmed.**

```{r}
#| label: q2-trimonly-summary

q2_first_event_detail <- q2_any |>
  filter(event_occurred == 1) |>
  left_join(lame_data |> select(id_animal, lact_number, date_event, event, trimonly),
            by = c("id_animal", "lact_number", "date_event"))

q2_trimonly_tab <- q2_first_event_detail |>
  group_by(tx_group) |>
  summarize(n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100, .groups = 'drop') |>
  bind_rows(
    q2_first_event_detail |> summarize(tx_group = 'Overall', n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100)
  )

# Title built from the cells beneath it. It used to be typed by hand and had
# already gone stale against a rebuild, quoting a range the table contradicted.
# level is 'TX (Trim)', not 'TX' - matching on 'TX' silently gave numeric(0)
tx_lvl <- setdiff(q2_trimonly_tab$tx_group, c('Control', 'Overall'))
stopifnot(length(tx_lvl) == 1)
pct_trimonly_tx  <- round(q2_trimonly_tab$pct_trimonly[q2_trimonly_tab$tx_group == tx_lvl], 1)
pct_trimonly_ctl <- round(q2_trimonly_tab$pct_trimonly[q2_trimonly_tab$tx_group == 'Control'], 1)
trimonly_title <- sprintf('%d-%d%% of Trims do not have a lesion',
                          floor(min(q2_trimonly_tab$pct_trimonly)), ceiling(max(q2_trimonly_tab$pct_trimonly)))

q2_trimonly_tab |>
  gt(rowname_col = 'tx_group') |>
  fmt_number(columns = pct_trimonly, decimals = 1) |>
  cols_label(n = 'N Cows with Trim', n_trimonly = 'No Lesion', pct_trimonly = '% Trim-Only') |>
  tab_header(title = trimonly_title)
```

Just over half the trims done straight off a camera alert found no lesion (`r pct_trimonly_tx`%), against `r pct_trimonly_ctl`% in the control group. But the control trims are mostly the routine schedule catching up rather than staff picking cows out - only `r n_ctl_staff` of the `r n_ctl_trimmed` control cows trimmed after an alert had a `CHKLAME` first. When staff did flag a cow themselves, `r pct_lesion_staff`% of those trims found something, against `r pct_lesion_routine`% for the routine ones. So this is not a camera-versus-staff comparison; it is the camera against a mixture that is mostly routine.

# How does the Farm staff do?

The `CHKLAME` event is used to flag cows in the need of trimming by farm staff.

The table below compares what % of cows with a NEDAP alert were flagged by farms staff \<=7 days of the NEDAP alert.

Overall farm staff flags about 20% of cows overall with more cows with low scores or strong declines being flagged.

```{r}
#| label: q3-build

invisible(list2env(fxn_build_q3(cohort, events_formatted), environment()))
```

::: panel-tabset
## By Treatment Group

```{r}
#| label: q3-treatment-group-plot
#| fig-height: 4

staff_id |>
  group_by(tx_group) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = tx_group, y = pct_staff_id, fill = tx_group)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "% Staff-Identified (CHKLAME)")
```

## By Alert Type

Staff catches a bigger % of Low and Strong Decline alerts

```{r}
#| label: q3-alert-type-by-group-plot
#| fig-height: 4

staff_id |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = alert_type, y = pct_staff_id, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(sprintf("%.1f", pct_staff_id), "%\n(n=", n, ")")),
            position = position_dodge(width = 0.7), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  labs(x = "Alert Type", y = "% Staff-Identified (CHKLAME)", fill = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Milk Production After 1st NEDLAME Alert

```{r}
#| label: q1-build

invisible(list2env(fxn_build_q1(cohort, milk_resolved, params), environment()))
```

The figures below compare milk production between the treatment groups. Cows were assigned to TX or Control at enrollment, so this is a like-for-like comparison rather than a descriptive one - and because the groups were randomised, seeing no difference here is itself evidence against a large effect, not just a lack of adjustment.

- `r n_milk_cows` of the `r nrow(cohort)` cohort cows had at least one milk record after their alert and a usable pre-alert baseline. Fewer cows sit behind the right-hand end of these curves than the left: a cow needs `r params$milk_window_post` days of milk records after her alert to appear at day `r params$milk_window_post`, and only `r n_milk_day30` have that so far. So the day-`r params$milk_window_post` end is not the same cows as the day-0 end - it is the ones alerted earliest in the study. That end of the curve can still move as the rest catch up, rather than just firm up around where it sits now.

Overall numerically there is an advantage to the trimmed group in milk production after an alert but there is a lot of variation (the shaded part around the solid line) so this is not something to put a lot of trust in.

**NOTE:** *Due to the fact that cows were at very different lactation stages when flagged milk production is not directly comparable across cows. Each cow's production is normalized to **% of her own pre-alert baseline** (mean production over the `r params$milk_baseline_days` days before the alert) before averaging by day and treatment group. This allow us to isolates the shape of the trend/recovery from each cow's absolute production level.*

::: panel-tabset
## By Treatment Group

```{r}
#| label: q1-plot
#| fig-height: 5

ggplot(trend, aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    x = "Days Relative to First NEDLAME Alert",
    y = "Milk Production (% of Own Pre-Alert Baseline)",
    color = "Treatment Group", fill = "Treatment Group"
  )
```

## By Alert Type

Does the milk trend differ by the type of the Nedap alert?

```{r}
#| label: q1-alert-type-plot
#| fig-height: 5

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```
:::

# Time to Culling After NEDAP Enrollment

Time to culling is compared between the treatment groups below.

```{r}
#| label: q6-build

invisible(list2env(fxn_build_q6(cohort, date_max_pull), environment()))

# Quoted in the caveat below. The curves used to be summarised as 'no real
# impact', which claims a demonstrated absence of effect - with this few
# events the honest statement is that the study cannot tell either way.
n_culls <- sum(q6$culled_ever)
n_culls_tx  <- sum(q6$culled_ever[q6$tx_group != 'Control'])
n_culls_ctl <- sum(q6$culled_ever[q6$tx_group == 'Control'])
```

```{r}
#| label: q6-plot
#| fig-height: 5

km_cull |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% Culled") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q6-table

km_cull |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% Culled After First NEDLAME Alert")
```

The curves sit close together, but only `r n_culls` cows were culled in total (`r n_culls_tx` TX, `r n_culls_ctl` Control). That is far too few to tell whether trimming off the alert changes culling - this is not evidence that it makes no difference, it is a question the study is too small and too short to answer.

# What cows were missed by the system?

This section looks at cows with lesion regardless of treatment group.

Among cows with a `LAME`-diagnosed lesion on or after `r params$history_after_date`, this checks whether that cow had a `NEDLAME` alert in the preceding lookback window (same lactation).

A "miss" means staff/routine trimming found a lesion that NEDAP hadn't already flagged.

**Note** *Lame events that were in heifers, dried off cows or cows trimmed within 28 days of another trim were excluded as the system does not monitor those cows or they were automatically excluded*.

```{r}
#| label: q4-build

invisible(list2env(fxn_build_q4(
  cohort = cohort, all_nedlame_cows = all_nedlame_cows, lame_data = lame_data,
  events_formatted = events_formatted, events_all = events_all,
  animal_lactations = animal_lactations, enrolled_ids = enrolled_ids,
  nedlame_start_date = nedlame_start_date, params = params), environment()))
```

- A total of `r nrow(q4)` cows had a lesion diagnosed after the project start date (`r params$history_after_date`).

  - Of those `r nrow(q4)` cows, **`r n_misses` (`r round(100 * n_misses / nrow(q4), 1)`%)** had no camera alert in DairyComp in the `r params$q4_lookback_days` days before the lesion was found.

  - Another `r n_warned_not_acted` are left out of that count on purpose. The camera *had* flagged them, further back than `r params$q4_lookback_days` days - a median of `r round(median(q4$days_alert_to_lesion[q4$warned_not_acted]))` days - and they were not trimmed between that alert and the lesion. The warning came and nothing was done with it, so it would be wrong to score that against the camera. Most of them are Control cows, where nobody was asked to act on an alert.

  - Two things to hold against that number. Cows trimmed in the `r params$recent_trim_suppression_days` days before their lesion are already left out (`r n_excluded_recent_trim` of them), because the DairyComp set-up could not have alerted on them - so this is not counting cases the system was never allowed to flag. But `r n_truncated_lookback` of the `r nrow(q4)` had less than the full `r params$q4_lookback_days` days available, because their lesion fell close to the camera going live. Those cows had a shorter run at being flagged, and they are still counted as misses.

## What Lesions are being missed?

The figure below shows the lesion types found at these first-diagnosis cases, broken out by NEDAP detection status:

- **Flagged, caught in time** - an alert reached DairyComp in the days before the lesion. The camera got there first.
- **Alerted earlier, no action taken** - the camera flagged her further back than `r params$q4_lookback_days` days and she was not trimmed in between. Not scored against the camera.
- **Flagged, but no alert in time** - the camera did see something, but nothing reached DairyComp early enough to act on.
- **Never flagged** - the camera never picked her up at all.


```{r}
#| label: q4-lesion-types

# The same detection split the full report uses: did the camera flag her, and
# did that flag reach DairyComp. This replaces an earlier labelling that split
# on whether she was ever in the pilot at all, which answered a different
# question and made the figure hard to read.
q4_sensor <- q4_group |>
  mutate(in_coverage = date_event <= nedap_coverage_end)

sensor_hits <- q4_sensor |>
  filter(in_coverage) |>
  select(id_animal, lact_number, date_event, lookback_days_used) |>
  left_join(attentions_resolved |> select(id_animal, lact_number, attention_date),
            by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(sensor_hit = !is.na(attention_date) & attention_date < date_event &
           attention_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event) |>
  summarize(sensor_fired_in_window = any(sensor_hit), .groups = "drop")

q4_sensor <- q4_sensor |>
  left_join(sensor_hits, by = c("id_animal", "lact_number", "date_event")) |>
  mutate(detection_group_3way = factor(case_when(
    !in_coverage           ~ "No Data (Outside Coverage)",
    caught_by_nedap        ~ "Flagged, Caught in Time",
    warned_not_acted       ~ "Alerted earlier, no action taken",
    sensor_fired_in_window ~ "Flagged, but no alert in time",
    TRUE                   ~ "Never Flagged (True Miss)"
  ), levels = c("Flagged, Caught in Time", "Alerted earlier, no action taken", "Flagged, but no alert in time",
                "Never Flagged (True Miss)", "No Data (Outside Coverage)")))

invisible(list2env(fxn_build_q4_groups(q4_sensor, lame_lesion_recent), environment()))

# Lesion mix of what the camera never flagged vs what it did flag. Quoted in
# the prose under the figure. Shares are within each group, so they answer
# 'what does a miss tend to be' rather than 'how often is DD missed'.
checkable <- q4_sensor |>
  filter(in_coverage) |>
  mutate(camera = if_else(detection_group_3way == 'Never Flagged (True Miss)',
                          'Camera never flagged', 'Camera flagged her'))
n_cam_miss <- sum(checkable$camera == 'Camera never flagged')
n_cam_flag <- sum(checkable$camera == 'Camera flagged her')

lesion_by_camera <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(checkable |> select(id_animal, lact_number, date_event, camera),
             by = c('id_animal', 'lact_number', 'date_event')) |>
  pivot_longer(all_of(lesion_type_cols), names_to = 'lesion_type', values_to = 'has_lesion') |>
  filter(has_lesion == 1) |>
  count(camera, lesion_type) |>
  mutate(pct = if_else(camera == 'Camera never flagged', 100 * n / n_cam_miss, 100 * n / n_cam_flag))

pctlc <- function(lt, cam) round(lesion_by_camera$pct[lesion_by_camera$lesion_type == lt & lesion_by_camera$camera == cam], 0)
pct_dd_miss  <- pctlc('dd',  'Camera never flagged')
pct_dd_flag  <- pctlc('dd',  'Camera flagged her')
pct_wld_miss <- pctlc('wld', 'Camera never flagged')
pct_wld_flag <- pctlc('wld', 'Camera flagged her')

ggplot(q4_lesion_types, aes(x = lesion_type, y = pct, fill = detection_group_3way)) +
  geom_col(width = 0.7) +
  geom_text(data = lesion_totals, inherit.aes = FALSE,
            aes(x = lesion_type, y = 101, label = paste0("n=", total_cases)),
            hjust = 0, size = 3, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 100, 25), labels = function(x) paste0(x, "%"),
                     limits = c(0, 118), expand = expansion(mult = c(0, 0))) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "% of cases with that lesion", fill = NULL) +
  theme(legend.position = "bottom")
```

Comparing what the camera missed against what it caught, digital dermatitis stands out: it is `r pct_dd_miss`% of the cases the camera never flagged against `r pct_dd_flag`% of the ones it did. White line runs the other way — `r pct_wld_miss`% against `r pct_wld_flag`% — so the camera picks that up well. "Other" on this farm is typically corkscrew claw.

None of that is a surprise. Locomotion scoring is known to detect DD poorly, and the camera is scoring locomotion, so a lesion that barely changes how a cow walks is the one it will tend to miss.

## Of the Misses, Did Farm Staff Catch Them Anyway?

```{r}
#| label: q4-staff-catch-build

invisible(list2env(fxn_build_staff_catch(q4, chklame), environment()))

# How many of the misses the camera genuinely never saw, as opposed to flagged
# without the alert arriving in time. Without this the miss count reads as the
# camera's own failure rate, which it is not - a good share of those cows were
# flagged and the alert simply did not land in the window.
n_never_flagged_farm <- sum(q4_sensor$detection_group_3way == 'Never Flagged (True Miss)')
n_flagged_no_alert   <- sum(q4_sensor$detection_group_3way == 'Flagged, but no alert in time')
n_outside_coverage   <- sum(q4_sensor$detection_group_3way == 'No Data (Outside Coverage)')
# the three buckets are exactly the misses - if this ever fails the prose below
# is quoting a split that does not add up
n_warned_earlier     <- sum(q4_sensor$detection_group_3way == 'Alerted earlier, no action taken')
# the buckets must partition the lesion cases exactly; a mistyped level would
# otherwise drop cows silently out of every count and plot on this page
stopifnot(n_never_flagged_farm + n_flagged_no_alert + n_outside_coverage == nrow(staff_catch),
          nrow(staff_catch) == n_misses,
          n_warned_earlier == n_warned_not_acted,
          !any(is.na(q4_sensor$detection_group_3way)))

# Why an alert did not land. Same four rules the full report uses, from the
# shared function, so the farm and the detailed report cannot disagree.
invisible(list2env(fxn_build_pipeline_by_design(
  q4_sensor, attentions_resolved, events_formatted, events_all, lame_data), environment()))
```

Of the `r nrow(staff_catch)` cows whose lesion had no `NEDLAME` alert recorded in DairyComp beforehand (they were trimmed on the routine round or because staff picked them out),

Worth being precise about what that number counts. It is cows with **no alert in DairyComp in time**, which is not the same as cows the camera failed to see. Splitting them:

- **`r n_never_flagged_farm` the camera never flagged at all.** No sign of a problem before the lesion was found. This is the real blind spot, and it is largely the DD cases above.

- **`r n_flagged_no_alert` the camera did flag, but no alert arrived in time.** Most of these are the system doing what it was set up to do: `r n_pipeline_loss_by_design` were deliberately held back, because she had been trimmed recently or had already been alerted and a repeat would only clutter the cowcard. Another `r n_pipeline_loss_unknowable` cannot be judged either way from the data available. That leaves **`r n_pipeline_loss_genuine` where a flag went missing between the camera and DairyComp** - a small number, and one for Nedap to chase rather than the farm.

`r if (n_outside_coverage > 0) paste0('- **', n_outside_coverage, ' fall outside the period we have camera data for**, so nothing can be said about them.')`

So the number that matters to the farm is the `r n_never_flagged_farm` the camera never saw. Getting alerts from the camera into DairyComp is largely working.

- `r sum(staff_catch$staff_flagged)` (`r round(100 * mean(staff_catch$staff_flagged), 1)`%) were flagged by the farm staff in the 7 days before diagnosis.

# **The bottom line**

- NEDAP camera seems to perform better than CattleEye did in the past.

- Cows with DD are being missed by the camera

- Where the camera gave no warning, it was mostly because it never flagged the cow, not because alerts were going missing on the way into DairyComp.

- NEDAP camera catches more cows than just farm staff/routine trimming

- Trimming a flagged cow straight away, rather than leaving it to staff or the routine round, did not produce a clear difference in milk or culling over the first 30 days. The milk curves slightly favour the trimmed group, but cows vary far more than the groups do, and there were too few culls either way to read anything into.

- The study was not long enough to determine if more cows being caught by camera do better long term as in less repeat lesions.
