---
title: "NEDLAME Alert Pilot: TX (Trim) vs Control"
editor: source
format:
 html:
   embed-resources: true
   toc: true
   toc-location: left
   code-fold: true
execute:
    echo: false
    message: false
    warning: false
params:
  milk_window_pre: 14
  milk_window_post: 30
  milk_baseline_days: 7
  q4_lookback_days: 21
  history_after_date: "2026-06-01"
---

```{r}
#| label: setup

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  arrow,
  lubridate,
  gt,
  gtsummary,
  survival,
  survminer,
  ggsurvfit,
  scales,
  viridis
)

source(file.path('functions/fxn_read_milk_folder.R'))

# GC's lame functions (fxn_collapse_lesions, fxn_trim_vars, fxn_dz_status) -
# fetched live from GitHub, same mechanism report_explore_lame_new.qmd uses.
# Two local overrides are sourced AFTER fxn_load_os_fxns() specifically so
# they win over the GitHub versions in the global environment:
#  - functions/fxn_code_lesions.R: the GitHub version's `trimonly` logic
#    only checks the *protocols* field for "NONE" - this farm's
#    routine/no-lesion trims are coded "NONE" in the *remark* field
#    instead, which the GitHub version doesn't check, so it silently
#    misclassifies every routine trim as having a lesion.
#  - functions/fxn_collapse_lesions.R: the GitHub version reassigns its
#    own `data` variable to an already-one-row-per-date subset before
#    computing the max() union across that date's lesion columns, so
#    when a cow has multiple event rows on the same date coding
#    DIFFERENT lesions, only whichever row survives the earlier
#    reduction keeps its lesion flags - the rest are silently dropped
#    (confirmed against a real cow with 4 same-day LAME rows: only 1 of
#    her 3 distinct lesion types survived under the GitHub version).
source(file.path('functions/fxn_load_os_fxns.R'))
fxn_load_os_fxns()
source(file.path('functions/fxn_code_lesions.R'))
source(file.path('functions/fxn_collapse_lesions.R'))

events_all <- read_parquet('data/intermediate_files/events_all_columns.parquet')
events_formatted <- read_parquet('data/intermediate_files/events_formatted.parquet')
animals <- read_parquet('data/intermediate_files/animals.parquet')
animal_lactations <- read_parquet('data/intermediate_files/animal_lactations.parquet')

date_max_pull <- max(events_all$date_event, na.rm = TRUE)
farm_name <- unique(events_all$location_event)

# NEDLAME go-live date, per the DairyComp setup: enrollment protocols were
# only turned on starting ~June 1 2026, so no NEDLAME alert can exist
# before this regardless of how "missed" a case looks.
nedlame_start_date <- min(events_all$date_event[events_all$Event == "NEDLAME"], na.rm = TRUE)
```

```{r}
#| label: cohort

# One row per cow-lactation that ever got a NEDLAME alert. Treatment group
# is read off the MNFRS value on the FIRST NEDLAME event for that cow -
# MNFRS is stable across a cow's NEDLAME rows even though it drifts over
# her broader event history (0 before enrollment, 1/2 once enrolled).
first_nedlame <- events_all |>
  filter(Event == "NEDLAME") |>
  group_by(id_animal) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  transmute(
    id_animal, id, lact_number,
    first_nedlame_date = date_event,
    alert_type = Remark,
    tx_group = case_when(
      MNFRS == "1" ~ "TX (Trim)",
      MNFRS == "2" ~ "Control",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(tx_group))

# Attach lactation-stage context (date_fresh/date_dry) so every downstream
# join can be scoped to id_animal + lact_number, never crossing into a
# different lactation for the same cow.
cohort <- first_nedlame |>
  left_join(
    animal_lactations |>
      select(id_animal, lact_number, date_fresh, date_dry, date_next_fresh),
    by = c("id_animal", "lact_number")
  ) |>
  mutate(
    dim_at_nedlame = as.numeric(first_nedlame_date - date_fresh),
    date_lact_end = pmin(
      coalesce(date_dry, date_next_fresh, date_max_pull),
      date_max_pull,
      na.rm = TRUE
    )
  )

n_tx <- sum(cohort$tx_group == "TX (Trim)")
n_control <- sum(cohort$tx_group == "Control")
```

```{r}
#| label: lesion-history-dataset

# Same lesion-coding / chronicity pipeline as report_explore_lame_new.qmd,
# just kept at the lactation grain (lact_number carried through) so the
# "history at enrollment" lookup below never crosses lactations.
all_lesions <- c("dd", "footrot", 'wld', "sole_ulcer", "injury",
                  "cork", "other", "hemorrhage",
                  "sole_fracture", "toe_ulcer", "thin",
                  "inf", "noninf", "toe", "axial", "lesion")
diseases <- c("inf", "noninf", "lesion")

lame_data <- events_formatted |>
  select(id_animal, lact_number, date_event, event,
         remark_letters1, protocols_letters1, location_event) |>
  filter(event %in% c("LAME", "TRIM", "FOOTRIM")) |>
  fxn_code_lesions(protocol_var = protocols_letters1, remark_var = remark_letters1) |>
  filter(!is.na(trimonly)) |>
  mutate(
    noninf = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 1,
      .default = noninf),
    other = case_when(
      other == 1 & str_detect(protocols_letters1, "Ab") ~ 0,
      .default = other)
  ) |>
  fxn_collapse_lesions(lesions = all_lesions) |>
  fxn_trim_vars(location_event, id_animal, date_var = date_event, trimonly_var = trimonly) |>
  fxn_dz_status(disease_cols = diseases, event_filter = trimonly == 0)

# Each NEDLAME cow's lameness-history label as of (at/before) her first
# NEDLAME date, using her FULL lifetime record (all lactations - lame_data
# goes back to 2021), not scoped to her current lactation. Unlike every
# other join in this report, this one is intentionally NOT lactation-scoped
# per explicit user direction: chronicity is a lifetime concept, and the
# ~5 years of history available should be used rather than artificially
# resetting it at each freshening.
#
# IMPORTANT: candidate history rows are filtered to "at/before enrollment"
# BEFORE joining to cohort, not after (same reasoning as fxn_time_to_next()
# above) - filtering after a left_join would silently DROP a cow from
# `cohort` entirely instead of correctly labeling her. An earlier version
# of this exact chunk had that bug and was quietly dropping 49 of 434 cows
# (~11%) from the ENTIRE cohort - not just from Q5, but from every section,
# since `cohort` is the shared base table.
history_lookup <- lame_data |>
  select(id_animal, date_event, status_lesion) |>
  filter(!is.na(status_lesion))

valid_history <- history_lookup |>
  inner_join(
    cohort |> select(id_animal, lact_number, first_nedlame_date),
    by = "id_animal"
  ) |>
  filter(date_event <= first_nedlame_date) |>
  select(id_animal, lact_number, date_event, status_lesion)

# A cow with literally no LAME/TRIM/FOOTRIM diagnosis anywhere in her life
# before her NEDLAME alert (still ~37% of the cohort even using the full
# 5-year history) has no status_lesion to inherit. Rather than invent a
# 4th bucket for this, she's folded into "New" - the same label her
# eventual first-ever diagnosis would get, and the closest existing
# category to "no established chronic pattern."
cohort <- cohort |>
  left_join(valid_history, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  group_by(id_animal, lact_number) |>
  slice_max(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
  ungroup() |>
  mutate(history_group = if_else(is.na(status_lesion), "New", status_lesion)) |>
  select(-date_event, -status_lesion)

history_levels <- c("New", "Repeat", "Chronic")
cohort <- cohort |>
  mutate(history_group = factor(history_group, levels = history_levels))
```

```{r}
#| label: milk-data

# data/milk/ is not wired into any existing pipeline script. Raw milk rows
# only carry a raw `id` (not the composite id_animal), so they're resolved
# to a specific animal-lactation by nearest-preceding date_fresh (see
# fxn_read_milk_folder.R for why a naive date_dry window isn't safe here).
milk_long <- fxn_read_milk_folder()
milk_resolved <- fxn_resolve_milk_to_lactation(milk_long, animal_lactations)
```

# Overview

`r n_tx` cows were enrolled in the **TX (Trim)** arm (`MNFRS = 1`, automatically trimmed after alert) and `r n_control` in the **Control** arm (`MNFRS = 2`, staff decide whether/when to act), based on `r nrow(first_nedlame)` cows that ever received a `NEDLAME` alert between `r format(min(first_nedlame$first_nedlame_date))` and `r format(max(first_nedlame$first_nedlame_date))`. This is a short (~5 week) history since NEDLAME is a newly-added event stream, which caps every time-window analysis below.

**A methodology note that applies throughout**: wherever this report looks for "the" trim, LAME diagnosis, staff check, or NEDLAME alert for a cow, it always uses her **first** qualifying occurrence in the relevant window, not every occurrence — a cow flagged or diagnosed more than once doesn't get counted multiple times. When a single visit/date has more than one distinct lesion type coded (e.g. white line disease *and* digital dermatitis found at the same exam), each distinct type is kept; the same lesion type coded more than once on one day only counts once (see the note in Q4 for the specific chart this matters most for).

```{r}
#| label: overview-table

cohort |>
  count(tx_group, alert_type) |>
  pivot_wider(names_from = alert_type, values_from = n, values_fill = 0) |>
  gt(rowname_col = "tx_group") |>
  tab_header(title = "Cows per Treatment Group by Alert Type") |>
  tab_stubhead(label = "Treatment Group")
```

```{r}
#| label: dim-balance-table

cohort |>
  group_by(tx_group) |>
  summarize(
    n = n(),
    mean_dim = round(mean(dim_at_nedlame, na.rm = TRUE), 0),
    median_dim = round(median(dim_at_nedlame, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  gt(rowname_col = "tx_group") |>
  tab_header(
    title = "Lactation Stage at Enrollment",
    subtitle = "Days in milk (DIM) when first NEDLAME alert fired"
  ) |>
  cols_label(n = "N Cows", mean_dim = "Mean DIM", median_dim = "Median DIM")
```

# Q1: Milk Production Trend After First NEDLAME Alert

Cows are at very different lactation stages when flagged, so raw milk pounds aren't directly comparable across cows. Each cow's series is normalized to **% of her own pre-alert baseline** (mean production over the `r params$milk_baseline_days` days before the alert) before averaging by day and treatment group — this isolates the *shape* of the trend/recovery from each cow's absolute production level.

```{r}
#| label: q1-build

milk_cohort <- cohort |>
  inner_join(
    milk_resolved |> select(id_animal, lact_number, milk_date, dmlk1),
    by = c("id_animal", "lact_number"),
    relationship = "many-to-many"
  ) |>
  mutate(day_rel = as.numeric(milk_date - first_nedlame_date)) |>
  filter(day_rel >= -params$milk_window_pre, day_rel <= params$milk_window_post)

baseline <- milk_cohort |>
  filter(day_rel >= -params$milk_baseline_days, day_rel <= -1) |>
  group_by(id_animal, lact_number) |>
  summarize(baseline_dmlk1 = mean(dmlk1, na.rm = TRUE), .groups = "drop")

milk_norm <- milk_cohort |>
  inner_join(baseline, by = c("id_animal", "lact_number")) |>
  filter(baseline_dmlk1 >= 5) |> # drop the rare near-zero baseline (unreliable % denominator)
  mutate(pct_of_baseline = dmlk1 / baseline_dmlk1 * 100)

n_milk_cows <- n_distinct(milk_norm$id_animal)

trend <- milk_norm |>
  group_by(tx_group, day_rel) |>
  summarize(
    mean_pct = mean(pct_of_baseline, na.rm = TRUE),
    se_pct = sd(pct_of_baseline, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )
```

`r n_milk_cows` of the `r nrow(cohort)` cohort cows had at least one resolvable milk record in this window and a usable pre-alert baseline.

```{r}
#| label: q1-plot
#| fig-height: 5

ggplot(trend, aes(x = day_rel, y = mean_pct, color = tx_group, fill = tx_group)) +
  geom_ribbon(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(
    x = "Days Relative to First NEDLAME Alert",
    y = "Milk Production (% of Own Pre-Alert Baseline)",
    color = "Treatment Group", fill = "Treatment Group"
  )
```

## Milk Trend by Alert Type

Does the milk trend differ by how severe the Nedap alert was (`Strong Decline` vs `Low Decline` vs `Low`)?

```{r}
#| label: q1-alert-type-plot
#| fig-height: 5

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

## Lactation-Stage Balance

```{r}
#| label: q1-dim-check

cohort |>
  filter(id_animal %in% milk_norm$id_animal) |>
  ggplot(aes(x = tx_group, y = dim_at_nedlame, fill = tx_group)) +
  geom_violin(alpha = 0.5, color = NA, trim = FALSE) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.4, size = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = "none") +
  labs(x = NULL, y = "Days in Milk at First NEDLAME Alert")
```

# Q2: Time to Next LAME/FOOTRIM/TRIM Event

For each cow, we look for the first `LAME`, `FOOTRIM`, or `TRIM` event strictly after her first NEDLAME alert, **within the same lactation** — censored at that lactation's dry-off date or the data pull date, whichever comes first. Because MNFRS=1 cows are automatically trimmed, "time to any trim" mechanically favors TX; a second version restricts to events where a lesion was actually identified, to separate "got trimmed on schedule" from "a problem was actually found."

```{r}
#| label: q2-build

next_events <- events_formatted |>
  filter(event %in% c("LAME", "FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event, event) |>
  distinct()

lesion_lookup <- lame_data |> select(id_animal, lact_number, date_event, event, lesion)

# IMPORTANT: candidate events are filtered to the valid post-NEDLAME window
# BEFORE joining to cohort, not after. Filtering after a left_join would
# silently DROP a cow entirely (rather than correctly censoring her)
# whenever she has trim/LAME events somewhere in her history but none
# inside the valid window - an earlier version of this code had exactly
# that bug and was quietly excluding ~45% of cows from the "any event"
# curve (and a different ~25% from the "with lesion" curve), which is
# also why the two curves looked identical - they weren't being computed
# over the same denominator.
fxn_time_to_next <- function(candidate_events) {
  valid_candidates <- candidate_events |>
    inner_join(
      cohort |> select(id_animal, lact_number, first_nedlame_date, date_lact_end),
      by = c("id_animal", "lact_number")
    ) |>
    filter(date_event > first_nedlame_date, date_event <= date_lact_end) |>
    select(id_animal, lact_number, date_event)

  cohort |>
    left_join(valid_candidates, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
    group_by(id_animal, lact_number) |>
    slice_min(date_event, n = 1, with_ties = FALSE, na_rm = FALSE) |>
    ungroup() |>
    mutate(
      event_occurred = as.integer(!is.na(date_event)),
      days_to_event = if_else(
        event_occurred == 1,
        as.numeric(date_event - first_nedlame_date),
        as.numeric(date_lact_end - first_nedlame_date)
      )
    )
}

q2_any <- fxn_time_to_next(next_events)

next_events_lesion <- next_events |>
  left_join(lesion_lookup, by = c("id_animal", "lact_number", "date_event", "event")) |>
  filter(!is.na(lesion) & lesion == 1)

q2_lesion <- fxn_time_to_next(next_events_lesion)
```

::: panel-tabset
## Time to Any Trim/LAME Event

```{r}
#| label: q2-any-plot
#| fig-height: 5

km_any <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_any)

km_any |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-any-table

km_any |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Trim/LAME Event")
```

## Time to a Trim/LAME Event *With a Lesion*

```{r}
#| label: q2-lesion-plot
#| fig-height: 5

km_lesion <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = q2_lesion)

km_lesion |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Lesion Found") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q2-lesion-table

km_lesion |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% With a Subsequent Lesion Found")
```
:::

## Trim Outcomes: How Many Found Nothing?

Of the cows who had a first post-NEDLAME LAME/FOOTRIM/TRIM event (the "any" curve above), what fraction of those visits were trim-only (`trimonly` - no lesion coded) vs. an actual lesion found - by treatment group.

```{r}
#| label: q2-trimonly-summary

q2_first_event_detail <- q2_any |>
  filter(event_occurred == 1) |>
  left_join(lame_data |> select(id_animal, lact_number, date_event, event, trimonly),
            by = c("id_animal", "lact_number", "date_event"))

q2_first_event_detail |>
  group_by(tx_group) |>
  summarize(n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100, .groups = "drop") |>
  bind_rows(
    q2_first_event_detail |> summarize(tx_group = "Overall", n = n(), n_trimonly = sum(trimonly, na.rm = TRUE), pct_trimonly = mean(trimonly, na.rm = TRUE) * 100)
  ) |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_trimonly, decimals = 1) |>
  cols_label(n = "N Cows With a First Event", n_trimonly = "Trim-Only (No Lesion)", pct_trimonly = "% Trim-Only") |>
  tab_header(title = "First Post-NEDLAME Visit Outcome")
```

# Q3: % of Cows Independently Identified by Farm Staff

`CHKLAME` is the event that represents farm staff independently flagging a lame cow (distinct from the automated `NEDLAME` alert). A cow counts as **staff-identified** here if she has a `CHKLAME` event, in the same lactation, either (a) in the 7 days before her first NEDLAME alert **or on the same day as it**, or (b) any time after it — excluding cows whose post-alert `CHKLAME` came at/after she'd already been trimmed, since a trim-triggered exam isn't independent detection.

```{r}
#| label: q3-build

chklame <- events_formatted |>
  filter(event == "CHKLAME") |>
  select(id_animal, lact_number, date_event) |>
  distinct()

trims <- events_formatted |>
  filter(event %in% c("FOOTRIM", "TRIM")) |>
  select(id_animal, lact_number, date_event) |>
  distinct()

# A CHKLAME on the SAME day as the NEDLAME alert counts as staff-identified
# too (folded into the "pre-week" bucket via <= rather than <), rather than
# falling through the gap between "before" (strict <) and "after" (strict >).
pre_week <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event >= first_nedlame_date - 7, date_event <= first_nedlame_date) |>
  distinct(id_animal, lact_number)

first_post_chklame <- cohort |>
  inner_join(chklame, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  rename(chklame_date = date_event)

first_post_trim <- cohort |>
  inner_join(trims, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  filter(date_event > first_nedlame_date) |>
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(id_animal, lact_number, trim_date = date_event)

post_flag <- first_post_chklame |>
  left_join(first_post_trim, by = c("id_animal", "lact_number")) |>
  filter(is.na(trim_date) | chklame_date < trim_date) |>
  distinct(id_animal, lact_number)

staff_id <- cohort |>
  mutate(
    staff_identified = as.integer(
      paste(id_animal, lact_number) %in% paste(pre_week$id_animal, pre_week$lact_number) |
      paste(id_animal, lact_number) %in% paste(post_flag$id_animal, post_flag$lact_number)
    )
  )
```

::: panel-tabset
## By Treatment Group

```{r}
#| label: q3-table

q3_summary <- staff_id |>
  group_by(tx_group) |>
  summarize(n = n(), n_staff_id = sum(staff_identified), pct_staff_id = mean(staff_identified) * 100, .groups = "drop")

q3_test <- fisher.test(table(staff_id$tx_group, staff_id$staff_identified))

q3_summary |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_staff_id, decimals = 1) |>
  cols_label(n = "N Cows", n_staff_id = "Staff-Identified", pct_staff_id = "% Staff-Identified") |>
  tab_header(title = "Independent Staff Identification (CHKLAME)") |>
  tab_source_note(source_note = paste0("Fisher's exact test p = ", signif(q3_test$p.value, 3)))
```

## By Alert Type

Does staff catch a cow on their own more often when the Nedap alert itself was more severe (`Strong Decline`) vs milder (`Low Decline`/`Low`)?

```{r}
#| label: q3-alert-type-by-group-plot
#| fig-height: 4

q3_alert_test <- fisher.test(table(staff_id$alert_type, staff_id$staff_identified))

staff_id |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type) |>
  summarize(pct_staff_id = mean(staff_identified) * 100, n = n(), .groups = "drop") |>
  ggplot(aes(x = alert_type, y = pct_staff_id, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(round(pct_staff_id, 1), "%\n(n=", n, ")")),
            position = position_dodge(width = 0.7), vjust = -0.2, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal() +
  labs(x = "Alert Type", y = "% Staff-Identified (CHKLAME)", fill = "Treatment Group",
       caption = paste0("Fisher's exact test (alert type vs. staff-identified, pooled across treatment groups): p = ", signif(q3_alert_test$p.value, 3))) +
  theme(legend.position = "bottom")
```
:::

# Q4: % of Post-June-1 Lesion Cases NEDAP Didn't Catch First

Among cows with a `LAME`-diagnosed lesion on or after `r params$history_after_date`, this checks whether that cow had a `NEDLAME` alert in the preceding lookback window (same lactation). A "miss" means staff/routine trimming found a lesion that NEDAP hadn't already flagged. Each cow is counted once here, at her **first** qualifying lesion diagnosis on/after that date (consistent with using the first event throughout this report) — a cow diagnosed lame twice after June 1 doesn't get double-counted.

Lesions diagnosed **after a cow was already dried off** are excluded entirely: NEDLAME's mobility/activity monitoring is production-linked, so a dry-cow lameness case isn't a fair "miss" to hold against it either way.

NEDLAME alerting only went live on `r format(nedlame_start_date)` (per the DairyComp setup), so a fixed `r params$q4_lookback_days`-day lookback isn't fair to lesion events found in the first few weeks after go-live — there simply wasn't `r params$q4_lookback_days` days of NEDLAME history yet for NEDAP to have caught anything in. The lookback window here is **rolling**: for each lesion event it's capped at however many days NEDLAME had actually been running by that date (never more than `r params$q4_lookback_days`, but shorter for events soon after go-live), so early cases are judged against the detection window that could actually have existed, not a fixed window reaching back before the system was on.

```{r}
#| label: q4-build

lame_lesion_recent_all <- lame_data |>
  filter(event == "LAME", lesion == 1, date_event >= ymd(params$history_after_date)) |>
  left_join(animal_lactations |> select(id_animal, lact_number, date_dry), by = c("id_animal", "lact_number"))

# exclude lesions diagnosed after the cow was already dried off
n_excluded_dry <- sum(!is.na(lame_lesion_recent_all$date_dry) & lame_lesion_recent_all$date_event > lame_lesion_recent_all$date_dry)

lame_lesion_recent <- lame_lesion_recent_all |>
  filter(is.na(date_dry) | date_event <= date_dry) |>
  select(-date_dry) |>
  # first qualifying lesion case per cow-lactation only
  group_by(id_animal, lact_number) |>
  slice_min(date_event, n = 1, with_ties = FALSE) |>
  ungroup()

nedlame_all <- events_all |>
  filter(Event == "NEDLAME") |>
  select(id_animal, lact_number, ned_date = date_event) |>
  distinct()

q4 <- lame_lesion_recent |>
  mutate(
    lookback_days_available = pmax(as.numeric(date_event - nedlame_start_date), 0),
    lookback_days_used = pmin(params$q4_lookback_days, lookback_days_available)
  ) |>
  left_join(nedlame_all, by = c("id_animal", "lact_number"), relationship = "many-to-many") |>
  mutate(lookback_hit = !is.na(ned_date) & ned_date <= date_event & ned_date >= date_event - lookback_days_used) |>
  group_by(id_animal, lact_number, date_event, lookback_days_used) |>
  summarize(caught_by_nedap = any(lookback_hit), .groups = "drop") |>
  left_join(cohort |> select(id_animal, lact_number, tx_group), by = c("id_animal", "lact_number")) |>
  mutate(tx_group = if_else(is.na(tx_group), "Never Alerted (Not in Pilot Cohort)", tx_group))

n_truncated_lookback <- sum(q4$lookback_days_used < params$q4_lookback_days)
```

`r nrow(q4)` qualifying lesion cases on/after `r params$history_after_date` (one per cow, first diagnosis only; `r n_excluded_dry` additional cases were excluded as diagnosed after dry-off), of which `r n_truncated_lookback` had a lookback window shorter than the full `r params$q4_lookback_days` days (because they fell within `r params$q4_lookback_days` days of NEDLAME go-live).

```{r}
#| label: q4-table

q4 |>
  group_by(tx_group) |>
  summarize(n_lesion_events = n(), pct_missed = mean(!caught_by_nedap) * 100, .groups = "drop") |>
  bind_rows(
    q4 |> summarize(tx_group = "Overall", n_lesion_events = n(), pct_missed = mean(!caught_by_nedap) * 100)
  ) |>
  gt(rowname_col = "tx_group") |>
  fmt_number(columns = pct_missed, decimals = 1) |>
  cols_label(n_lesion_events = "N Lesion Cases", pct_missed = "% Not Caught by NEDAP First") |>
  tab_header(title = paste("Lesion Cases On/After", params$history_after_date)) |>
  tab_source_note(source_note = "\"Never Alerted (Not in Pilot Cohort)\" = cows with a lesion case here that never received a NEDLAME alert at all (not just a late one) - the most complete misses, distinct from TX/Control cows who were alerted but not in time.")
```

## What Lesions Is NEDAP Missing?

Lesion types found at these first-diagnosis cases, broken out by NEDAP detection status:

- **Never Alerted (Not in Pilot Cohort)** — this cow never received a NEDLAME alert at all, in the whole dataset. The most complete kind of miss.
- **Alerted, Missed** — this cow did get a NEDLAME alert at some point (she's in the TX or Control cohort), but not within the lookback window before *this* lesion was found. NEDAP was watching her, just not in time for this case.
- **Alerted, Caught in Time** — a NEDLAME alert fired within the lookback window before this lesion was diagnosed; NEDAP flagged it first.

If a cow had more than one distinct lesion type diagnosed on the same day (e.g. white line disease *and* digital dermatitis at the same exam), each distinct type is counted separately here; the same lesion type coded more than once on one day only counts once.

```{r}
#| label: q4-lesion-types

lesion_type_cols <- c("dd", "footrot", "wld", "sole_ulcer", "injury", "cork",
                       "hemorrhage", "sole_fracture", "toe_ulcer", "thin", "other")

q4_group <- q4 |>
  mutate(detection_group = if_else(tx_group == "Never Alerted (Not in Pilot Cohort)",
                                    "Never Alerted (Not in Pilot Cohort)",
                                    if_else(caught_by_nedap, "Alerted, Caught in Time", "Alerted, Missed")))

q4_lesion_types <- lame_lesion_recent |>
  select(id_animal, lact_number, date_event, all_of(lesion_type_cols)) |>
  inner_join(q4_group |> select(id_animal, lact_number, date_event, detection_group),
             by = c("id_animal", "lact_number", "date_event")) |>
  pivot_longer(cols = all_of(lesion_type_cols), names_to = "lesion_type", values_to = "has_lesion") |>
  filter(has_lesion == 1) |>
  count(detection_group, lesion_type)

ggplot(q4_lesion_types, aes(x = fct_reorder(lesion_type, n, .fun = sum), y = n, fill = detection_group)) +
  geom_col() +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = NULL, y = "Number of Lesion Cases", fill = "NEDAP Detection Status") +
  theme(legend.position = "bottom")
```

# Q5: Influence of Lameness History

Each cow's lameness-history label (`New` / `Repeat` / `Chronic`, from the same chronicity classifier `report_explore_lame_new.qmd` uses) reflects her **full lifetime record** as of her first NEDLAME alert — `lame_data` goes back to 2021, and this deliberately isn't restricted to her current lactation, since chronicity is a lifetime concept. A cow with no LAME/TRIM/FOOTRIM diagnosis anywhere in her life before her alert (still about a third of the cohort even with 5 years of history to draw on) is folded into `New`, the closest existing category to "no established chronic pattern" — there's no separate "never diagnosed" bucket. With only `r nrow(cohort)` cows split across 2 treatment arms x 3 history categories, cells here are thin (some under 20) — treat this section as descriptive/exploratory, not a confirmatory subgroup test. The milk-trend and time-to-event plots below are faceted by **alert type** rather than history group, since that split turned out more informative; history's own distribution is shown just below.

```{r}
#| label: q5-balance-plot
#| fig-height: 4

cohort |>
  count(tx_group, history_group) |>
  group_by(tx_group) |>
  mutate(pct = n / sum(n) * 100) |>
  ungroup() |>
  mutate(history_group = fct_relabel(history_group, \(x) str_wrap(x, width = 14))) |>
  ggplot(aes(x = history_group, y = n, fill = tx_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = paste0(n, " (", round(pct, 0), "%)")),
            position = position_dodge(width = 0.7), vjust = -0.3, size = 3) +
  scale_fill_viridis_d() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal() +
  labs(x = NULL, y = "N Cows", fill = "Treatment Group",
       title = "Lameness-History Distribution at Enrollment") +
  theme(legend.position = "bottom",
        axis.text.x = element_text(lineheight = 0.9),
        plot.margin = margin(10, 15, 10, 15))
```

## Milk Trend by Alert Type

```{r}
#| label: q5-milk-plot
#| fig-height: 6

milk_norm |>
  mutate(alert_type = factor(alert_type, levels = c("LOW", "LOW DECLINE", "STRONG DECLINE"))) |>
  group_by(tx_group, alert_type, day_rel) |>
  summarize(mean_pct = mean(pct_of_baseline, na.rm = TRUE), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = day_rel, y = mean_pct, color = tx_group)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Days Relative to First NEDLAME Alert", y = "Milk Production (% of Baseline)", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

## Time to Next Trim/LAME Event by Alert Type

```{r}
#| label: q5-km-plot
#| fig-height: 6

alert_levels <- c("LOW", "LOW DECLINE", "STRONG DECLINE")

# ggsurvfit only exposes a single combined `strata` column, so a KM plot
# faceted by alert_type is built manually: fit one survfit2() per
# alert level (skipping any level without both treatment arms
# represented) and stack the tidy() estimates before plotting.
fittable_alert_levels <- alert_levels |>
  keep(\(a) {
    d <- q2_any |> filter(alert_type == a)
    n_distinct(d$tx_group) == 2 && nrow(d) >= 10
  })

q5_km_data <- fittable_alert_levels |>
  map(\(a) {
    d <- q2_any |> filter(alert_type == a)
    fit <- survfit2(Surv(days_to_event, event_occurred) ~ tx_group, data = d)
    broom::tidy(fit) |>
      mutate(
        tx_group = str_remove(strata, "^tx_group="),
        alert_type = a,
        risk = 1 - estimate
      )
  }) |>
  list_rbind() |>
  mutate(alert_type = factor(alert_type, levels = alert_levels))

skipped_alert_levels <- setdiff(alert_levels, fittable_alert_levels)

ggplot(q5_km_data, aes(x = time, y = risk, color = tx_group)) +
  geom_step(linewidth = 1) +
  facet_wrap(~alert_type) +
  scale_color_viridis_d() +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  labs(x = "Days After First NEDLAME Alert", y = "% With a Subsequent Trim/LAME Event", color = "Treatment Group") +
  theme(legend.position = "bottom")
```

`r if (length(skipped_alert_levels) > 0) paste0("*Alert type(s) omitted for insufficient data in one or both arms: ", paste(skipped_alert_levels, collapse = ", "), ".*")`

# Q6: Time to Culling After NEDAP Enrollment

Same culling-outcome pattern as `report_explore_lame_new.qmd`, anchored on the first NEDLAME date instead of a lesion-event date.

```{r}
#| label: q6-build

q6 <- cohort |>
  left_join(animals |> select(id_animal, date_sold, date_died, date_left), by = "id_animal") |>
  mutate(
    culled_ever = as.integer(!is.na(date_left)),
    days_to_censor_cull = if_else(
      culled_ever == 1,
      as.numeric(date_left - first_nedlame_date),
      as.numeric(date_max_pull - first_nedlame_date)
    )
  ) |>
  filter(days_to_censor_cull >= 0)

km_cull <- survfit2(Surv(days_to_censor_cull, culled_ever) ~ tx_group, data = q6)
```

```{r}
#| label: q6-plot
#| fig-height: 5

km_cull |>
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  scale_ggsurvfit() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  labs(x = "Days After First NEDLAME Alert", y = "% Culled") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: q6-table

km_cull |>
  tbl_survfit(times = c(7, 14, 30), label = " ", type = "risk", label_header = "{time} Days") |>
  as_gt() |>
  tab_header(title = "% Culled After First NEDLAME Alert")
```

# Caveats

- **MNFRS coding is now confirmed from the DairyComp setup documentation**, not just inferred: the `NDTLIST` command explicitly enrolls `MNFRS = 1` cows onto the trim list ("put nedap cows on trim list"), matching the empirical pattern found here (`MNFRS = 1` cows get trimmed within 14 days of their alert 81.5% of the time vs 42.9% for `MNFRS = 2`).
- **Two data-processing bugs were found and fixed while building this report**, both of which materially changed earlier draft results:
  1. The externally-fetched `fxn_collapse_lesions()`/`fxn_trim_vars()`/`fxn_dz_status()` functions (pulled live from GitHub) come bundled with their *own* `fxn_code_lesions()`, whose `trimonly` logic only checks the `protocols` field for "NONE" - this farm's routine/no-lesion trims are coded "NONE" in the `remark` field instead, which that version doesn't check. The GitHub function was silently overwriting this repo's own (correctly-calibrated) `functions/fxn_code_lesions.R`, so *every* LAME/FOOTRIM/TRIM row was being misclassified as having a lesion. Fixed by sourcing the local file last so it wins. This is why Q2's "any trim" and "with a lesion" curves were identical in an earlier draft.
  2. Q2's time-to-next-event join was filtering candidate events to the valid post-NEDLAME window *after* joining to the cohort rather than before, which silently dropped a cow from the analysis entirely (instead of correctly counting her as censored) whenever she had trim/LAME events somewhere in her history but none inside the valid window. This was quietly excluding ~45% of cows from the "any event" curve.
- **Short history**: `NEDLAME` only spans `r as.numeric(max(first_nedlame$first_nedlame_date) - min(first_nedlame$first_nedlame_date))` days of data (go-live `r format(nedlame_start_date)`), capping every time-to-event and milk-trend window above — results should be treated as an early read, not a mature-pilot conclusion. Q4's lookback window is rolling for this same reason (see Q4).
- **Small sample for Q5**: 2 treatment arms x 4 history categories over `r nrow(cohort)` cows produces thin cells; the two arms are not perfectly balanced on prior lameness history at enrollment either (see the Q5 balance chart), which should temper causal interpretation of any history-stratified differences.
- **Milk join resolution**: `r round(100 * n_milk_cows / nrow(cohort), 1)`% of cohort cows had a usable milk baseline in the analysis window; the remainder lacked resolvable milk records (e.g. dried off, sold, or outside the pulled milk-file date range).
